### 2020-05-15 manuscript/PLOS-MSc-Thesis.tex
add manuscript

% Template for PLoS
% Version 3.4 January 2017
%
% % % % % % % % % % % % % % % % % % % % % %
%
% -- IMPORTANT NOTE
%
% This template contains comments intended 
% to minimize problems and delays during our production 
% process. Please follow the template instructions
% whenever possible.
%
% % % % % % % % % % % % % % % % % % % % % % % 
%
% Once your paper is accepted for publication, 
% PLEASE REMOVE ALL TRACKED CHANGES in this file 
% and leave only the final text of your manuscript. 
% PLOS recommends the use of latexdiff to track changes during review, as this will help to maintain a clean tex file.
% Visit https://www.ctan.org/pkg/latexdiff?lang=en for info or contact us at latex@plos.org.
%
%
% There are no restrictions on package use within the LaTeX files except that 
% no packages listed in the template may be deleted.
%
% Please do not include colors or graphics in the text.
%
% The manuscript LaTeX source should be contained within a single file (do not use \input, \externaldocument, or similar commands).
%
% % % % % % % % % % % % % % % % % % % % % % %
%
% -- FIGURES AND TABLES
%
% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
%
% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
% - Figures should be uploaded separately from your manuscript file. 
% - Figures generated using LaTeX should be extracted and removed from the PDF before submission. 
% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
% For figure citations, please use "Fig" instead of "Figure".
% See http://journals.plos.org/plosone/s/figures for PLOS figure guidelines.
%
% Tables should be cell-based and may not contain:
% - spacing/line breaks within cells to alter layout or alignment
% - do not nest tabular environments (no tabular environments within tabular environments)
% - no graphics or colored text (cell background color/shading OK)
% See http://journals.plos.org/plosone/s/tables for table guidelines.
%
% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
%
% % % % % % % % % % % % % % % % % % % % % % % %
%
% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
%
% IMPORTANT
% Below are a few tips to help format your equations and other special characters according to our specifications. For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://journals.plos.org/plosone/s/latex
%
% For inline equations, please be sure to include all portions of an equation in the math environment.  For example, x$^2$ is incorrect; this should be formatted as $x^2$ (or $\mathrm{x}^2$ if the romanized font is desired).
%
% Do not include text that is not math in the math environment. For example, CO2 should be written as CO\textsubscript{2} instead of CO$_2$.
%
% Please add line breaks to long display equations when possible in order to fit size of the column. 
%
% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
%
% When adding superscript or subscripts outside of brackets/braces, please group using {}.  For example, change "[U(D,E,\gamma)]^2" to "{[U(D,E,\gamma)]}^2". 
%
% Do not use \cal for caligraphic font.  Instead, use \mathcal{}
%
% % % % % % % % % % % % % % % % % % % % % % % % 
%
% Please contact latex@plos.org with any questions.
%
% % % % % % % % % % % % % % % % % % % % % % % %

\documentclass[10pt,letterpaper]{article}
\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}

% amsmath and amssymb packages, useful for mathematical formulas and symbols
\usepackage{amsmath,amssymb}

% Use adjustwidth environment to exceed column width (see example table in text)
\usepackage{changepage}

% Use Unicode characters when possible
\usepackage[utf8x]{inputenc}

% textcomp package and marvosym package for additional characters
\usepackage{textcomp,marvosym}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

% Use nameref to cite supporting information files (see Supporting Information section for more info)
\usepackage{nameref,hyperref}

% line numbers
\usepackage[right]{lineno}

% ligatures disabled
\usepackage{microtype}
\DisableLigatures[f]{encoding = *, family = * }

% color can be used to apply background shading to table cells only
\usepackage[table]{xcolor}

% array package and thick rules for tables
\usepackage{array}

% create "+" rule type for thick vertical lines
\newcolumntype{+}{!{\vrule width 2pt}}

% create \thickcline for thick horizontal lines of variable length
\newlength\savedwidth
\newcommand\thickcline[1]{%
  \noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
  \cline{#1}%
  \noalign{\vskip\arrayrulewidth}%
  \noalign{\global\arrayrulewidth\savedwidth}%
}

% \thickhline command for thick horizontal lines that span the table
\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
\hline
\noalign{\global\arrayrulewidth\savedwidth}}

% Remove comment for double spacing
%\usepackage{setspace} 
%\doublespacing

% Text layout
\raggedright
\setlength{\parindent}{0.5cm}
\textwidth 5.25in 
\textheight 8.75in

% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
% Captions will be left justified
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
\renewcommand{\figurename}{Fig}

% Use the PLoS provided BiBTeX style
\bibliographystyle{plos2015}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Leave date blank
\date{}

% Header and Footer with logo
\usepackage{lastpage,fancyhdr,graphicx}
\usepackage{epstopdf}
\pagestyle{myheadings}
\pagestyle{fancy}
\fancyhf{}
\setlength{\headheight}{27.023pt}
\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
\fancyheadoffset[L]{2.25in}
\fancyfootoffset[L]{2.25in}
\lfoot{\sf PLOS}

%% Include all macros below

\newcommand{\lorem}{{\bf LOREM}}
\newcommand{\ipsum}{{\bf IPSUM}}

%% END MACROS SECTION

\begin{document}
\vspace*{0.2in}

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Gene regulatory networks producing a striped band of gene 
expression} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Juan Camilo Arboleda-Rivera\textsuperscript{1\Yinyang},
Gloria Machado-Rodríguez\textsuperscript{1\Yinyang},
Boris Rodríguez-Rey\textsuperscript{2\textcurrency},
Jayson Gutiérrez\textsuperscript{3},
% Name5 Surname\textsuperscript{2\ddag},
% Name6 Surname\textsuperscript{2\ddag},
% Name7 Surname\textsuperscript{1,2,3*},
% with the Lorem Ipsum Consortium\textsuperscript{\textpilcrow}
\\
\bigskip
\textbf{1} Instituto de Biología, Universidad de Antioquia, Medellín, Antioquia, Colombia
\\
\textbf{2} Instituto de Física, Universidad de Antioquia, Medellín, Antioquia, Colombia
\\
\textbf{3} Affiliation Dept/Program/Center, Ghent University, Ghent, State, Belgium
% \textbf{3} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\bigskip

% Insert additional author notes using the symbols described below. Insert symbol callouts after author names as necessary.
% 
% Remove or comment out the author notes below if they aren't used.
%
% Primary Equal Contribution Note
\Yinyang These authors contributed equally to this work.

% Additional Equal Contribution Note
% Also use this double-dagger symbol for special authorship notes, such as senior authorship.
\ddag These authors also contributed equally to this work.

% Current address notes
\textcurrency Current Address: Dept/Program/Center, Universidad de Antioquia, 
Medellín, Antioquia, Colombia % change symbol to "\textcurrency a" if more than one current address note
% \textcurrency b Insert second current address 
% \textcurrency c Insert third current address

% Deceased author note
% \dag Deceased

% Group/Consortium Author Note
\textpilcrow Membership list can be found in the Acknowledgments section.

% Use the asterisk to denote corresponding authorship and provide email address in note below.
* juan.arboleda2@udea.edu.co

\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur eget porta 
erat. Morbi consectetur est vel gravida pretium. Suspendisse ut dui eu ante 
cursus gravida non sed sem. Nullam sapien tellus, commodo id velit id, eleifend 
volutpat quam. Phasellus mauris velit, dapibus finibus elementum vel, pulvinar 
non tellus. Nunc pellentesque pretium diam, quis maximus dolor faucibus id. Nunc
convallis sodales ante, ut ullamcorper est egestas vitae. Nam sit amet enim 
ultrices, ultrices elit pulvinar, volutpat risus.

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLOS ONE authors please skip this step. 
% Author Summary not valid for PLOS ONE submissions.   
\section*{Author summary}
Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur eget porta 
erat. Morbi consectetur est vel gravida pretium. Suspendisse ut dui eu ante 
cursus gravida non sed sem. Nullam sapien tellus, commodo id velit id, eleifend 
volutpat quam. Phasellus mauris velit, dapibus finibus elementum vel, pulvinar 
non tellus. Nunc pellentesque pretium diam, quis maximus dolor faucibus id. Nunc
convallis sodales ante, ut ullamcorper est egestas vitae. Nam sit amet enim 
ultrices, ultrices elit pulvinar, volutpat risus.

\linenumbers

% Use "Eq" instead of "Equation" for equation citations.
\section*{Introduction}

Cells interact continuously with their environment and they must detect signals
from it. In many cases Gene Regulatory Networks (GRNs) interpret these signals
through gene interaction and produce a particular response. This information
processing function is of particular importance in developmental processes such
as those in which a group of cells differentiate in response to a signalling
molecule such as a morphogen \cite{Turing1952, Wolpert1969}. However, as
information processing by GRNs is often a complex process, one major goal of
developmental biology is to understand mechanistically how positional
information is translated in spatial differentiation.\\

Our understanding of how GRNs process information has increased thanks to the
concept and theory of network motifs, which are defined as patterns of
interconnections occurring in networks at numbers that are significantly higher
than those in randomized networks \cite{Milo2002}. This novel approach, along
with mathematical modeling and computational methods have been shown to be a
powerful tool to study GRNs and the predictions of studies using these methods
have been validated by experimental studies
\cite{Kalir2005,Kalir2004,Mangan2003a,ODonnell2005,Basu2005}.

However, although several studies have analyzed the properties of GRNs that
process positional information, they have focused on particular natural networks
\cite{Schaerli2014,Jaeger2004}, in synthetic networks
\cite{Basu2005,Reinitz1995,Schaerli2018}, or have constrained the design of the
GRNs \cite{Elowitz2000}.

In this study, we used a set of 2061 three-node GRNs selected by a genetic
algorithm to interpret a morphogen signal and respond to it generating a
striped pattern of gene expression in order to analyse the role of network 
motifs in this process and its dynamic properties.

% Lorem ipsum dolor sit~\cite{bib1} amet, consectetur adipiscing elit. Nullam Eq~(\ref{eq:schemeP}) sapien tellus, commodo id velit id, eleifend volutpat quam. Nunc pellentesque pretium diam, quis maximus dolor faucibus id.~\cite{bib2} Nunc convallis sodales ante, ut ullamcorper est egestas vitae.

% Equation provided by the LaTeX template.
% \begin{eqnarray}
% \label{eq:schemeP}
% 	\mathrm{P_Y} = \underbrace{H(Y_n) - H(Y_n|\mathbf{V}^{Y}_{n})}_{S_Y} + \underbrace{H(Y_n|\mathbf{V}^{Y}_{n})- H(Y_n|\mathbf{V}^{X,Y}_{n})}_{T_{X\rightarrow Y}},
% \end{eqnarray}

\section*{Materials and methods}
\subsection*{Gene Regulatory Networks and morphogenetic field}

% For figure citations, please use "Fig" instead of "Figure".
In order to study three-node GRNs we represented them by three sets of real 
numbers, the first set was composed of the interaction values
between genes in the network; the second of the diffusion rates (D) of each of 
the three gene products; and the third of the degradation rates ($\delta$) of 
these same gene products, with $D$ and $\delta  \in [0, 0.05]$.\\

The set of interaction values was represented as adjacency matrices $W \in 
\mathbb{R}^{3\times4}$ where $i$ represents a regulated node of the GRN and $j$ 
represents a regulator node of the GRN Fig \ref{fig:adjmat} with 
$w_{\textit{ij}}~\in [-10,\ 10]$. Values in these matrices can be negative, zero
or positive and represent represion, no interaction, or activation respectively.
The magnitude of the value is proportional to the strength of the interaction.\\

\begin{figure}[!h]
%  \centering
 \includegraphics[width=\textwidth]{figures/metodos/matriz-de-adyacencia}
    \caption{\bf Modelling of GRNs.}
    A: GRN represented as a directed graph with activation (blue arrows) and 
    inhibition (red arrows) interactions. Black circles
    tagged with a \textbf{M} represent the morphogen; \textbf{A}, \textbf{B} and
    \textbf{C} represent genes in the network.
    B: The same GRN represented as an adjacency matrix in which positive values 
    represent activation, negative ones represent inhibition and
    zeros represent no interaction.
 \label{fig:adjmat}
\end{figure}

In our model the morphogen could interact with any of the three genes on the 
GRN, but could not be affected by them. The morphogenetic field was defined by 
an unidimensional array of 30 isogenic cells exposed to the morphogen 
concentration gradient (Fig~\ref{fig:morfgrad}). The initial concentration of 
each gen product in all cells was set to 0.1 in all simulations.\\

\begin{figure}[!h]
 \includegraphics[width=\textwidth]{figures/metodos/morphogen-gradient}
 \caption{\bf Morphogenetic field.}
 In our model the field was composed of a linear array of 30 isogenic cells and 
 a morphogen gradient concentration described by an exponential decay function. 
 Horizontal arrows between cells represent diffusion of gene products between 
 adjacent cells.
 \label{fig:morfgrad}
\end{figure}

\subsection*{Mathematical model}

Our model is a modification of the model used by Cotterell \& Sharpe 
\cite{Cotterell2010} and proposed by Reinitz et al. \cite{Reinitz1995}. The 
model is a dynamical system that describes the change in concentration of gene 
product \emph{i} in time, as shown in Eq~\ref{ec:modelo-gral}.

 \begin{eqnarray}
  \frac{d[G]^i_n}{dt}
  = g(u^i) + D_i \cdot [ ([G]^i_{n-1}-[G]^i_n) +([G]^i_{n+1}-[G]^i_n)]-\delta_i 
  \cdot [ G ]^i_n ,
  \label{ec:modelo-gral}
 \end{eqnarray}

in which $[G]^{i}_{n}$ is the concentration of the i-th gene in the \emph{n}-th 
cell ($[G]^{i}_{n} ≥ 0$), $g(u^i)$ is a function describing the relationship 
between the interactions on the \emph{i}-th gene and its expression 
Eq~(\ref{ec:input-func}) and described in more detail below. $D_i$ and 
$\delta_i$ are the diffusion and the degradation rate of \emph{i}-th gene 
product, respectively.\\

The input function is a sigmoid function described by Eq~\ref{ec:input-func}.

\begin{equation}
 g(u^i) = \frac{1}{1 + \exp(a - b \cdot u^i)},
 \label{ec:input-func}
\end{equation}

in which $a$ is the sigmoid steepness, equal to 5; $a/b$ is the threshold value,
set to 1 for all simulations, and $u^i$ is the following equation:

\begin{equation}
 u^i = \sum_j w_{ij} \cdot [G]^j_n + w_{im} \cdot [M]_n.
 \label{ec:mat-sum}
\end{equation}

This equation sums the interactions acting upon the \emph{i}-th gene, being 
$w_{ij}$ the interaction strength of the \emph{j}-th gene upon the \emph{i}-th 
gene and $w_{im}$ the interaction strength of the morphogen upon the \emph{i}-th
gene. $[G]^j_n$ and $[M]_n$ are the concentrations of the \emph{j}-th gene 
product and the morphogen in the \emph{n}-th cell.

\subsection*{Morphogen spatial distribution}

The morphogen concentration along the morphogenetic field is described by 
Eq~\ref{ec:morph}.

\begin{equation}
 M = A_0 \cdot \exp(-c/h),
 \label{ec:morph}
\end{equation}

were $A_0$ is the concentration of the morphogen in the position zero of the 
morphogenetic field and was set to 1 in our experiments; $c$ is the cell index, 
defined as the ratio between the \emph{n}-th cell from left to right and the 
total number of cells, and $h$ is a decay parameter, whose value in our model 
was set to 0.4.

The phenotype of a GRN was defined as the expression pattern of each gene along 
the morphogenetic field after 500 time steps of integration of the dynamical 
system. We chose that number of time steps because numerical experiments showed 
that GRNs reached the steady state in approximately 300 steps.

\subsection*{Optimal pattern definition}

The optimal pattern of gene expression defined by us, consisted in cells at the 
sides of the field ($n<11$ and $n>20$) displaying expression levels lower than 
10\% of the maximal level observed along the field of cells for the output gene,
and cells at the middle of the field ($n ∈ [11,20]$) displaying expression 
levels greater than 90\% of the maximal level observed along the entire field 
for the output gene (Fig~\ref{fig:exp-pattern}).

\begin{figure}[!h]
 \includegraphics[width=\textwidth]{figures/metodos/optimal-pattern}
 \caption{\bf Optimal gene expression pattern.}
 \label{fig:exp-pattern}
\end{figure}

\subsection*{Genetic algorithm}

The genetic algorithm used to produce the set of gene regulatory networks (GRNs)
consisted in the following steps:

\begin{enumerate}
 \item{\bf Generate a set of 200 random GRNs.}

 \item{\bf Numerically solve the dynamical system for each GRN from t=0 until 
 t=500.} We integrated the differential equation using the function NDSolve from
 Wolfram Mathematica 11 with the option “EquationSimplification” and the 
 “Residual” simplification method.

 \item{\bf Evaluate fitness by comparing the phenotypes with the optimal 
 phenotype (Fig~\ref{fig:exp-pattern}).} In order to compute the fitness 
 function for each GRN we used three different filters. The first filter 
 assesses whether the expression profile of the output node reaches a 
 quasi-steady state:

 \begin{equation}
  S_{filter} = \left( \frac{1}{I \cdot N}\right) \sqrt{\sum_{i=1}^{I=3}
  \sum_{n=1}^{N=30} ([G]_n^i(t=500) - [G]^i_n(t=250))^2}.
 \end{equation}

 The expression profile was considered to have reached the steady state if 
 $S_{filter} < 0.001$.\\

 The second filter measures the spatial heterogeneity in the field:

 \begin{equation}
  P_{filter} = \left( \frac{1}{I \cdot N} \right) \sqrt{ \sum_{i=3}^{I=3}
  \sum_{n=1}^{N=30} \left( [G]^i_n - \langle [G]^i_n \rangle \right)^2 },
 \end{equation}

 where $\langle [G]^i_n \rangle$ is the mean concentration of the \emph{i}-th 
 gene along the field at $t=500$.\\

 The third filter relates the Manhattan Distance ($D_{obs}$) between an
 expression pattern and the optimal pattern with the maximum distance achievable
 ($D_{max}$):

 \begin{equation}
  \mathit{PF}_{eff} = 1 - \frac{D_{obs}}{D_{max}}.
 \end{equation}

 In this function, the expression profile of the output gene was normalized and
 discretized so that each expression value for each cell in the field was an
 integer ranging from 1 to 10.\\

 The filters mentioned above are integrated in the following functions that 
 asses the quality of a given phenotype:

 \begin{equation}
  Q_1(P_{filter}) = \left( \frac{ {P_{filter}}^{10} }{ {P_{filter}}^{10} + 
  0.1^{10} } \right),
 \end{equation}

 \begin{equation}
  Q_2(S_{filter}) = \left( \frac{ 0.1^2 }{ {S_{filter}}^2 + 0.1^2 } \right).
 \end{equation}

 Finally, these quality functions were used to compute the fitness score with
 the following fitness function:

 \begin{equation}
  F = \mathit{PF_{eff}} \cdot Q_1(P_{filter}) \cdot Q_2(S_{filter}).
 \end{equation}

 \item{\bf Select GRNs with a probability proportional to its fitness.}

 \item{\bf Recombine and mutate GRNs.} Mutations included changes in the
 numerical values of degradation and diffusion parameters and interaction
 strengths, and addition and removal of interactions among genes and between the
 genes and the morphogen.
\end{enumerate}

Steps 2 to 5 were iterated 1000 times, resulting in a set of 2061 GRNs with a
striped pattern of gene expression in at least one gene.

\subsection*{Classification of GRNs}

In order to group all the isomorphic networks in the set of 2061 GRNs we used a
classification algorithm that selected a random GRN, then it generated all 
isomorphs of that GRN by performing permutations in rows and columns. The 
algorithm then compared all six isomorphs with other GRN and classified the 
topologies in the same group if they were isomorphic networks.

\subsection*{Neutral network}

The neutral network was obtained by generating all possible isomorphs of a 
network topology and then calculating the minimal Hamming distance between these
isomorphs and other network topologies. The Hamming distance between two 
topologies was calculated using the following equation:

\begin{equation}
 D_H(W^A, W^B) = \sum_{ij} \mid sgn(w_{ij}^A) - sgn(w^B_{ij}) \mid .
\end{equation}

We adopted the same definition of neighborhood as Cotterell \& Sharpe 
\cite{Cotterell2010}, in which two topologies were neighbors if the gain or 
removal of any one interaction can transform one of the topologies into the 
other.

\subsection*{Robustness analysis}

In order to evaluate the robustness of the different topologies to perturbations
in parameters of interactions between genes we took a GRN belonging to each one 
of the network topologies and we modified each interaction independently by 
increasing and decreasing its value in a 20\%. For topologies with 3 or more 
GRNs we chose the most similar to the mean configuration for that topology.\\

As a measure of the robustness of the topology we chose the proportion of times 
in which a perturbation resulted in a GRN with a fitness greater or equal to 
0.95. Additionally, we considered a topology to be robust if the previous 
measure was greater or equal to 0.5.\\

As a measure of the robustness of subgraphs (see “Classification of topologies”)
we calculated the mean of the robustness measure of all the topologies 
containing a particular subgraph.\\

\subsubsection*{Robustness of topology 9 to perturbations in non-network parameters}

In order to estimate the robustness of topology 9 to changes in parameters that 
control morphogen gradient and changes in initial concentrations of gene 
products, we generated 100 fitness values for each one of the 19 GRNs 
belonging to topology 9; in each of these 100 assays we chose $A_0$, $h$ or 
initial concentrations randomly from a Normal Distribution with $\mu = 1$ for 
$A_0$, $\mu = 0.4$ for $h$ and $\mu = 1$ for initial concentrations and a 
Coefficient of Variation of 30\% in each case.

\subsection*{Shannon entropy}

We calculated the Shannon entropy Eq~(\ref{ec:shannon}) of the network 
topologies distribution in order to test if this distribution could be obtained 
by chance with a probability greater than 0.05. In order to obtain a 95\% 
confidence interval, we generated 30 samples of a set of 2061 random matrices 
and calculated the Shannon entropy in bits for each sample and tested these data
for normality.

\begin{equation}
 H = -\sum p(x) \cdot \log_{2}p(x).
 \label{ec:shannon}
\end{equation}

\subsection*{Classification of GRNs phenotypes}

In order to classify each GRN by its spatiotemporal expression profile (its 
dynamics of expression), we selected expression profiles of GRNs from time step 
0 to time step 250 each 10 time steps, for a total of 26 expression profiles 
over time. These spatiotemporal expression profiles were represented as tensors 
$S \in \mathbb{R}^{3 \times 26 \times 30}$, in which each element ($s_{itn} \geq
0$) is the expression level of the \emph{i}-th gene at time $t$ in cell $n$.\\

Later we calculated the distance between all the pairs of tensors and we 
performed a Neighbor Joining clustering using the package Scikit Bio 0.5.5 from 
Python 3.

\subsection*{Classification of topologies}

In order to classify each topology by its subgraph composition, we first created
a matrix  $T = (t_{ij})$ with 714 rows corresponding to each different topology 
and 17 columns corresponding to different subgraphs. Each element $t_{ij}$ of 
the matrix was 1 if the \emph{j}-th subgraph was present in the \emph{i}-th 
topology or 0 if the subgraph was not present.\\

With this matrix we then calculated a distance matrix in which each element 
consisted in the Hamming distance between the row vectors in matrix $T$ between 
each pair of topologies. We then performed a clustering of these topologies 
using the Neighbor Joining method in the Scikit Bio package.

\subsection*{Complexity index and network descriptors}

As a measure of network complexity, we used the complexity index based on 
Shannon entropy proposed by Bonchev \& Rovray \cite{D.2005}. We calculated this 
index using the following equation:

\begin{equation}
 I_{vd} = \sum_{i=1}^V a_i \cdot \log_{2} a_i,
\end{equation}

being $a_i$ the degree of \emph{i}-th node and $V$ the number of nodes in the 
network. The network descriptors were calculated using built-in functions from 
Wolfram Mathematica software.\\

We performed a Pearson’s chi-square goodness of fit test for the node degree 
distribution using Wolfram Mathematica.

% For placing video, \nameref{S1_Video} vel sagittis arcu lobortis.

% Place figure captions after the first paragraph in which they are cited.
% \begin{figure}[!h]
% \caption{{\bf Bold the figure title.}
% Figure caption text here, please use this space for the figure panel descriptions instead of using subfigure commands. A: Lorem ipsum dolor sit amet. B: Consectetur adipiscing elit.}
% \label{fig1}
% \end{figure}

% Results and Discussion can be combined.
\section*{Results}

The classification of GRNs showed that in our initial set of 2061 GRNs exists a 
number of 714 different network topologies, each one of them containing a 
different number of GRNs (Fig~\ref{fig:distopol}). This abundances distribution 
was not produced by chance (Shannon entropy = 8.52963, 95\% confidence interval 
= [10.8915, 10.9275]), and instead it shows a structuring, indicating the 
possible existence of network motifs.

\begin{figure}[!h]
 \center
 \includegraphics[width=\textwidth]{figures/results/distribution_topologies_en}
 \caption{\bf Distribution of abundances of network topologies.}
 \label{fig:distopol}
\end{figure}

The most abundant network topologies are shown in Fig~\ref{fig:topol}. 
Interestingly, among the 15 most abundant network topologies, eleven of them 
presented the Incoherent type 3 Feed-Forward Loop (I3-FFL) network motif. In 
addition, we found that topology number 9 could produce the striped phenotype 
with only two nodes.

\begin{figure}[!h]
 \includegraphics[width=\textwidth]{figures/results/most-abundant-topologies}
 \caption{\bf Most frequent network topologies.}
 In this figure we show the most frequent network topologies, which contain the 
 I3-FFL network motif. Top left of each box is presented the network topology 
 number in order of abundance. Top center is presented the number of GRNs 
 belonging to that network topology. Circled in red are the input nodes of 
 I3-FFL in network topologies 1 to 5, that display the striped pattern of gene 
 expression using three nodes, in contrast, network topology number 9 displays 
 the same phenotype using a two-node GRN.
 \label{fig:topol}
\end{figure}

The neutral network shows that most of the network topologies are grouped into a
connected graph. This graph is composed of 639 nodes that are accessible by one 
mutational step. The 15 most abundant network topologies, with the exception of 
number 10, are located in this graph (Fig~\ref{fig:neutral-network}). We also 
found nodes inside the neutral network that had more connections between them 
than with other nodes, generating thus clusters of nodes, interestingly in 
almost every cluster is present one or more of the most abundant network 
topologies. In addition, we found that the first eight most abundant network 
topologies were located in the largest cluster.

\begin{figure}[!h]
 \includegraphics[width=\textwidth]{figures/results/grafo-hamming-15-nodes-inhibition-bar}
 \caption{\bf Neutral network.}
 This graph shows network topologies connected by one mutational step. Most 
 abundant network topologies are shown as black nodes. All nodes with the same 
 color belong to a cluster of nodes more connected between them than with other 
 nodes in the graph.
 \label{fig:neutral-network}
\end{figure}

The degree distribution of this network seems to follow a Poisson distribution
(Pearson’s $\chi^2 = 12.8902$, $\text{p-value} = 0.115684$) characteristic of a 
Erdös-Rényi network \cite{Erdos1959}. This kind of network presents a great 
number of nodes with low degree and a few nodes with high degree 
(Fig~\ref{fig:deg-dist}). Nodes corresponding to topologies number one and three
showed the highest degree ($deg(1) = deg (3) = 14$). Curiously, nodes with 
higher degree resulted to be also those network topologies with higher abundance
(Fig~\ref{fig:ab-deg}), indicating a high robustness to topological mutations as
well as robustness to mutations in the parameter values. This relationship 
between abundance and node degree was corroborated by the Spearman correlation 
($r_S = −0.353752$, $p = 6.2933\times10^{-23} $).

\begin{figure}[!h]
 \includegraphics[width=\textwidth]{figures/results/neutral-network-degree-distribution}
 \caption{\bf Histogram of vertex degree distribution in the neutral network.}
 \label{fig:deg-dist}
\end{figure}

\begin{figure}[!h]
 \includegraphics[width=\textwidth]{figures/results/abundance-vs-node-degree}
 \caption{\bf Scatter plot of Topology abundance vs. Node degree in the neutral
 network.}
 \label{fig:ab-deg}
\end{figure}

We wanted to know if connecting topologies separated by two mutational steps
would produce a connected graph, for that reason we generated another neutral
network joining neighbors at two mutational steps and we found that only 
topology number 420 remained unconnected from the rest of topologies in the 
graph (Fig~\ref{fig:2neut-net}).

\begin{figure}
 \includegraphics[width=\textwidth]{figures/results/hamming-2pasos}
 \caption{\bf Network of topologies separated by two mutational steps.}
 The network shows that all the topologies are connected by two or less 
 mutational steps, except for topology number 420 (big red circle). Big black 
 circles correspond to the 15 topologies with higher abundance.
 \label{fig:2neut-net}
\end{figure}

Besides topological robustness, we wanted to know if network topologies were
robust to perturbations in the values of the parameters. After performing the
perturbations to a GRN from each topology, we plotted the abundance of the
topology vs. its robustness. We did not find a significant correlation between
these two variables ($r_S = −0.0389724$, $\text{p-value} = 0.298475$) 
(Fig~\ref{fig:ab-rob}). However, when we grouped those topologies containing 
the same subgraph and calculated their mean robustness, we found that between 
the most robust subgraphs is the ``Bistable'' motif reported by Cotterell \& 
Sharpe \cite{Cotterell2010}, the I3-FFL and the ``Overlapping domains'' motif 
(which is a variation of the I3-FFL) also reported by Cotterell \& Sharpe (see 
Supplementary Material).

\begin{figure}[!h]
 \includegraphics[width=\textwidth]{figures/results/abundance-vs-robustness}
 \caption{\bf Scatter plot of Abundance vs. robustness.}
 The robustness of each topology was calculated as the number of perturbations
 in which the fitness was greater or equal to 0.95 with respect to the total of
 perturbations performed.
 \label{fig:ab-rob}
\end{figure}

The average vertex degree indicates that in general all the network topologies
display high robustness and that they could resist an average of approximately 4
topological mutations without losing their fitness. Additionally, the average
path length indicates that the evolutionary landscape is easily accessible for
most network topologies (Table~\ref{table1}).

\begin{table}[!ht]
% \begin{adjustwidth}{-2.25in}{0in} % Comment out/remove adjustwidth environment if table fits in text column.
 \centering
 \caption{{\bf Network descriptors of neutral network of topologies.}}
 \begin{tabular}{|l|l|}
 \hline
 {\bf Network descriptor} & {\bf Value}\\ \thickhline
 Average vertex degree  & 3.85714 \\ \hline
 Clustering coefficient \cite{Watts1998} & 0.117097 \\ \hline
 Average path length    & 7.25261       \\ \hline
 Graph connectedness    & 0.00540974    \\ \hline
 Graph diameter*        & 25            \\ \hline
 Average path length*   & 9.0499        \\ \hline
 \end{tabular}
 \begin{flushleft} *These descriptors were calculated for the main connected
 subgraph of the network.
 \end{flushleft}
 \label{table1}
% \end{adjustwidth}
 \end{table}

The average complexity index of network topologies was 39.2435 ($\pm 11.6768$)
and was found to be negatively correlated with the abundance ($r_S = -0.316653$,
$\text{p-value} = 2.25161\times10^{-18}$) (Fig~\ref{fig:ab-comp}).

\begin{figure}[!h]
 \includegraphics[width=\textwidth]{figures/results/abundance-vs-complexity-index}
 \caption{\bf Scatter plot of Abundance vs. Complexity index.}
 Each point in the plot represents one of the 714 network topologies.
 \label{fig:ab-comp}
\end{figure}

Spatiotemporal expression profiles showed that all GRNs reached a steady state
expression profile before $t = 250$ and most of them indeed reached the steady
state before $t = 150$ (Fig~\ref{fig:profile}).

\begin{figure}[!h]
 \includegraphics[width=\textwidth]{figures/results/GRN_660_topol_33_spatio-temp_exp_profile}
 \caption{\bf Most GRNs reach the steady state before 250 time steps.}
 GRN number 660 displaying the topology number 33 (A) and its spatiotemporal
 expression profile (B).
 \label{fig:profile}
\end{figure}

Hierarchical clustering of spatiotemporal expression profiles showed that GRNs
with topologies that seemed to be very different have indeed very similar
expression dynamics. We expected to find GRNs with the same topology to group
together but instead we found GRNs with different topologies grouped as sister
taxa (Fig~\ref{fig:exp-tree}). More interestingly, some GRNs displaying very
abundant network topologies are grouped together with GRNs displaying very
infrequent network topologies.

\begin{figure}[!h]
 \includegraphics[width=\textwidth]{figures/results/spatio-temporal-profile-clustering-i3-ffl}
 \caption{\bf Clustering of spatiotemporal expression profiles obtained by
 neighbor joining showing the presence of I3-FFL network motif.}
 \label{fig:exp-tree}
\end{figure}

Clustering of topologies revealed that there are six main groups of topologies
in our set of 2061 GRNs (Fig~\ref{fig:subg-tree}).

\begin{figure}[!h]
 \includegraphics[width=\textwidth]{figures/results/subgraph-profile-clustering-i3-ffl}
 \caption{\bf Clustering of network topologies.}
 Blue circles correspond to topologies that contain the I3-FFL network motif.
 \label{fig:subg-tree}
\end{figure}

Among the new network topologies we report here, perhaps the most interesting
is the topology number 9, which is able to express the striped gene expression
pattern with only two genes and indeed expresses this pattern in both genes
(Fig~\ref{fig:topol9}).

\begin{figure}[!h]
 \includegraphics[width=\textwidth]{figures/results/topol-9}
 \caption{\bf Expression dynamics of topology 9.}
 A: Network topology number 9. B: Spatiotemporal expression profile of a GRN
 with topology 9.
 \label{fig:topol9}
\end{figure}

Although the topology 9 consists of only two genes, its mechanism of expression
is dynamically more complex than that of other topologies such as the simple
I3-FFL. Initially the expression pattern of the gene A is uniform in the
morphogenetic field, but later it is shaped by the morphogen gradient in such a
way that two important events occur, in first place the overall expression
level of the gene A is decreased by the inhibition of the morphogen, releasing
gene C from its repression by gene A; in second place, appears a non-homogeneous
spatial distribution in which gene A is expressed at higher levels in the cells
at the rightmost part of the morphogenetic field.\\

These two events conduce to the repression of the gene C by the gene A in the
rightmost part of the morphogenetic field and by the morphogen in the leftmost
part of the morphogenetic field, thus generating a lightly higher level of
expression of gene C in the center of the field. Because gene C activates its
own expression, a band of expression begins to form. This band of expression
then begins to activate the gene A also in the center of the field
(Fig~\ref{fig:topol9-dyn}).

\begin{figure}[!h]
 \includegraphics[width=\textwidth]{figures/results/topol-9-dynamics}
 \caption{\bf Dynamics of expression in topology number 9.}
 \label{fig:topol9-dyn}
\end{figure}

In this way two bands of expression are formed that depend on each other. The
high level of gene C in the middle of the field counteract its inhibition by
gene A, while in the ends of the bands the inhibition of gene A on gene C
overcomes the self-activation of gene C, preventing the expression of the
latter to extend out of the middle of the field. In turn, the band of
expression of gene A can only exist in the middle of the field because its
expression depends on gene C.

This is a complex mechanism described by the first time in this study, and it
would be interesting to study its feasibility and possible implementation in
synthetic gene regulatory networks.

%%%%%%%%%%%%%%%%%%%%%%%%%% copy of table environment %%%%%%%%%%%%%%%%%%%%%%%%
% \begin{table}[!ht]
% \begin{adjustwidth}{-2.25in}{0in} % Comment out/remove adjustwidth environment if table fits in text column.
% \centering
% \caption{{\bf Network descriptors of neutral network of topologies.}}
% \begin{tabular}{|l+l|l|l|l|l|l|l|}
% \hline
% \multicolumn{4}{|l|}{\bf Heading1} & \multicolumn{4}{|l|}{\bf Heading2}\\ \thickhline
% $cell1 row1$ & cell2 row 1 & cell3 row 1 & cell4 row 1 & cell5 row 1 & cell6 row 1 & cell7 row 1 & cell8 row 1\\ \hline
% $cell1 row2$ & cell2 row 2 & cell3 row 2 & cell4 row 2 & cell5 row 2 & cell6 row 2 & cell7 row 2 & cell8 row 2\\ \hline
% $cell1 row3$ & cell2 row 3 & cell3 row 3 & cell4 row 3 & cell5 row 3 & cell6 row 3 & cell7 row 3 & cell8 row 3\\ \hline
% \end{tabular}
% \begin{flushleft} Table notes Phasellus venenatis, tortor nec vestibulum mattis, massa tortor interdum felis, nec pellentesque metus tortor nec nisl. Ut ornare mauris tellus, vel dapibus arcu suscipit sed.
% \end{flushleft}
% \label{table1}
% \end{adjustwidth}
% \end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%PLOS does not support heading levels beyond the 3rd (no 4th level headings).
% \subsection*{\lorem\ and \ipsum\ nunc blandit a tortor}
% \subsubsection*{3rd level heading} 
% Sed suscipit iaculis mi at feugiat. 
% 
% \begin{enumerate}
% 	\item{react}
% 	\item{diffuse free particles}
% 	\item{increment time by dt and go to 1}
% \end{enumerate}
% 
% 
% \subsection*{Sed ac quam id nisi malesuada congue}
% 
% Praesent faucibus semper libero.
% 
% \begin{itemize}
% 	\item First bulleted item.
% 	\item Second bulleted item.
% 	\item Third bulleted item.
% \end{itemize}

\section*{Discussion}

We found the presence of network motifs in the GRNs produced by the genetic
algorithm. However, we also found a great diversity of network topologies that
could produce a striped gene expression pattern.\\

We found that the I3-FFL is one of the network motifs that appears more
frequently in these GRNs, as has also been reported by Cotterell \& Sharpe
\cite{Cotterell2010}, and has been found in transcriptional networks involved 
in cell fate definition \cite{Li2019}. Moreover, our results indicate that this 
is a robust motif that can be usefull for construction of synthetic gene 
regulatory networks that are exposed to stochastic fluctuations in the 
environment and mutations.

However, as our model was less stringent, we found a larger number of possible 
topologies with the same striped phenotype.\\

Although it has been experimentally demonstrated that the Incoherent Feed 
Forward Loop type 2 (I2-FFL) is able to generate a striped gene expression 
pattern, \cite{Schaerli2014, Basu2005}, in our study this motif was present in 
only 10.92\% of the topologies. This fact may be because genetic algorithms 
can not always find global minima but local minima or its surroundings, however,
it is also possible that the I2-FFL motif is not as robust as other motifs and 
therefore be less represented.

We also found topologies with a similar topology to that of the “opposing
gradients” reported by Schaerli et al. \cite{Schaerli2018,Schaerli2014}.
Although the “opposing gradients” mechanism requires the constitutive expression
of two of the genes in the GRN, and our model does not take constitutive
expression into account, we realized that the GRNs that presented the “opposing
gradients” topology, had also commonly positive auto-regulation interactions on
these two nodes, supplying in this way the lack of constitutive expression.

The correlation between the complexity index of a topology and its abundance 
indicates that networks with simpler designs are more robust to changes in 
interaction parameters and that more complex networks are less robusts. For 
example, topology number 420, which presents high complexity, was one of the 
topologies with lower abundance and was disconnected from all other topologies 
in the two-step neighbors in the neutral network.

It is interesting that the one-step neutral network

\section*{Conclusion}

This study can help us to understand the mechanisms by which the GRNs work in
nature as modules that make part of genomes, but it also can provide guidance
in the construction of new synthetic GRNs in the field of synthetic biology.

%CO\textsubscript{2} Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque. Quisque augue sem, tincidunt sit amet feugiat eget, ullamcorper sed velit.

%Aliquam tincidunt, ex in facilisis elementum, libero lectus luctus est, non vulputate nisl augue at dolor. For more information, see \nameref{S1_Appendix}.

\section*{Supporting information}

% Include only the SI item label in the paragraph heading. Use the \nameref{label} command to cite SI items in the text.
\paragraph*{S1 Fig.}
\label{S1_Fig}
{\bf Bold the title sentence.} Add descriptive text after the title of the item (optional).

\paragraph*{S2 Fig.}
\label{S2_Fig}
{\bf Lorem ipsum.} Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque.

\paragraph*{S1 File.}
\label{S1_File}
{\bf Lorem ipsum.}  Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque.

\paragraph*{S1 Video.}
\label{S1_Video}
{\bf Lorem ipsum.}  Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque.

\paragraph*{S1 Appendix.}
\label{S1_Appendix}
{\bf Lorem ipsum.} Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque.

\paragraph*{S1 Table.}
\label{S1_Table}
{\bf Lorem ipsum.} Maecenas convallis mauris sit amet sem ultrices gravida. Etiam eget sapien nibh. Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula. Curabitur fringilla pulvinar lectus consectetur pellentesque.

\section*{Acknowledgments}
Cras egestas velit mauris, eu mollis turpis pellentesque sit amet. Interdum et malesuada fames ac ante ipsum primis in faucibus. Nam id pretium nisi. Sed ac quam id nisi malesuada congue. Sed interdum aliquet augue, at pellentesque quam rhoncus vitae.

\nolinenumbers

% Either type in your references using
% \begin{thebibliography}{}
% \bibitem{}
% Text
% \end{thebibliography}
%
% or
%
% Compile your BiBTeX database using our plos2015.bst
% style file and paste the contents of your .bbl file
% here. See http://journals.plos.org/plosone/s/latex for 
% step-by-step instructions.
% 
% \begin{thebibliography}{10}
% 
% \bibitem{bib1}
% Conant GC, Wolfe KH.
% \newblock {{T}urning a hobby into a job: how duplicated genes find new
%   functions}.
% \newblock Nat Rev Genet. 2008 Dec;9(12):938--950.
% 
% \bibitem{bib2}
% Ohno S.
% \newblock Evolution by gene duplication.
% \newblock London: George Alien \& Unwin Ltd. Berlin, Heidelberg and New York:
%   Springer-Verlag.; 1970.
% 
% \bibitem{bib3}
% Magwire MM, Bayer F, Webster CL, Cao C, Jiggins FM.
% \newblock {{S}uccessive increases in the resistance of {D}rosophila to viral
%   infection through a transposon insertion followed by a {D}uplication}.
% \newblock PLoS Genet. 2011 Oct;7(10):e1002337.
% 
% \end{thebibliography}

\bibliography{referencias-tesis}{}

\end{document}

### 2020-06-04 manuscript/PLOS-MSc-Thesis.tex
general changes i don't remember

@@
 then begins to activate the gene A also in the center of the field
-(Fig~\ref{fig:topol9-dyn}).
-
-\begin{figure}[!h]
- \includegraphics[width=\textwidth]{figures/results/topol-9-dynamics}
- \caption{\bf Dynamics of expression in topology number 9.}
- \label{fig:topol9-dyn}
-\end{figure}
+(\nameref{S2_Fig}).
+% (Fig~\ref{fig:topol9-dyn}).
+
+% \begin{figure}[!h]
+%  \includegraphics[width=\textwidth]{figures/results/topol-9-dynamics}
+%  \caption{\bf Dynamics of expression in topology number 9.}
+%  \label{fig:topol9-dyn}
+% \end{figure}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%% copy of table environment %%%%%%%%%%%%%%%%%%%%%%%%
-% \begin{table}[!ht]
-% \begin{adjustwidth}{-2.25in}{0in} % Comment out/remove adjustwidth environment if table fits in text column.
-% \centering
-% \caption{{\bf Network descriptors of neutral network of topologies.}}
-% \begin{tabular}{|l+l|l|l|l|l|l|l|}
-% \hline
-% \multicolumn{4}{|l|}{\bf Heading1} & \multicolumn{4}{|l|}{\bf Heading2}\\ \thickhline
-% $cell1 row1$ & cell2 row 1 & cell3 row 1 & cell4 row 1 & cell5 row 1 & cell6 row 1 & cell7 row 1 & cell8 row 1\\ \hline
-% $cell1 row2$ & cell2 row 2 & cell3 row 2 & cell4 row 2 & cell5 row 2 & cell6 row 2 & cell7 row 2 & cell8 row 2\\ \hline
-% $cell1 row3$ & cell2 row 3 & cell3 row 3 & cell4 row 3 & cell5 row 3 & cell6 row 3 & cell7 row 3 & cell8 row 3\\ \hline
-% \end{tabular}
-% \begin{flushleft} Table notes Phasellus venenatis, tortor nec vestibulum mattis, massa tortor interdum felis, nec pellentesque metus tortor nec nisl.
-Ut ornare mauris tellus, vel dapibus arcu suscipit sed.
-% \end{flushleft}
-% \label{table1}
-% \end{adjustwidth}
-% \end{table}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 %PLOS does not support heading levels beyond the 3rd (no 4th level headings).
@@
 
-It is interesting that the one-step neutral network
+The cladogram of spatiotemporal expression profiles shows that the genetic 
+algorithm selected similar expression mechanisms that can be produced by very 
+different topologies.
+
+Although GRNs that produce a band of gene expression with only two nodes have 
+been described (e.g. the I-zero motif, \cite{Schaerli2014}) this is to our 
+knowledge the first time a GRN with a topology such as that of topology nine 
+and its expression mechanism are reported, increasing our knowledge of the 
+genotype-phenotype map in this kind of system.
+
+However, it is still necessary to confirm experimentally the ability of this 
+topology to generate the described pattern and it would be interesting to 
+study its feasibility and possible implementation in synthetic gene regulatory 
+networks.
 
@@
 This study can help us to understand the mechanisms by which the GRNs work in
-nature as modules that make part of genomes, but it also can provide guidance
+nature as modules that make part of genomes, and it also can provide guidance
 in the construction of new synthetic GRNs in the field of synthetic biology.
@@
 \label{S1_Fig}
-{\bf Bold the title sentence.} Add descriptive text after the title of the item (optional).
+{\bf Relevant subgraphs present in the set of GRNs.}
+%Add descriptive text after the title of the item (optional).
 
@@
 \label{S2_Fig}
-{\bf Lorem ipsum.} Maecenas convallis mauris sit amet sem ultrices gravida.
+{\bf Dynamics of expression of topology number 9.}
+
+% \paragraph*{S1 File.}
+% \label{S1_File}
+% {\bf Lorem ipsum.}  Maecenas convallis mauris sit amet sem ultrices gravida.
 Etiam eget sapien nibh.
@@
 Curabitur fringilla pulvinar lectus consectetur pellentesque.
-
-\paragraph*{S1 File.}
-\label{S1_File}
-{\bf Lorem ipsum.}  Maecenas convallis mauris sit amet sem ultrices gravida.
+% 
+% \paragraph*{S1 Video.}
+% \label{S1_Video}
+% {\bf Lorem ipsum.}  Maecenas convallis mauris sit amet sem ultrices gravida.
 Etiam eget sapien nibh.
@@
 Curabitur fringilla pulvinar lectus consectetur pellentesque.
-
-\paragraph*{S1 Video.}
-\label{S1_Video}
-{\bf Lorem ipsum.}  Maecenas convallis mauris sit amet sem ultrices gravida.
+% 
+% \paragraph*{S1 Appendix.}
+% \label{S1_Appendix}
+% {\bf Lorem ipsum.} Maecenas convallis mauris sit amet sem ultrices gravida.
 Etiam eget sapien nibh.
@@
 
-\paragraph*{S1 Appendix.}
-\label{S1_Appendix}
-{\bf Lorem ipsum.} Maecenas convallis mauris sit amet sem ultrices gravida.
-Etiam eget sapien nibh.
-Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula.
-Curabitur fringilla pulvinar lectus consectetur pellentesque.
-
 \paragraph*{S1 Table.}
 \label{S1_Table}
-{\bf Lorem ipsum.} Maecenas convallis mauris sit amet sem ultrices gravida.
-Etiam eget sapien nibh.
-Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula.
-Curabitur fringilla pulvinar lectus consectetur pellentesque.
+{\bf Proportion of relevant subgraphs in the set of GRNs.}
 

### 2020-06-04 manuscript/PLOS-MSc-Thesis.tex
write abstract

@@
 \section*{Abstract}
-Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur eget porta 
-erat. Morbi consectetur est vel gravida pretium. Suspendisse ut dui eu ante 
-cursus gravida non sed sem. Nullam sapien tellus, commodo id velit id, eleifend 
-volutpat quam. Phasellus mauris velit, dapibus finibus elementum vel, pulvinar 
-non tellus. Nunc pellentesque pretium diam, quis maximus dolor faucibus id. Nunc
-convallis sodales ante, ut ullamcorper est egestas vitae. Nam sit amet enim 
-ultrices, ultrices elit pulvinar, volutpat risus.
+One of the central problems in developmental and synthetic biology is to 
+understand how do cells obtain information about its relative position to 
+other cells in a developing embryo, a tissue or in a petri dish. It has been 
+stated that a gradient of concentration of certain molecule is enough to 
+provide such a positional information and produce cell differentiation by 
+interacting with the cell's gene regulatory networks (GRNs). In this study we 
+used a genetic algorithm in order to explore the genotypic space of GRNs 
+capable of producing a band expression pattern in a group of cells.
+
+Out of 2061 GRNs selected using the genetic algorithm, 714 groups of topologies
+were found that were able to interpret a morphogen input. The main network 
+motif involved in the formation of the band expression pattern was the type 3
+Incoherent Feed Forward Loop (I3-FFL); we also present here other network
+topologies that have not been previously reported.
 

### 2020-06-14 manuscript/PLOS-MSc-Thesis.tex
fix Jay's affiliation and reformat the Abstract

--- 2020-06-04 manuscript/PLOS-MSc-Thesis.tex
@@
 \bigskip
-\textbf{1} Instituto de Biología, Universidad de Antioquia, Medellín, Antioquia, Colombia
+\textbf{1} Instituto de Biología, Universidad de Antioquia, Medellín,
+Antioquia, Colombia
 \\
-\textbf{2} Instituto de Física, Universidad de Antioquia, Medellín, Antioquia, Colombia
+\textbf{2} Instituto de Física, Universidad de Antioquia, Medellín, Antioquia,
+Colombia
 \\
-\textbf{3} Affiliation Dept/Program/Center, Ghent University, Ghent, State, Belgium
+\textbf{3} Flanders Marine Institute (VLIZ), Wandelaarkaai 7, 8400 Oostende,
+Belgium
 % \textbf{3} Affiliation Dept/Program/Center, Institution Name, City, State, Country
@@
 \section*{Abstract}
-Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur eget porta 
-erat. Morbi consectetur est vel gravida pretium. Suspendisse ut dui eu ante 
-cursus gravida non sed sem. Nullam sapien tellus, commodo id velit id, eleifend 
-volutpat quam. Phasellus mauris velit, dapibus finibus elementum vel, pulvinar 
-non tellus. Nunc pellentesque pretium diam, quis maximus dolor faucibus id. Nunc
-convallis sodales ante, ut ullamcorper est egestas vitae. Nam sit amet enim 
-ultrices, ultrices elit pulvinar, volutpat risus.
+
+\subsection*{Background}
+One of the central problems in developmental and synthetic biology is to 
+understand how cells process external cues on its relative position to other 
+cells in a developing embryo, a tissue or in a petri dish and transform such
+information into a coherent response, e.g. terminal differentiation state. It 
+has been stated that a gradient of concentration of certain molecule is enough 
+to provide such positional information and produce cell differentiation by 
+interacting with the cell's gene regulatory networks (GRNs).
+
+\subsection*{Methods}
+In this study we used a genetic algorithm in order to explore the genotypic 
+space of GRNs capable of producing a band expression pattern in a group of 
+cells.
+
+\subsection*{Results}
+Out of 2061 GRNs selected using the genetic algorithm, 714 groups of 
+topologies were found that were able to interpret a morphogen input. The 
+main network motif involved in the formation of the band expression pattern 
+was the type 3 Incoherent Feed Forward Loop (I3-FFL); we also present here 
+other network topologies that have not been previously reported.
+
+\subsection*{Conclusions}
 
@@
 have been validated by experimental studies
-\cite{Kalir2005,Kalir2004,Mangan2003a,ODonnell2005,Basu2005}.
+\cite{Kalir2005,Kalir2004,Mangan2003a,ODonnell2005}, and some of them 
+have used synthetic GRNs to produce artificial cell differentiation in bacteria
+cell cultures~\cite{Basu2005}.
 
@@
 
-The classification of GRNs showed that in our initial set of 2061 GRNs exists a 
-number of 714 different network topologies, each one of them containing a 
-different number of GRNs (Fig~\ref{fig:distopol}). This abundances distribution 
-was not produced by chance (Shannon entropy = 8.52963, 95\% confidence interval 
-= [10.8915, 10.9275]), and instead it shows a structuring, indicating the 
-possible existence of network motifs.
+The classification of GRNs showed that our initial set of 2061 GRNs can be 
+grouped in a number of 714 distinct network topologies, each one of them
+containing a different number of GRNs (Fig~\ref{fig:distopol}). This abundances
+distribution was not produced by chance (Shannon entropy = 8.52963, 95\%
+confidence interval = [10.8915, 10.9275]), and instead it displays some level 
+of structure, indicating the existence of network motifs.
 
@@
 \begin{figure}[!h]
- \includegraphics[width=\textwidth]{figures/results/most-abundant-topologies}
+ \includegraphics[width=\textwidth]{figures/results/most-abu-topol}
  \caption{\bf Most frequent network topologies.}
- In this figure we show the most frequent network topologies, which contain the 
- I3-FFL network motif. Top left of each box is presented the network topology 
- number in order of abundance. Top center is presented the number of GRNs 
- belonging to that network topology. Circled in red are the input nodes of 
- I3-FFL in network topologies 1 to 5, that display the striped pattern of gene 
- expression using three nodes, in contrast, network topology number 9 displays 
- the same phenotype using a two-node GRN.
+ Most frequent network topologies contain the I3-FFL network motif. Topologies 
+ are numbered in order of decreasing abundance.
+ The number of GRNs belonging to each network topology is presented below 
+ topology number.
+ Network topology number 9 produces the striped phenotype using a two-node GRN.
  \label{fig:topol}
@@
 
-The neutral network shows that most of the network topologies are grouped into a
-connected graph. This graph is composed of 639 nodes that are accessible by one 
-mutational step. The 15 most abundant network topologies, with the exception of 
-number 10, are located in this graph (Fig~\ref{fig:neutral-network}). We also 
+The neutral network (a metagraph in wich each node is a network topology) 
+shows that most of the network topologies are grouped into a connected graph.
+This graph is composed of 639 nodes that are accessible by one mutational step.
+The 15 most abundant network topologies, with the exception of 
+number 10, are part of this graph (Fig~\ref{fig:neutral-network}). We also 
 found nodes inside the neutral network that had more connections between them 
-than with other nodes, generating thus clusters of nodes, interestingly in 
-almost every cluster is present one or more of the most abundant network 
+than with other nodes, generating clusters of nodes, interestingly in 
+almost every cluster is present one or more of the 15 most abundant network 
 topologies. In addition, we found that the first eight most abundant network 
@@
 higher degree resulted to be also those network topologies with higher abundance
-(Fig~\ref{fig:ab-deg}), indicating a high robustness to topological mutations as
-well as robustness to mutations in the parameter values. This relationship 
-between abundance and node degree was corroborated by the Spearman correlation 
+(Fig~\ref{fig:ab-deg}). This relationship between abundance and node degree 
+was corroborated by the Spearman's rank correlation coefficient 
 ($r_S = −0.353752$, $p = 6.2933\times10^{-23} $).
@@
 expression dynamics. We expected to find GRNs with the same topology to group
-together but instead we found GRNs with different topologies grouped as sister
-taxa (Fig~\ref{fig:exp-tree}). More interestingly, some GRNs displaying very
-abundant network topologies are grouped together with GRNs displaying very
-infrequent network topologies.
+together but instead we found GRNs with different topologies grouped as close
+neighbors (Fig~\ref{fig:exp-tree}). More interestingly, some GRNs displaying 
+very abundant network topologies were grouped together with GRNs displaying 
+very infrequent network topologies.
 
@@
 level of the gene A is decreased by the inhibition of the morphogen, releasing
-gene C from its repression by gene A; in second place, appears a non-homogeneous
-spatial distribution in which gene A is expressed at higher levels in the cells
-at the rightmost part of the morphogenetic field.\\
+gene C from its repression by gene A; in second place, it appears a 
+non-homogeneous spatial distribution in which gene A is expressed at higher 
+levels in the cells at the rightmost part of the morphogenetic field.\\
 
@@
 rightmost part of the morphogenetic field and by the morphogen in the leftmost
-part of the morphogenetic field, thus generating a lightly higher level of
+part of the morphogenetic field, thus generating a slightly higher level of
 expression of gene C in the center of the field. Because gene C activates its
@@
 
-In this way two bands of expression are formed that depend on each other. The
+In this way two bands of expression that depend on each other are formed. The
 high level of gene C in the middle of the field counteract its inhibition by
-gene A, while in the ends of the bands the inhibition of gene A on gene C
+gene A, while at the ends of the bands the inhibition of gene A on gene C
 overcomes the self-activation of gene C, preventing the expression of the
@@
 
-We found the presence of network motifs in the GRNs produced by the genetic
-algorithm. However, we also found a great diversity of network topologies that
-could produce a striped gene expression pattern.\\
+We found the presence of network motifs in the set of GRNs produced by the
+genetic algorithm. However, we also found a great diversity of network 
+topologies that could produce a striped gene expression pattern.\\
+
+That the most abundant topologies were found to be part of highly connected 
+clusters in the neutral network indicates that these topologies are both robust 
+to quantitative changes in the strength of interactions and diffusion and 
+degradation parameters as well as to qualitative changes in the topology 
+itself.
 

### 2020-06-14 manuscript/PLOS-MSc-Thesis.tex
change title

@@
 
-\newcommand{\lorem}{{\bf LOREM}}
-\newcommand{\ipsum}{{\bf IPSUM}}
+%\newcommand{\lorem}{{\bf LOREM}}
+%\newcommand{\ipsum}{{\bf IPSUM}}
 
@@
 {\Large
-\textbf\newline{Gene regulatory networks producing a striped band of gene 
-expression} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
+\textbf\newline{Elucidating multi-input processing 3-node gene regulatory 
+network topologies capable of generating striped gene expression patterns}
+% Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
 }

### 2020-08-25 manuscript/PLOS-MSc-Thesis.tex
change first paragraph of abstract according to J. Gutiérrez recomendations

@@
 \subsection*{Background}
-One of the central problems in developmental and synthetic biology is to 
-understand how cells process external cues on its relative position to other 
-cells in a developing embryo, a tissue or in a petri dish and transform such
-information into a coherent response, e.g. terminal differentiation state. It 
-has been stated that a gradient of concentration of certain molecule is enough 
-to provide such positional information and produce cell differentiation by 
-interacting with the cell's gene regulatory networks (GRNs).
+A central problem in developmental and synthetic biology is to understand the 
+mechanisms by which cells in a tissue or in a petri dish process external cues 
+and transform such information into a coherent response, e.g. a terminal 
+differentiation state. It has been stated that a gradient of concentration of 
+certain molecule is enough to provide such positional information and produce 
+cell differentiation by interacting with the cell's gene regulatory networks 
+(GRNs).
 

### 2020-08-25 manuscript/PLOS-MSc-Thesis.tex
change abstract introduction to morphogen and GRNs

@@
 and transform such information into a coherent response, e.g. a terminal 
-differentiation state. It has been stated that a gradient of concentration of 
-certain molecule is enough to provide such positional information and produce 
-cell differentiation by interacting with the cell's gene regulatory networks 
-(GRNs).
+differentiation state. It was long believed that this type of positional 
+information could be entirely attributed to a gradient of concentration of a 
+certain signaling molecule (i.e. a morphogen). However, advances in experimental 
+methodologies and computer modeling have demonstrated the crucial role of the 
+dynamics of a cell's gene regulatory network (GRN) in decoding the information 
+carried by the morphogen, which is eventually translated into a spatial 
+pattern. This morphogen interpretation mechanism has gained a lot of attention 
+in systems biology as a tractable system to investigate the emergent properties 
+of complex genotype-phenotype maps.
 

### 2020-08-25 manuscript/PLOS-MSc-Thesis.tex
extend abstract methods section

@@
 In this study we used a genetic algorithm in order to explore the genotypic 
-space of GRNs capable of producing a band expression pattern in a group of 
-cells.
+space of three-node GRNs capable of producing a band expression pattern in a 
+group of cells. In contrast to previous studies, we explore multiple input 
+network designs that enable more morphogen interpretation mechanisms to arise.
 

### 2020-08-25 manuscript/PLOS-MSc-Thesis.tex
extend abstract results section

@@
 was the type 3 Incoherent Feed Forward Loop (I3-FFL); we also present here 
-other network topologies that have not been previously reported.
+other network topologies that have not been previously reported and that can 
+be useful as synthetic regulatory networks due to its relatively simple design 
+and robust morphogen interpretation mechanisms.
 

### 2020-08-26 modules/DesignMorphogeneResponsiveGRCs.m
wrap long comment line

@@ -151,8 +151,8 @@ TimeSeriesSpatialExpOutput=StripeFormingGRCs4SSMorpGradientSFGRM[SSExpValues,Mor
 
 (************************************************************************************************************************)
 
-(*NOTE: to assess the degenerate structure of a GRC we have to slightly modify the previous methods used to assess a GRC for its ability to
-	generate a stripe expression pattern*)
+(*NOTE: to assess the degenerate structure of a GRC we have to slightly modify the previous methods used to assess a GRC
+ for its ability to	generate a stripe expression pattern*)
 AssessDegenerateStruc4EngineeredMorpResponsiveGRC[GRCGenotype_List,
   FitnessThrs_, MaxNumDeletedEdges_, ReportData_String] :=
  Block[{ParamGRCWiring, MorphogeneInput, DiffParams, DegParams,

### 2020-11-09 docs/manuscript/PLOS-MSc-Thesis.tex
update Genetic algorithm section to describe Markov chain algorithm

--- 2020-08-25 manuscript/PLOS-MSc-Thesis.tex
@@
 \cite{Kalir2005,Kalir2004,Mangan2003a,ODonnell2005}, and some of them 
-have used synthetic GRNs to produce artificial cell differentiation in bacteria
-cell cultures~\cite{Basu2005}.
+have even used synthetic GRNs to produce artificial cell differentiation~\cite{Basu2005} 
+and circadian gene expression~\cite{atkinson_development_2003}.
 
@@
 \begin{enumerate}
- \item{\bf Generate a set of 200 random GRNs.}
-
- \item{\bf Numerically solve the dynamical system for each GRN from t=0 until 
+ \item{\bf Generate a random GRN.}
+
+ \item{\bf{Numerically solve the dynamical system for the GRN from t=0 until 
 t=500.} We integrated the differential equation using the function NDSolve from 
 Wolfram Mathematica 11 with the option “EquationSimplification” and the 
- “Residual” simplification method.
-
- \item{\bf Evaluate fitness by comparing the phenotypes with the optimal 
+“Residual” simplification method.}
+
+ \item{\bf Evaluate fitness by comparing the GRN phenotype with the optimal 
  phenotype (Fig~\ref{fig:exp-pattern}).} In order to compute the fitness 
@@
 
- \item{\bf Select GRNs with a probability proportional to its fitness.}
-
- \item{\bf Recombine and mutate GRNs.} Mutations included changes in the
- numerical values of degradation and diffusion parameters and interaction
- strengths, and addition and removal of interactions among genes and between the
- genes and the morphogen.
+ \item{\bf Mutate GRN network parameters or gene product diffusion or 
+degradation rates randomly and evaluate fitness again.} If fitness of the 
+mutated GRN is greater than that of the previous GRN, save GRN and continue 
+mutating.
+
+ \item {\bf Mutate GRNs randomly and evaluate fitness again.} Mutations 
+included changes in the  numerical values of degradation and diffusion 
+parameters and interaction strengths, and addition and removal of interactions 
+among genes and between the genes and the morphogen. If fitness of the mutated 
+GRN was greater or equal than that of the previous GRN, mutated GRN was saved 
+and the cycle continued from step 2. This cycle was repeated 5000000 times and 
+GRNs with a fitness greater or equal than 0.95 were saved into a file.
+
 \end{enumerate}
 
-Steps 2 to 5 were iterated 1000 times, resulting in a set of 2061 GRNs with a
+Steps 1 to 5 were iterated 500 times, resulting in a set of 2061 GRNs with a 
 striped pattern of gene expression in at least one gene.

### 2020-11-09 docs/manuscript/PLOS-MSc-Thesis.tex
write clearer Results section in Abstract

@@
 \subsection*{Results}
-Out of 2061 GRNs selected using the genetic algorithm, 714 groups of 
-topologies were found that were able to interpret a morphogen input. The 
-main network motif involved in the formation of the band expression pattern 
-was the type 3 Incoherent Feed Forward Loop (I3-FFL); we also present here 
-other network topologies that have not been previously reported and that can 
-be useful as synthetic regulatory networks due to its relatively simple design 
-and robust morphogen interpretation mechanisms.
+Out of 2061 GRNs selected using the genetic algorithm, we found 714 groups of 
+topologies that were able to interpret a morphogen input. The main network motif 
+involved in the formation of the band expression pattern was the type 3 
+Incoherent Feed Forward Loop (I3-FFL); we also present here other network 
+topologies that have not been previously reported and that can be useful as 
+synthetic regulatory networks due to its simple design and robust morphogen 
+interpretation mechanisms.
 

### 2020-11-13 docs/manuscript/PLOS-MSc-Thesis.tex
citate santos-moreno&schaerli 2019
add author summary and reformat affiliations

@@
 \bigskip
-\textbf{1} Instituto de Biología, Universidad de Antioquia, Medellín,
-Antioquia, Colombia
+\textbf{1} Grupo de Fundamentos y Enseñanza de la Física y los Sistemas Dinámicos,
+Instituto de Biología, Facultad de Ciencias Exactas y Naturales, Universidad de Antioquia
+UdeA, Calle 70 No. 52-21, Medellín, Colombia.
 \\
-\textbf{2} Instituto de Física, Universidad de Antioquia, Medellín, Antioquia,
-Colombia
+\textbf{2} Grupo de Fundamentos y Enseñanza de la Física y los Sistemas Dinámicos,
+Instituto de Física, Facultad de Ciencias Exactas y Naturales, Universidad de Antioquia
+UdeA, Calle 70 No. 52-21, Medellín, Colombia.
 \\
@@
 \section*{Author summary}
-Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur eget porta 
-erat. Morbi consectetur est vel gravida pretium. Suspendisse ut dui eu ante 
-cursus gravida non sed sem. Nullam sapien tellus, commodo id velit id, eleifend 
-volutpat quam. Phasellus mauris velit, dapibus finibus elementum vel, pulvinar 
-non tellus. Nunc pellentesque pretium diam, quis maximus dolor faucibus id. Nunc
-convallis sodales ante, ut ullamcorper est egestas vitae. Nam sit amet enim 
-ultrices, ultrices elit pulvinar, volutpat risus.
+Systems biology is growing very quickly thanks to the advances in computing and
+mathematical modeling of biological systems, and has proven to be a powerful tool that
+can help us to understand complex biological processes such as pattern formation and
+development. In this article we use an \textit{in silico} approach in order to find and study
+the design space of three node Gene Regulatory Networks that are able to produce a 
+striped pattern of gene expression.
 
@@
 have even used synthetic GRNs to produce artificial cell differentiation~\cite{Basu2005} 
-and circadian gene expression~\cite{atkinson_development_2003}.
+and circadian gene expression~\cite{atkinson_development_2003}. Moreover, 
+the design and implementation of GRNs through synthetic biology is emerging as 
+a promising tool to study biological phenomena as pattern formation 
+\cite{santos-moreno_using_2019}.
 

### 2020-11-13 docs/manuscript/PLOS-MSc-Thesis.tex
correct typos \noindent and equations

@@
 A central problem in developmental and synthetic biology is to understand the 
-mechanisms by which cells in a tissue or in a petri dish process external cues 
+mechanisms by which cells in a tissue or in a Petri dish process external cues 
 and transform such information into a coherent response, e.g. a terminal 
@@
 \section*{Author summary}
-Systems biology is growing very quickly thanks to the advances in computing and
+Systems biology is a very fast growing field thanks to the advances in computing and
 mathematical modeling of biological systems, and has proven to be a powerful tool that
-can help us to understand complex biological processes such as pattern formation and
+can help us to understand complex biological processes such as pattern formation in
 development. In this article we use an \textit{in silico} approach in order to find and study
@@
  \end{eqnarray}
-
+\noindent
 in which $[G]^{i}_{n}$ is the concentration of the i-th gene in the \emph{n}-th 
@@
 The input function is a sigmoid function described by Eq~\ref{ec:input-func}.
-
 \begin{equation}
@@
 \end{equation}
-
+\noindent
 in which $a$ is the sigmoid steepness, equal to 5; $a/b$ is the threshold value,
 set to 1 for all simulations, and $u^i$ is the following equation:
-
 \begin{equation}
@@
 Eq~\ref{ec:morph}.
-
 \begin{equation}
@@
 \end{equation}
-
+\noindent
 were $A_0$ is the concentration of the morphogen in the position zero of the 
@@
 
- \item{\bf{Numerically solve the dynamical system for the GRN from t=0 until 
+ \item{\bf Numerically solve the dynamical system for the GRN from t=0 until 
  t=500.} We integrated the differential equation using the function NDSolve from 
  Wolfram Mathematica 11 with the option “EquationSimplification” and the 
-“Residual” simplification method.}
+ “Residual” simplification method.
 
@@
  quasi-steady state:
-
  \begin{equation}
@@
  The second filter measures the spatial heterogeneity in the field:
-
  \begin{equation}
@@
  ($D_{max}$):
-
  \begin{equation}
@@
  The filters mentioned above are integrated in the following functions that 
- asses the quality of a given phenotype:
+ evaluates the  quality of a given phenotype:
 
@@
  \end{equation}
-
  \begin{equation}
@@
  the following fitness function:
-
  \begin{equation}
@@
 topologies was calculated using the following equation:
-
 \begin{equation}
@@
 of the network topologies and we modified each interaction independently by 
-increasing and decreasing its value in a 20\%. For topologies with 3 or more 
+increasing and decreasing its value by 20\%. For topologies with 3 or more 
 GRNs we chose the most similar to the mean configuration for that topology.\\
@@
 for normality.
-
 \begin{equation}
@@
 index using the following equation:
-
 \begin{equation}
@@
 \end{equation}
-
+\noindent
 being $a_i$ the degree of \emph{i}-th node and $V$ the number of nodes in the 
@@
 in cell fate definition \cite{Li2019}. Moreover, our results indicate that this 
-is a robust motif that can be usefull for construction of synthetic gene 
+is a robust motif that can be useful for construction of synthetic gene 
 regulatory networks that are exposed to stochastic fluctuations in the 

### 2020-12-23 docs/manuscript/PLOS-MSc-Thesis.tex
Remove trailing whitespaces and change abundance vs complexity fig to log
Delete paragraph about GRNs reaching quickly the steady state

@@
 \begin{figure}[!h]
- \includegraphics[width=\textwidth]{figures/results/abundance-vs-complexity-index}
+ \includegraphics[width=\textwidth]{figures/results/abundance-vs-complexity-index-log}
  \caption{\bf Scatter plot of Abundance vs. Complexity index.}
@@
 \end{figure}
-
-Spatiotemporal expression profiles showed that all GRNs reached a steady state
-expression profile before $t = 250$ and most of them indeed reached the steady
-state before $t = 150$ (Fig~\ref{fig:profile}).
 

### 2021-01-31 docs/manuscript/PLOS-MSc-Thesis.tex
Improve introduction

@@
 as those in which a group of cells differentiate in response to a signalling
-molecule such as a morphogen \cite{Turing1952, Wolpert1969}. However, as
-information processing by GRNs is often a complex process, one major goal of
-developmental biology is to understand mechanistically how positional
-information is translated in spatial differentiation.\\
+molecule. Such molecules were referred to as morphogens by Turing in 1952
+\cite{Turing1952} and posterior theoretical studies on morphogens led to the
+french flag model \cite{Wolpert1969}, in which a gradient of concentration of
+the morphogen induces differentiation in a one-dimensional field of cells.
+Although the Bicoid protein (bcd), the first example of a molecule that acted
+as a morphogen, was only found in the 80's in the developing embryos of
+\textit{Drosophila melanogaster}
+\cite{driever_gradient_1988, driever_bicoid_1988}, now it is known that there
+are many other examples such as the Decapentaplegic (Dpp) protein in
+\textit{Drosophila} wing imaginal discs \cite{affolter_decapentaplegic_2007},
+as well as Hedgehog and Wnt.
+However, information processing by GRNs has proven to be a complex process, and
+a major goal of developmental biology is to understand mechanistically how
+positional information given by morphogens is translated in spatial
+differentiation.\\
 
@@
 
-\begin{figure}[!h]
- \includegraphics[width=\textwidth]{figures/results/GRN_660_topol_33_spatio-temp_exp_profile}
- \caption{\bf Most GRNs reach the steady state before 250 time steps.}
- GRN number 660 displaying the topology number 33 (A) and its spatiotemporal
- expression profile (B).
- \label{fig:profile}
-\end{figure}
-
 Hierarchical clustering of spatiotemporal expression profiles showed that GRNs
@@
 knowledge the first time a GRN with a topology such as that of topology nine
+(Fig \ref{fig:topol9})
 and its expression mechanism are reported, increasing our knowledge of the
-genotype-phenotype map in this kind of system.
-
-However, it is still necessary to confirm experimentally the ability of this
-topology to generate the described pattern and it would be interesting to
-study its feasibility and possible implementation in synthetic gene regulatory
-networks.
+genotype-phenotype map in this kind of system, and it would be very interesting
+to confirm experimentally the ability of this topology to generate the described
+pattern and it would be interesting to study its feasibility and possible
+implementation in synthetic gene regulatory networks.
 
@@
 {\bf Dynamics of expression of topology number 9.}
+
+\paragraph*{S3 Fig.}
+\begin{figure}[!h]
+ \includegraphics[width=\textwidth]{figures/supplementary-figures/GRN_660_topol_33_spatio-temp_exp_profile}
+ \caption{\bf Most GRNs reach the steady state before 250 time steps.}
+ GRN number 660 displaying the topology number 33 (A) and its spatiotemporal
+ expression profile (B).
+ \label{S3_Fig}
+\end{figure}
 

### 2021-02-10 docs/manuscript/PLOS-MSc-Thesis.tex
Improve introduction, figures and references

@@
 
-\subsection*{Conclusions}
+% \subsection*{Conclusions}
 
@@
 \textit{Drosophila} wing imaginal discs \cite{affolter_decapentaplegic_2007},
-as well as Hedgehog and Wnt.
-However, information processing by GRNs has proven to be a complex process, and
-a major goal of developmental biology is to understand mechanistically how
-positional information given by morphogens is translated in spatial
-differentiation.\\
+as well as Sonic Hedgehog \cite{dessaud_dynamic_2010,cohen_theoretical_2014}
+and Wnt \cite{raspopovic_digit_2014}. However, information processing by GRNs
+has proven to be a complex process, and a major goal of developmental biology
+is to understand mechanistically how positional information given by morphogens
+is translated in spatial differentiation.\\
 
@@
 interconnections occurring in networks at numbers that are significantly higher
-than those in randomized networks \cite{Milo2002}. This novel approach, along
-with mathematical modeling and computational methods have been shown to be a
-powerful tool to study GRNs and the predictions of studies using these methods
-have been validated by experimental studies
-\cite{Kalir2005,Kalir2004,Mangan2003a,ODonnell2005}, and some of them
-have even used synthetic GRNs to produce artificial cell differentiation~\cite{Basu2005}
-and circadian gene expression~\cite{atkinson_development_2003}. Moreover,
-the design and implementation of GRNs through synthetic biology is emerging as
-a promising tool to study biological phenomena as pattern formation
-\cite{santos-moreno_using_2019}.
+than those in randomized networks~\cite{Milo2002}. This novel approach, along
+with mathematical modeling and computational systems biology methods have been
+shown to be a powerful tool to study GRNs~\cite{kitano_computational_2002}, and
+the predictions of studies using these methods have been validated by
+experimental studies~\cite{Kalir2005,Kalir2004,Mangan2003a,ODonnell2005}, and
+some of them have even used synthetic GRNs to produce artificial cell
+differentiation~\cite{Basu2005}, circadian gene
+expression~\cite{atkinson_development_2003}, counting
+devices~\cite{friedland_synthetic_2009} and systems that respond to light.
+Moreover, the design and implementation of GRNs through synthetic biology is
+emerging as a promising tool to study biological phenomena as pattern formation
+\cite{santos-moreno_using_2019} and as a novel therapeutic tool with amazing
+biomedical applications~\cite{karlsson_therapeutic_2012,
+higashikuni_advancing_2017,abil_synthetic_2015,healy_genetic_2019,
+kitada_programming_2018}.
 
@@
 \mathbb{R}^{3\times4}$ where $i$ represents a regulated node of the GRN and $j$
-represents a regulator node of the GRN Fig \ref{fig:adjmat} with
-$w_{\textit{ij}}~\in [-10,\ 10]$. Values in these matrices can be negative, zero
-or positive and represent represion, no interaction, or activation respectively.
-The magnitude of the value is proportional to the strength of the interaction.\\
+represents a regulator node of the GRN (Fig \ref{fig:model}A and
+\ref{fig:model}B) with $w_{\textit{ij}}~\in [-10,\ 10]$. Values in these
+matrices can be negative, zero or positive and represent represion, no
+interaction, or activation respectively. The magnitude of the value is
+proportional to the strength of the interaction.\\
 
@@
 %  \centering
- \includegraphics[width=\textwidth]{figures/metodos/matriz-de-adyacencia}
+ \includegraphics[width=\textwidth]{figures/metodos/fig_1}
     \caption{\bf Modelling of GRNs.}
-    A: GRN represented as a directed graph with activation (blue arrows) and
+    (A) GRN represented as a directed graph with activation (blue arrows) and
     inhibition (red arrows) interactions. Black circles
@@
     \textbf{C} represent genes in the network.
-    B: The same GRN represented as an adjacency matrix in which positive values
+    (B) The same GRN represented as an adjacency matrix in which positive values
     represent activation, negative ones represent inhibition and
     zeros represent no interaction.
- \label{fig:adjmat}
+    (C) Morphogenetic field. In our model the field was composed of a linear
+    array of 30 isogenic cells and a morphogen gradient concentration described
+    by an exponential decay function. Horizontal arrows between cells represent
+    diffusion of gene products between adjacent cells. Optimal gene expression
+    pattern is shown top right.
+    (D) Spatiotemporal expression profile of a GRN.
+ \label{fig:model}
 \end{figure}
@@
 an unidimensional array of 30 isogenic cells exposed to the morphogen
-concentration gradient (Fig~\ref{fig:morfgrad}). The initial concentration of
+concentration gradient (Fig~\ref{fig:model}C). The initial concentration of
 each gen product in all cells was set to 0.1 in all simulations.\\
-
-\begin{figure}[!h]
- \includegraphics[width=\textwidth]{figures/metodos/morphogen-gradient}
- \caption{\bf Morphogenetic field.}
- In our model the field was composed of a linear array of 30 isogenic cells and
- a morphogen gradient concentration described by an exponential decay function.
- Horizontal arrows between cells represent diffusion of gene products between
- adjacent cells.
- \label{fig:morfgrad}
-\end{figure}
 
@@
 levels greater than 90\% of the maximal level observed along the entire field
-for the output gene (Fig~\ref{fig:exp-pattern}).
-
-\begin{figure}[!h]
- \includegraphics[width=\textwidth]{figures/metodos/optimal-pattern}
- \caption{\bf Optimal gene expression pattern.}
- \label{fig:exp-pattern}
-\end{figure}
+for the output gene (Fig~\ref{fig:model}C).
 
@@
 \begin{figure}[!h]
- \center
- \includegraphics[width=\textwidth]{figures/results/distribution_topologies_en}
+ \includegraphics[width=\textwidth]{figures/results/distribution_most_abundant_topologies}
  \caption{\bf Distribution of abundances of network topologies.}
+ Most frequent network topologies shown in the grid contain the I3-FFL network
+ motif. Topologies are numbered in order of decreasing abundance. The number of
+ GRNs belonging to each network topology is presented below topology number.
+ Note that network topology number 9 produces the striped phenotype using a
+ two-node GRN.
  \label{fig:distopol}
@@
 
-The most abundant network topologies are shown in Fig~\ref{fig:topol}.
+The most abundant network topologies are shown in Fig~\ref{fig:distopol}.
 Interestingly, among the 15 most abundant network topologies, eleven of them
@@
 with only two nodes.
-
-\begin{figure}[!h]
- \includegraphics[width=\textwidth]{figures/results/most-abu-topol}
- \caption{\bf Most frequent network topologies.}
- Most frequent network topologies contain the I3-FFL network motif. Topologies
- are numbered in order of decreasing abundance.
- The number of GRNs belonging to each network topology is presented below
- topology number.
- Network topology number 9 produces the striped phenotype using a two-node GRN.
- \label{fig:topol}
-\end{figure}
 
@@
 number of nodes with low degree and a few nodes with high degree
-(Fig~\ref{fig:deg-dist}). Nodes corresponding to topologies number one and three
-showed the highest degree ($deg(1) = deg (3) = 14$). Curiously, nodes with
+(Fig~\ref{fig:deg-dist}A). Nodes corresponding to topologies number one and
+three showed the highest degree ($deg(1) = deg (3) = 14$). Curiously, nodes with
 higher degree resulted to be also those network topologies with higher abundance
-(Fig~\ref{fig:ab-deg}). This relationship between abundance and node degree
+(Fig~\ref{fig:deg-dist}B). This relationship between abundance and node degree
 was corroborated by the Spearman's rank correlation coefficient
@@
 \begin{figure}[!h]
- \includegraphics[width=\textwidth]{figures/results/neutral-network-degree-distribution}
- \caption{\bf Histogram of vertex degree distribution in the neutral network.}
+ \includegraphics[width=\textwidth]{figures/results/vertex_degree_dist-abundance_vs_node_degree}
+ \caption{\bf Node degree in the neutral network.}
+ (A) Histogram of vertex degree distribution in the neutral network.
+ (B) Scatter plot of Topology abundance vs. Node degree in the neutral
+ network.
  \label{fig:deg-dist}
-\end{figure}
-
-\begin{figure}[!h]
- \includegraphics[width=\textwidth]{figures/results/abundance-vs-node-degree}
- \caption{\bf Scatter plot of Topology abundance vs. Node degree in the neutral
- network.}
- \label{fig:ab-deg}
 \end{figure}
@@
 these two variables ($r_S = −0.0389724$, $\text{p-value} = 0.298475$)
-(Fig~\ref{fig:ab-rob}). However, when we grouped those topologies containing
+(Fig~\ref{fig:ab-rob}A). However, when we grouped those topologies containing
 the same subgraph and calculated their mean robustness, we found that between
@@
 \begin{figure}[!h]
- \includegraphics[width=\textwidth]{figures/results/abundance-vs-robustness}
- \caption{\bf Scatter plot of Abundance vs. robustness.}
- The robustness of each topology was calculated as the number of perturbations
- in which the fitness was greater or equal to 0.95 with respect to the total of
- perturbations performed.
+ \includegraphics[width=\textwidth]{figures/results/abundance-robustness-complexity-index}
+ \caption{\bf Relationship between Abundance and robustness and complexity
+ index.} (A) Abundance vs. Robustness. The robustness of each topology was
+ calculated as the number of perturbations in which the fitness was greater or
+ equal to 0.95 with respect to the total of perturbations performed.
+ (B) Abundance vs. Complexity index. Each point in the plot represents one of
+ the 714 network topologies.
  \label{fig:ab-rob}
@@
 and was found to be negatively correlated with the abundance ($r_S = -0.316653$,
-$\text{p-value} = 2.25161\times10^{-18}$) (Fig~\ref{fig:ab-comp}).
-
-\begin{figure}[!h]
- \includegraphics[width=\textwidth]{figures/results/abundance-vs-complexity-index-log}
- \caption{\bf Scatter plot of Abundance vs. Complexity index.}
- Each point in the plot represents one of the 714 network topologies.
- \label{fig:ab-comp}
-\end{figure}
+$\text{p-value} = 2.25161\times10^{-18}$) (Fig~\ref{fig:ab-rob}B).
 
@@
 together but instead we found GRNs with different topologies grouped as close
-neighbors (Fig~\ref{fig:exp-tree}). More interestingly, some GRNs displaying
+neighbors (Fig~\ref{fig:clustering}A). More interestingly, some GRNs displaying
 very abundant network topologies were grouped together with GRNs displaying
@@
 \begin{figure}[!h]
- \includegraphics[width=\textwidth]{figures/results/spatio-temporal-profile-clustering-i3-ffl}
- \caption{\bf Clustering of spatiotemporal expression profiles obtained by
- neighbor joining showing the presence of I3-FFL network motif.}
- \label{fig:exp-tree}
+ \includegraphics[width=\textwidth]{figures/results/clustering_of_grns_and_topologies}
+ \caption{\bf Clustering of expression profiles and topologies.}
+ (A) Clustering of spatiotemporal expression profiles obtained by
+ neighbor joining. (B) Clustering of network topologies. Blue circles
+ correspond to topologies that contain the I3-FFL network motif.
+ \label{fig:clustering}
 \end{figure}
@@
 Clustering of topologies revealed that there are six main groups of topologies
-in our set of 2061 GRNs (Fig~\ref{fig:subg-tree}).
-
-\begin{figure}[!h]
- \includegraphics[width=\textwidth]{figures/results/subgraph-profile-clustering-i3-ffl}
- \caption{\bf Clustering of network topologies.}
- Blue circles correspond to topologies that contain the I3-FFL network motif.
- \label{fig:subg-tree}
-\end{figure}
+in our set of 2061 GRNs (Fig~\ref{fig:clustering}B).
 
@@
 
-%PLOS does not support heading levels beyond the 3rd (no 4th level headings).
-% \subsection*{\lorem\ and \ipsum\ nunc blandit a tortor}
-% \subsubsection*{3rd level heading}
-% Sed suscipit iaculis mi at feugiat.
-%
-% \begin{enumerate}
-% 	\item{react}
-% 	\item{diffuse free particles}
-% 	\item{increment time by dt and go to 1}
-% \end{enumerate}
-%
-%
-% \subsection*{Sed ac quam id nisi malesuada congue}
-%
-% Praesent faucibus semper libero.
-%
-% \begin{itemize}
-% 	\item First bulleted item.
-% 	\item Second bulleted item.
-% 	\item Third bulleted item.
-% \end{itemize}
-
 \section*{Discussion}
+
+This study can help us to understand the mechanisms by which the GRNs work in
+nature as modules that make part of genomes, and it also can provide guidance
+in the construction of new synthetic GRNs in the field of synthetic biology.
 
@@
 
-That the most abundant topologies were found to be part of highly connected
-clusters in the neutral network indicates that these topologies are both robust
-to quantitative changes in the strength of interactions and diffusion and
-degradation parameters as well as to qualitative changes in the topology
-itself.
-
-We found that the I3-FFL is one of the network motifs that appears more
+Robustness is an important feature of patterning mediated by morphogens
+\cite{lo_robust_2015}. In our study, the most abundant topologies were found to
+be part of highly connected clusters in the neutral network, and this fact
+indicates that these topologies are both robust to quantitative changes in the
+strength of interactions and diffusion and degradation parameters as well as to
+qualitative changes in the topology itself. As robustness and evolvability are
+not mutually exclusive properties \cite{wagner_robustness_2008}, it is likely
+that these topologies are also evolvable.
+
+Other studies have shown that Incoherent Feed Forward Loops are able to form
+spatial stripes of gene expression \cite{ishihara_cross_2005}. In particular,
+we found that the I3-FFL is one of the network motifs that appears more
 frequently in these GRNs, as has also been reported by Cotterell \& Sharpe
@@
 
-\section*{Conclusion}
-
-This study can help us to understand the mechanisms by which the GRNs work in
-nature as modules that make part of genomes, and it also can provide guidance
-in the construction of new synthetic GRNs in the field of synthetic biology.
-
-%CO\textsubscript{2} Maecenas convallis mauris sit amet sem ultrices gravida.
-Etiam eget sapien nibh.
-Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula.
-Curabitur fringilla pulvinar lectus consectetur pellentesque.
-Quisque augue sem, tincidunt sit amet feugiat eget, ullamcorper sed velit.
+%\section*{Conclusion}
 
@@
 \paragraph*{S3 Fig.}
-\begin{figure}[!h]
- \includegraphics[width=\textwidth]{figures/supplementary-figures/GRN_660_topol_33_spatio-temp_exp_profile}
- \caption{\bf Most GRNs reach the steady state before 250 time steps.}
+\label{S3_Fig}
+{\bf Most GRNs reach the steady state before 250 time steps.}
 GRN number 660 displaying the topology number 33 (A) and its spatiotemporal
 expression profile (B).
- \label{S3_Fig}
-\end{figure}
-
-% \paragraph*{S1 File.}
-% \label{S1_File}
-% {\bf Lorem ipsum.}  Maecenas convallis mauris sit amet sem ultrices gravida.
-Etiam eget sapien nibh.
-Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula.
-Curabitur fringilla pulvinar lectus consectetur pellentesque.
+
+\paragraph*{S4 Fig.}
+\label{S4_Fig}
+{\bf Bridge topologies in the neutral network.}
 %

### 2021-02-10 docs/manuscript/PLOS-MSc-Thesis.tex
Remove acknow and authors symbols

@@
 \\
-Juan Camilo Arboleda-Rivera\textsuperscript{1\Yinyang},
-Gloria Machado-Rodríguez\textsuperscript{1\Yinyang},
-Boris Rodríguez-Rey\textsuperscript{2\textcurrency},
+Juan Camilo Arboleda-Rivera\textsuperscript{1},
+Gloria Machado-Rodríguez\textsuperscript{1},
+Boris Rodríguez-Rey\textsuperscript{2},
 Jayson Gutiérrez\textsuperscript{3},
@@
 % Primary Equal Contribution Note
-\Yinyang These authors contributed equally to this work.
+% \Yinyang These authors contributed equally to this work.
 
@@
 % Also use this double-dagger symbol for special authorship notes, such as senior authorship.
-\ddag These authors also contributed equally to this work.
+% \ddag These authors also contributed equally to this work.
 
 % Current address notes
-\textcurrency Current Address: Dept/Program/Center, Universidad de Antioquia,
-Medellín, Antioquia, Colombia % change symbol to "\textcurrency a" if more than one current address note
+% \textcurrency Current Address: Dept/Program/Center, Universidad de Antioquia,
+% Medellín, Antioquia, Colombia % change symbol to "\textcurrency a" if more than one current address note
 % \textcurrency b Insert second current address
@@
 % Group/Consortium Author Note
-\textpilcrow Membership list can be found in the Acknowledgments section.
+% \textpilcrow Membership list can be found in the Acknowledgments section.
 
@@
 
-\section*{Acknowledgments}
-Cras egestas velit mauris, eu mollis turpis pellentesque sit amet.
-Interdum et malesuada fames ac ante ipsum primis in faucibus.
-Nam id pretium nisi.
-Sed ac quam id nisi malesuada congue.
-Sed interdum aliquet augue, at pellentesque quam rhoncus vitae.
-
 \nolinenumbers

### 2021-02-10 docs/manuscript/PLOS-MSc-Thesis.tex
Implement new figures in TIFF

@@
 %  \centering
- \includegraphics[width=\textwidth]{figures/metodos/fig_1}
+ \includegraphics[width=\textwidth]{figures/metodos/Fig1}
     \caption{\bf Modelling of GRNs.}
@@
  \item{\bf Evaluate fitness by comparing the GRN phenotype with the optimal
- phenotype (Fig~\ref{fig:exp-pattern}).} In order to compute the fitness
+ phenotype (Fig~\ref{fig:model}C).} In order to compute the fitness
  function for each GRN we used three different filters. The first filter
@@
 \begin{figure}[!h]
- \includegraphics[width=\textwidth]{figures/results/distribution_most_abundant_topologies}
+ \includegraphics[width=\textwidth]{figures/results/Fig2}
  \caption{\bf Distribution of abundances of network topologies.}
@@
 \begin{figure}[!h]
- \includegraphics[width=\textwidth]{figures/results/grafo-hamming-15-nodes-inhibition-bar}
+ \includegraphics[width=\textwidth]{figures/results/Fig3}
  \caption{\bf Neutral network.}
@@
 \begin{figure}[!h]
- \includegraphics[width=\textwidth]{figures/results/vertex_degree_dist-abundance_vs_node_degree}
+ \includegraphics[width=\textwidth]{figures/results/Fig4}
  \caption{\bf Node degree in the neutral network.}
@@
 \begin{figure}
- \includegraphics[width=\textwidth]{figures/results/hamming-2pasos}
+ \includegraphics[width=\textwidth]{figures/results/Fig5}
  \caption{\bf Network of topologies separated by two mutational steps.}
@@
 \begin{figure}[!h]
- \includegraphics[width=\textwidth]{figures/results/abundance-robustness-complexity-index}
+ \includegraphics[width=\textwidth]{figures/results/Fig6}
  \caption{\bf Relationship between Abundance and robustness and complexity
@@
 \begin{figure}[!h]
- \includegraphics[width=\textwidth]{figures/results/clustering_of_grns_and_topologies}
+ \includegraphics[width=\textwidth]{figures/results/Fig7}
  \caption{\bf Clustering of expression profiles and topologies.}
@@
 \begin{figure}[!h]
- \includegraphics[width=\textwidth]{figures/results/topol-9}
+ \includegraphics[width=\textwidth]{figures/results/Fig8}
  \caption{\bf Expression dynamics of topology 9.}
@@
 % Include only the SI item label in the paragraph heading. Use the \nameref{label} command to cite SI items in the text.
+
 \paragraph*{S1 Fig.}
-\label{S1_Fig}
-{\bf Relevant subgraphs present in the set of GRNs.}
-%Add descriptive text after the title of the item (optional).
-
-\paragraph*{S2 Fig.}
 \label{S2_Fig}
@@
 
-\paragraph*{S3 Fig.}
+\paragraph*{S2 Fig.}
 \label{S3_Fig}
@@
 
-\paragraph*{S4 Fig.}
+\paragraph*{S3 Fig.}
 \label{S4_Fig}
 {\bf Bridge topologies in the neutral network.}
-%
-% \paragraph*{S1 Video.}
-% \label{S1_Video}
-% {\bf Lorem ipsum.}  Maecenas convallis mauris sit amet sem ultrices gravida.
-Etiam eget sapien nibh.
-Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula.
-Curabitur fringilla pulvinar lectus consectetur pellentesque.
-%
-% \paragraph*{S1 Appendix.}
-% \label{S1_Appendix}
-% {\bf Lorem ipsum.} Maecenas convallis mauris sit amet sem ultrices gravida.
-Etiam eget sapien nibh.
-Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula.
-Curabitur fringilla pulvinar lectus consectetur pellentesque.
+
+\paragraph*{S4 Fig.}
+\label{S1_Fig}
+{\bf Relevant subgraphs present in the set of GRNs.}
 

### 2021-04-22 docs/manuscript/PLOS-MSc-Thesis.tex
Add asterisk to main author

@@
 \\
-Juan Camilo Arboleda-Rivera\textsuperscript{1},
+Juan Camilo Arboleda-Rivera\textsuperscript{1*},
 Gloria Machado-Rodríguez\textsuperscript{1},

### 2021-04-22 modules/VideosExpressionDynamics.m
Translate VideosExpressionDynamics.m

@@ -1,23 +1,19 @@
 (* ::Package:: *)
 
-(*Paquete para analizar la dinámica de expresión de cada genotipo.
+(*Package to analyze the expression dynamics of each genotype.
 This package depends on EvolAlgorithm4PatternFormingGRCModel2MathPackageV9*)
 <<EvolAlgorithm4PatternFormingGRCModel2MathPackageV9`
 
-videoExp::usage = "videoExp[ genotipo ] crea un video de la dinámica de
-expresión de un genotipo, acepta como argumentos 'gen', con las opciones 'a','b'
- y 'c', tamaño del gráfico y 'fps' (cuadros por segundo)."
+videoExp::usage = "videoExp[ genotipo ] create a video of the expression dynamics of a GRN, it takes as argument 'gen', with options 'a','b' and 'c', image size and 'fps' (frames per second)."
 
-(*Esta línea importa la lista de motivos clasificados, ordenados no isomorfos*)
+(*Import list of classified, ordered, non isomorphic motifs*)
 motivosClasificadosOrdenadosNoIsomorfos = Flatten[ToExpression[#],1] &/@
-Import["/home/nesper94/Documentos/biologia-de-sistemas/RobustnessModularityProject/RobustnessModularityProject/motivos_clasificados_ordenados_no_isomorfos.txt","TSV"];
+Import["../data/ordered-non-isomorph-grns.txt","TSV"];
 
-(*Esta función me permite saber en qué posiciones están los genotipos que pertenecen a determinado motivo de red*)
-posiciónMotivos[motivo_Integer]:=Flatten[{First[#],Last[#]}&/@{Position[motivosClasificadosOrdenadosNoIsomorfos[[All,2]],motivo]}]
+(*This function shows the positions of GRNs with some topology*)
+posici\[OAcute]nMotivos[topology_Integer]:=Flatten[{First[#],Last[#]}&/@{Position[motivosClasificadosOrdenadosNoIsomorfos[[All,2]],topology]}]
 
-(*Aquí redefino la función "SpaceTimeFeats4StripeFormingGRCs4SSMorpGradientWithSumAndFilterModel0" de tal manera que acepte los genotipos sin necesidad de usar como argumento el input de morfógeno*)
[…]

### 2021-04-22 docs/manuscript/PLOS-MSc-Thesis.tex
Fix typo: wich -> which

@@
 
-The neutral network (a metagraph in wich each node is a network topology)
+The neutral network (a metagraph in which each node is a network topology)
 shows that most of the network topologies are grouped into a connected graph.

### 2021-04-27 docs/manuscript/PLOS-MSc-Thesis.tex
Mention Cotterell and Sharpe work in introduction

@@
 
+Cotterell and Sharpe studied what kind of 3-node network topologies could
+interact with a morphogen gradient and produce a stripe of gene expression
+in a one-dimensional field of cells and found a variety of networks that
+implemented at least 7 different mechanisms and had different levels of
+complexity \cite{Cotterell2010}.
+
 However, although several studies have analyzed the properties of GRNs that

### 2021-04-27 docs/manuscript/PLOS-MSc-Thesis.tex
Rewrite French Flag section in light of Sharpe 2019 paper

@@
 molecule. Such molecules were referred to as morphogens by Turing in 1952
-\cite{Turing1952} and posterior theoretical studies on morphogens led to the
-french flag model \cite{Wolpert1969}, in which a gradient of concentration of
-the morphogen induces differentiation in a one-dimensional field of cells.
+\cite{Turing1952} and posterior theoretical studies on patterning led to the
+conceptualization of the French Flag Problem by Wolpert, who also stated
+in this respect that a gradient of concentration of a morphogen could
+trigger cell differentiation in a one-dimensional field of cells
+\cite{Wolpert1969, sharpe_2019}.
 Although the Bicoid protein (bcd), the first example of a molecule that acted

### 2021-04-28 example/EvolAlgorithm4PatternFormingGRCModel2MathPackageV9.m
Use Matplotlib Viridian colormap

@@ -29,32 +29,32 @@ SpaceTimeFeats4StripeFormingGRCs4SSMorpGradientWithSumAndFilterModel0::usage = "
 
 CreatSpaceTimePlots4GRC::usage="Function to create the spatio-temporal expression portraits of the GRCs";
 
-EquilibFilter::usage = "This function is to assess whether the expression profile of the output node in the GRC models 
-						reaches a quasi-steady state. GRC dynamics are simulated over 500 time steps, and the spatial 
-						profile over the total number of nuclei considered attained at time step = 500 is compared to 
-						that attained previously at time step = 250. The difference between expression profiles is 
-						assessed based of the normalized Euclidean Distance. We focus on GRCs capable of achieving a 
-						relatively fast developmental patterning task (single stripe formation) because: 
-						1) computing time restrictions; and 2) the speed of developmental patterning is likely to be 
-						critical during embryogenesis (potential fitness correlate). 
-						Ideally, the value of this score should be \[LessEqual] 0.001, which is indicative that the expression 
+EquilibFilter::usage = "This function is to assess whether the expression profile of the output node in the GRC models
+						reaches a quasi-steady state. GRC dynamics are simulated over 500 time steps, and the spatial
+						profile over the total number of nuclei considered attained at time step = 500 is compared to
+						that attained previously at time step = 250. The difference between expression profiles is
+						assessed based of the normalized Euclidean Distance. We focus on GRCs capable of achieving a
+						relatively fast developmental patterning task (single stripe formation) because:
+						1) computing time restrictions; and 2) the speed of developmental patterning is likely to be
+						critical during embryogenesis (potential fitness correlate).
+						Ideally, the value of this score should be \[LessEqual] 0.001, which is indicative that the expression
 						profile at time tfinal has reached steady state";
 
-PatternFilter::usage = "A patterning score describes how much \[OpenCurlyQuote]pattern\[CloseCurlyQuote] or spatial heterogeneity there is. Ideally, a value 
[…]

### 2021-04-28 modules/EvolAlgorithm4PatternFormingGRCModel2MathPackageV9.m

@@ -74,6 +74,9 @@ Begin["`Private`"]
 
 (**************************************************************************************************************)
 
+(*Load Viridis ColorMap*)
+ClearAll[MPLColorMap] << "http://pastebin.com/raw/pFsb4ZBS";
+
 ColorF01[Node_] := Switch[Node,"A",Darker[Gray,0.2],"B",Darker[Gray,0.2],"C",Darker[Gray,0.2],"\!\(\*SubscriptBox[\(M\), \(\!\(\*
 StyleBox[\"A\",\nFontSize->0]\)\)]\)",Black,"\!\(\*SubscriptBox[\(M\), \(\!\(\*
 StyleBox[\"B\",\nFontSize->0]\)\)]\)",Black,"\!\(\*SubscriptBox[\(M\), \(\!\(\*
@@ -221,9 +224,9 @@ GRCPhenotReadout= Transpose[Thread[Table[Flatten[Table[Flatten[Evaluate[Prot[i,n
 
 (**************************************************************************************************************)
 
-colorbar[{min_,max_},divs_: 150]:=DensityPlot[y,{x,0,0.1},{y,min,max},AspectRatio->12,PlotRangePadding->0,PlotPoints->{2,divs},MaxRecursion->0,FrameTicks->{None,Automatic,None,None},ColorFunction->ColorData["TemperatureMap"]]
+colorbar[{min_,max_},divs_: 150]:=DensityPlot[y,{x,0,0.1},{y,min,max},AspectRatio->12,PlotRangePadding->0,PlotPoints->{2,divs},MaxRecursion->0,FrameTicks->{None,Automatic,None,None},ColorFunction->MPLColorMap["Viridis"]]
 
-CreatSpaceTimePlots4GRC[TimeSeries_List]:=GraphicsRow[MapIndexed[With[{opts={ImageSize->{Automatic,130},ImagePadding->5}},ArrayPlot[#,ColorFunction->ColorData["TemperatureMap"],FrameTicks->None,opts]]&,((#/Max[Flatten[#]])&/@TimeSeries)],Spacings->{-10,0}]
+CreatSpaceTimePlots4GRC[TimeSeries_List]:=GraphicsRow[MapIndexed[With[{opts={ImageSize->{Automatic,130},ImagePadding->5}},ArrayPlot[#,ColorFunction->MPLColorMap["Viridis"],FrameTicks->None,opts]]&,((#/Max[Flatten[#]])&/@TimeSeries)],Spacings->{-10,0}]
 
 (**************************************************************************************************************)
[…]

### 2021-05-04 docs/manuscript/PLOS-MSc-Thesis.tex
Explain heatmap in Fig1.D
Improve Fig3 caption
- Explain that each node is a topology
- Say that most abundant topologies are depicted as
in Fig1.

@@
     pattern is shown top right.
-    (D) Spatiotemporal expression profile of a GRN.
+    (D) Heatmap of spatiotemporal expression profile of a GRN where blue
+    colors represent lower expression and red colors represent higher
+    expression.
  \label{fig:model}
@@
  \caption{\bf Neutral network.}
- This graph shows network topologies connected by one mutational step. Most
- abundant network topologies are shown as black nodes. All nodes with the same
+ This graph shows network topologies connected by one mutational step, where
+ each node represents a single topology. Black nodes represent the most
+ abundant topologies. All nodes with the same
  color belong to a cluster of nodes more connected between them than with other
- nodes in the graph.
+ nodes in the graph. The most abundant topologies are represented as
+ directed graphs in the same way as in Fig~\ref{fig:model}A, except for
+ the topologies 1 to 8 and 15 (orange and red clusters), in which only
+ topology 1 is depicted.
  \label{fig:neutral-network}

### 2021-05-04 docs/manuscript/PLOS-MSc-Thesis.tex
Improve Fig5 caption
- Add that black circles are as in Fig3
- Add that colors represent clusters.

@@
  \caption{\bf Network of topologies separated by two mutational steps.}
- The network shows that all the topologies are connected by two or less
- mutational steps, except for topology number 420 (big red circle). Big black
- circles correspond to the 15 topologies with higher abundance.
+ The network shows that almost all the topologies are connected by two or less
+ mutational steps. Big red circle represents topology 420 that remains
+ unconnected from the main graph. Big black circles correspond to the 15
+ most abundant topologies described in Fig~\ref{fig:neutral-network}.
+ Different colors in the graph represent different clusters of topologies.
  \label{fig:2neut-net}

### 2021-05-04 docs/manuscript/PLOS-MSc-Thesis.tex
Fix negative sign in correlation coefficient
Response to Reviewer #1, MINOR POINTS, 2c

--- 2021-04-27 docs/manuscript/PLOS-MSc-Thesis.tex
@@
 was corroborated by the Spearman's rank correlation coefficient
-($r_S = −0.353752$, $p = 6.2933\times10^{-23} $).
+($r_S = 0.353752$, $p = 6.2933\times10^{-23} $).
 

### 2021-05-04 docs/manuscript/PLOS-MSc-Thesis.tex
Improve Fig6 caption
- A) Change "number" for "the ratio" to explain better how robutsness
was calculated.
- B) Add main results related to this graph.

--- 2021-05-04 docs/manuscript/PLOS-MSc-Thesis.tex
@@
  index.} (A) Abundance vs. Robustness. The robustness of each topology was
- calculated as the number of perturbations in which the fitness was greater or
+ calculated as the ratio of perturbations in which the fitness was greater or
  equal to 0.95 with respect to the total of perturbations performed.
  (B) Abundance vs. Complexity index. Each point in the plot represents one of
- the 714 network topologies.
+ the 714 network topologies. More complex topologies tend to be less abundant
+ than simple topologies
+ ($r_S = -0.316653$, $\text{p-value} = 2.25161\times10^{-18}$).
  \label{fig:ab-rob}

### 2021-05-09 docs/manuscript/PLOS-MSc-Thesis.tex
Implement highlighting of changes

--- 2021-05-04 docs/manuscript/PLOS-MSc-Thesis.tex
@@
 %\newcommand{\ipsum}{{\bf IPSUM}}
+
+% Highlight changes
+\usepackage{color,soul}
+\newcommand{\hil}[1]{\hl{#1}} % Highlight changes
+% \newcommand{\hil}[1]{#1} % Do nothing, clean version
 
@@
 molecule. Such molecules were referred to as morphogens by Turing in 1952
-\cite{Turing1952} and posterior theoretical studies on patterning led to the
+\cite{Turing1952} \hil{and posterior theoretical studies on patterning led to the
 conceptualization of the French Flag Problem by Wolpert, who also stated
 in this respect that a gradient of concentration of a morphogen could
-trigger cell differentiation in a one-dimensional field of cells
+trigger cell differentiation in a one-dimensional field of cells}
 \cite{Wolpert1969, sharpe_2019}.
@@
 
-Cotterell and Sharpe studied what kind of 3-node network topologies could
+\hil{Cotterell and Sharpe studied what kind of 3-node network topologies could
 interact with a morphogen gradient and produce a stripe of gene expression
@@
 implemented at least 7 different mechanisms and had different levels of
-complexity \cite{Cotterell2010}.
+complexity} \cite{Cotterell2010}.
 
@@
 was corroborated by the Spearman's rank correlation coefficient
-($r_S = 0.353752$, $p = 6.2933\times10^{-23} $).
+(\hil{$r_S = 0.353752$}, $p = 6.2933\times10^{-23} $).
 

### 2021-05-10 docs/manuscript/PLOS-MSc-Thesis.tex
Improve Fig8 legend

--- 2021-05-04 docs/manuscript/PLOS-MSc-Thesis.tex
@@
  \caption{\bf Expression dynamics of topology 9.}
- A: Network topology number 9. B: Spatiotemporal expression profile of a GRN
- with topology 9.
+ (A) Network topology number 9. Although this topology consists only of two
+ genes, it display a striped pattern of gene expression for both of them as
+ can be observed in (B). (B) Spatiotemporal expression profile of a GRN
+ with topology 9. Gene \textbf{B} is not shown as it was lost in this topology
+ in the evolutionary process simulated by the genetic algorithm.
  \label{fig:topol9}

### 2021-05-11 docs/manuscript/PLOS-MSc-Thesis.tex
Highlight changes in figure legends

--- 2021-05-09 docs/manuscript/PLOS-MSc-Thesis.tex
@@
     pattern is shown top right.
-    (D) Spatiotemporal expression profile of a GRN.
+    \hil{(D) Heatmap of spatiotemporal expression profile of a GRN where blue
+    colors represent lower expression and red colors represent higher
+    expression.}
  \label{fig:model}
@@
  \caption{\bf Neutral network.}
- This graph shows network topologies connected by one mutational step. Most
- abundant network topologies are shown as black nodes. All nodes with the same
+ \hil{This graph shows network topologies connected by one mutational step, where
+ each node represents a single topology. Black nodes represent the most
+ abundant topologies. All nodes with the same}
  color belong to a cluster of nodes more connected between them than with other
- nodes in the graph.
+ nodes in the graph. \hil{The most abundant topologies are represented as
+ directed graphs in the same way as in Fig}~\ref{fig:model}A, \hil{except for
+ the topologies 1 to 8 and 15 (orange and red clusters), in which only
+ topology 1 is depicted.}
  \label{fig:neutral-network}
@@
  \caption{\bf Network of topologies separated by two mutational steps.}
- The network shows that all the topologies are connected by two or less
- mutational steps, except for topology number 420 (big red circle). Big black
- circles correspond to the 15 topologies with higher abundance.
+ \hil{The network shows that almost all the topologies are connected by two or less
+ mutational steps. Big red circle represents topology 420 that remains
+ unconnected from the main graph. Big black circles correspond to the 15
+ most abundant topologies described in} Fig~\ref{fig:neutral-network}.
+ \hil{Different colors in the graph represent different clusters of topologies.}
  \label{fig:2neut-net}
@@
  index.} (A) Abundance vs. Robustness. The robustness of each topology was
- calculated as the number of perturbations in which the fitness was greater or
+ calculated as the \hil{ratio} of perturbations in which the fitness was greater or
  equal to 0.95 with respect to the total of perturbations performed.
  (B) Abundance vs. Complexity index. Each point in the plot represents one of
- the 714 network topologies.
+ the 714 network topologies. \hil{More complex topologies tend to be less abundant
+ than simple topologies
+ ($r_S = -0.316653$, $\text{p-value} = 2.25161\times10^{-18}$).}
  \label{fig:ab-rob}
@@
  \caption{\bf Expression dynamics of topology 9.}
- A: Network topology number 9. B: Spatiotemporal expression profile of a GRN
- with topology 9.
+ \hil{(A) Network topology number 9. Although this topology consists only of two
+ genes, it display a striped pattern of gene expression for both of them as
+ can be observed in (B). (B) Spatiotemporal expression profile of a GRN
+ with topology 9. Gene \textbf{B} is not shown as it was lost in this topology
+ in the evolutionary process simulated by the genetic algorithm.}
  \label{fig:topol9}

### 2021-05-11 docs/manuscript/PLOS-MSc-Thesis.tex
Add definition of genotype

--- 2021-05-11 docs/manuscript/PLOS-MSc-Thesis.tex
@@
 interaction, or activation respectively. The magnitude of the value is
-proportional to the strength of the interaction.\\
+proportional to the strength of the interaction. \hil{In this study we define
+a 'genotype' as the set of interaction, diffusion and degradation values that
+represent a GRN.}\\
 
@@
     pattern is shown top right.
-    (D) Heatmap of spatiotemporal expression profile of a GRN where blue
-    colors represent lower expression and red colors represent higher
-    expression.
+    (D) Spatiotemporal expression profile of a GRN.
  \label{fig:model}
@@
  \caption{\bf Neutral network.}
- This graph shows network topologies connected by one mutational step, where
- each node represents a single topology. Black nodes represent the most
- abundant topologies. All nodes with the same
+ This graph shows network topologies connected by one mutational step. Most
+ abundant network topologies are shown as black nodes. All nodes with the same
  color belong to a cluster of nodes more connected between them than with other
- nodes in the graph. The most abundant topologies are represented as
- directed graphs in the same way as in Fig~\ref{fig:model}A, except for
- the topologies 1 to 8 and 15 (orange and red clusters), in which only
- topology 1 is depicted.
+ nodes in the graph.
  \label{fig:neutral-network}
@@
  \caption{\bf Network of topologies separated by two mutational steps.}
- The network shows that almost all the topologies are connected by two or less
- mutational steps. Big red circle represents topology 420 that remains
- unconnected from the main graph. Big black circles correspond to the 15
- most abundant topologies described in Fig~\ref{fig:neutral-network}.
- Different colors in the graph represent different clusters of topologies.
+ The network shows that all the topologies are connected by two or less
+ mutational steps, except for topology number 420 (big red circle). Big black
+ circles correspond to the 15 topologies with higher abundance.
  \label{fig:2neut-net}
@@
  index.} (A) Abundance vs. Robustness. The robustness of each topology was
- calculated as the ratio of perturbations in which the fitness was greater or
+ calculated as the number of perturbations in which the fitness was greater or
  equal to 0.95 with respect to the total of perturbations performed.
  (B) Abundance vs. Complexity index. Each point in the plot represents one of
- the 714 network topologies. More complex topologies tend to be less abundant
- than simple topologies
- ($r_S = -0.316653$, $\text{p-value} = 2.25161\times10^{-18}$).
+ the 714 network topologies.
  \label{fig:ab-rob}
@@
  \caption{\bf Expression dynamics of topology 9.}
- (A) Network topology number 9. Although this topology consists only of two
- genes, it display a striped pattern of gene expression for both of them as
- can be observed in (B). (B) Spatiotemporal expression profile of a GRN
- with topology 9. Gene \textbf{B} is not shown as it was lost in this topology
- in the evolutionary process simulated by the genetic algorithm.
+ A: Network topology number 9. B: Spatiotemporal expression profile of a GRN
+ with topology 9.
  \label{fig:topol9}

### 2021-05-12 docs/manuscript/PLOS-MSc-Thesis.tex
Improve Fig2 caption

--- 2021-05-11 docs/manuscript/PLOS-MSc-Thesis.tex
@@
  \caption{\bf Distribution of abundances of network topologies.}
- Most frequent network topologies shown in the grid contain the I3-FFL network
- motif. Topologies are numbered in order of decreasing abundance. The number of
- GRNs belonging to each network topology is presented below topology number.
+ \hil{Some of the most} frequent network topologies shown in the grid
+ contain the I3-FFL network motif \hil{(shown on the left). In this network
+ motif the gene B is activated in an indirect way by the morphogen through
+ the activation of A (positive interactions), and at the same time, it
+ is inactivated in a direct way by the morphogen (negative interaction).}
+ Topologies are numbered in order of decreasing abundance, \hil{being the
+ Topology 1 the most abundant with a number of 95 genotypes}. The number of
+ \hil{genotypes} belonging to each network topology is presented below
+ topology number.
  Note that network topology number 9 produces the striped phenotype using a

### 2021-05-12 docs/manuscript/PLOS-MSc-Thesis.tex
Add link to Fig 5 when mentioning topology 420

@@
 interaction parameters and that more complex networks are less robusts. For
-example, topology number 420, which presents high complexity, was one of the
+example, topology number 420 \hil{(Fig}~\ref{fig:2neut-net}), which
+presents high complexity, was one of the
 topologies with lower abundance and was disconnected from all other topologies

### 2021-05-12 docs/manuscript/PLOS-MSc-Thesis.tex
Remove mention to 714 to avoid confusion
Response to Reviewer #3, 2.b

--- 2021-05-11 docs/manuscript/PLOS-MSc-Thesis.tex
@@
 In order to classify each topology by its subgraph composition, we first created
-a matrix  $T = (t_{ij})$ with 714 rows corresponding to each different topology
+a matrix  $T = (t_{ij})$ with \hil{each row} corresponding to \hil{a}
+different topology
 and 17 columns corresponding to different subgraphs. Each element $t_{ij}$ of

### 2021-05-14 docs/manuscript/PLOS-MSc-Thesis.tex
Describe properly supplementary fig: expression of topol 9

--- 2021-05-12 docs/manuscript/PLOS-MSc-Thesis.tex
@@
 {\bf Dynamics of expression of topology number 9.}
+The dynamics of expression is shown from $t = 5$ to $t = 150$. The red line
+represents the expression level of the gene C along the morphogenetic field,
+whereas the dotted line represents the expression level of gene A. The
+striped pattern of gene expression can be seen for both genes since $t = 80$.
 

### 2021-05-14 docs/manuscript/PLOS-MSc-Thesis.tex
Relabel supplementary figures in order of appearance

@@
 then begins to activate the gene A also in the center of the field
-(\nameref{S2_Fig}).
+(\nameref{S1_Fig}).
 % (Fig~\ref{fig:topol9-dyn}).
@@
 \paragraph*{S1 Fig.}
-\label{S2_Fig}
+\label{S1_Fig}
 {\bf Dynamics of expression of topology number 9.}
@@
 \paragraph*{S2 Fig.}
-\label{S3_Fig}
+\label{S2_Fig}
 {\bf Most GRNs reach the steady state before 250 time steps.}
@@
 \paragraph*{S3 Fig.}
+\label{S3_Fig}
+{\bf Bridge topologies in the neutral network.}
+
+\paragraph*{S4 Fig.}
 \label{S4_Fig}
-{\bf Bridge topologies in the neutral network.}
-
-\paragraph*{S4 Fig.}
-\label{S1_Fig}
 {\bf Relevant subgraphs present in the set of GRNs.}

### 2021-05-14 docs/manuscript/PLOS-MSc-Thesis.tex
Add description of figure S3

@@
 {\bf Most GRNs reach the steady state before 250 time steps.}
-GRN number 660 displaying the topology number 33 (A) and its spatiotemporal
-expression profile (B).
+(A) Gene regulatory network displaying the topology number 33. (B) Spatiotemporal
+expression profile of the gene regulatory network shown in (A).
 
@@
 {\bf Bridge topologies in the neutral network.}
+``Bridge topologies'' are those topologies that connect two clusters in the neutral
+network (Fig~\ref{fig:neutral-network}), and as such can be interpreted as
+intermediary steps in the evolution from a cluster of related topologies into
+another cluster. For example the topology 441 could be an initial step to reach the
+topology 9, as in this topology nodes B and C are not connected.
 

### 2021-05-14 docs/manuscript/PLOS-MSc-Thesis.tex
Add description of Fig S4 and cite it in text

@@
 a matrix  $T = (t_{ij})$ with 714 rows corresponding to each different topology
-and 17 columns corresponding to different subgraphs. Each element $t_{ij}$ of
+and 17 columns corresponding to subgraphs shown in \nameref{S4_Fig}.
+Each element $t_{ij}$ of
 the matrix was 1 if the \emph{j}-th subgraph was present in the \emph{i}-th
@@
 {\bf Relevant subgraphs present in the set of GRNs.}
+These are subgraphs that have been reported in previous studies as networks
+involved in morphogenesis and development \cite{Cotterell2010, Schaerli2014,
+Schaerli2018}. These were used to calculate the subgraph profile and the results
+reported in \nameref{S1_Table}.
 

### 2021-05-14 docs/manuscript/PLOS-MSc-Thesis.tex

@@
 In order to classify each topology by its subgraph composition, we first created
-a matrix  $T = (t_{ij})$ with 714 rows corresponding to each different topology
+a matrix  $T = (t_{ij})$ with each row corresponding to a different topology
 and 17 columns corresponding to subgraphs shown in \nameref{S4_Fig}.

### 2021-05-14 docs/manuscript/PLOS-MSc-Thesis.tex
Remove \hil command for highlighting changes
Change tracking will be implemented with latexdiff.

@@
 %\newcommand{\ipsum}{{\bf IPSUM}}
-
-% Highlight changes
-\usepackage{color,soul}
-\newcommand{\hil}[1]{\hl{#1}} % Highlight changes
-% \newcommand{\hil}[1]{#1} % Do nothing, clean version
 
@@
 molecule. Such molecules were referred to as morphogens by Turing in 1952
-\cite{Turing1952} \hil{and posterior theoretical studies on patterning led to the
+\cite{Turing1952} and posterior theoretical studies on patterning led to the
 conceptualization of the French Flag Problem by Wolpert, who also stated
 in this respect that a gradient of concentration of a morphogen could
-trigger cell differentiation in a one-dimensional field of cells}
+trigger cell differentiation in a one-dimensional field of cells
 \cite{Wolpert1969, sharpe_2019}.
@@
 
-\hil{Cotterell and Sharpe studied what kind of 3-node network topologies could
+Cotterell and Sharpe studied what kind of 3-node network topologies could
 interact with a morphogen gradient and produce a stripe of gene expression
@@
 implemented at least 7 different mechanisms and had different levels of
-complexity} \cite{Cotterell2010}.
+complexity \cite{Cotterell2010}.
 
@@
 interaction, or activation respectively. The magnitude of the value is
-proportional to the strength of the interaction. \hil{In this study we define
+proportional to the strength of the interaction. In this study we define
 a 'genotype' as the set of interaction, diffusion and degradation values that
-represent a GRN.}\\
+represent a GRN.\\
 
@@
     pattern is shown top right.
-    \hil{(D) Heatmap of spatiotemporal expression profile of a GRN where blue
+    (D) Heatmap of spatiotemporal expression profile of a GRN where blue
     colors represent lower expression and red colors represent higher
-    expression.}
+    expression.
  \label{fig:model}
@@
  \caption{\bf Distribution of abundances of network topologies.}
- \hil{Some of the most} frequent network topologies shown in the grid
- contain the I3-FFL network motif \hil{(shown on the left). In this network
+ Some of the most frequent network topologies shown in the grid
+ contain the I3-FFL network motif (shown on the left). In this network
  motif the gene B is activated in an indirect way by the morphogen through
  the activation of A (positive interactions), and at the same time, it
- is inactivated in a direct way by the morphogen (negative interaction).}
- Topologies are numbered in order of decreasing abundance, \hil{being the
- Topology 1 the most abundant with a number of 95 genotypes}. The number of
- \hil{genotypes} belonging to each network topology is presented below
+ is inactivated in a direct way by the morphogen (negative interaction).
+ Topologies are numbered in order of decreasing abundance, being the
+ Topology 1 the most abundant with a number of 95 genotypes. The number of
+ genotypes belonging to each network topology is presented below
  topology number.
@@
  \caption{\bf Neutral network.}
- \hil{This graph shows network topologies connected by one mutational step, where
+ This graph shows network topologies connected by one mutational step, where
  each node represents a single topology. Black nodes represent the most
- abundant topologies. All nodes with the same}
+ abundant topologies. All nodes with the same
  color belong to a cluster of nodes more connected between them than with other
- nodes in the graph. \hil{The most abundant topologies are represented as
- directed graphs in the same way as in Fig}~\ref{fig:model}A, \hil{except for
+ nodes in the graph. The most abundant topologies are represented as
+ directed graphs in the same way as in Fig~\ref{fig:model}A, except for
  the topologies 1 to 8 and 15 (orange and red clusters), in which only
- topology 1 is depicted.}
+ topology 1 is depicted.
  \label{fig:neutral-network}
@@
 was corroborated by the Spearman's rank correlation coefficient
-(\hil{$r_S = 0.353752$}, $p = 6.2933\times10^{-23} $).
+($r_S = 0.353752$, $p = 6.2933\times10^{-23} $).
 
@@
  \caption{\bf Network of topologies separated by two mutational steps.}
- \hil{The network shows that almost all the topologies are connected by two or less
+ The network shows that almost all the topologies are connected by two or less
  mutational steps. Big red circle represents topology 420 that remains
  unconnected from the main graph. Big black circles correspond to the 15
- most abundant topologies described in} Fig~\ref{fig:neutral-network}.
- \hil{Different colors in the graph represent different clusters of topologies.}
+ most abundant topologies described in Fig~\ref{fig:neutral-network}.
+ Different colors in the graph represent different clusters of topologies.
  \label{fig:2neut-net}
@@
  index.} (A) Abundance vs. Robustness. The robustness of each topology was
- calculated as the \hil{ratio} of perturbations in which the fitness was greater or
+ calculated as the ratio of perturbations in which the fitness was greater or
  equal to 0.95 with respect to the total of perturbations performed.
  (B) Abundance vs. Complexity index. Each point in the plot represents one of
- the 714 network topologies. \hil{More complex topologies tend to be less abundant
+ the 714 network topologies. More complex topologies tend to be less abundant
  than simple topologies
- ($r_S = -0.316653$, $\text{p-value} = 2.25161\times10^{-18}$).}
+ ($r_S = -0.316653$, $\text{p-value} = 2.25161\times10^{-18}$).
  \label{fig:ab-rob}
@@
  \caption{\bf Expression dynamics of topology 9.}
- \hil{(A) Network topology number 9. Although this topology consists only of two
+ (A) Network topology number 9. Although this topology consists only of two
  genes, it display a striped pattern of gene expression for both of them as
@@
  with topology 9. Gene \textbf{B} is not shown as it was lost in this topology
- in the evolutionary process simulated by the genetic algorithm.}
+ in the evolutionary process simulated by the genetic algorithm.
  \label{fig:topol9}
@@
 interaction parameters and that more complex networks are less robusts. For
-example, topology number 420 \hil{(Fig}~\ref{fig:2neut-net}), which
+example, topology number 420 (Fig~\ref{fig:2neut-net}), which
 presents high complexity, was one of the

### 2021-05-14 docs/manuscript/manuscript.tex
Rename main file as manuscript.tex

--- 2021-05-14 docs/manuscript/PLOS-MSc-Thesis.tex
@@
 \setlength{\headheight}{27.023pt}
-\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
+% \lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
 \rfoot{\thepage/\pageref{LastPage}}
@@
 \fancyfootoffset[L]{2.25in}
-\lfoot{\sf PLOS}
+% \lfoot{\sf PLOS}
 
@@
 {\bf Proportion of relevant subgraphs in the set of GRNs.}
+The 19 subgraphs presented in this table are those presented in \nameref{S4_Fig}.
+Most of the topologies presented at least one negative feedback loop (70.73\%),
+and 62.46\% of them presented the I3-FFL network motif. The robustness of each
+subgraph was calculated as the mean robustness of the topologies presenting that
+subgraph.
 

### 2021-05-23 docs/manuscript/manuscript.tex
Include Munteanu et al. 2014 in introduction

@@
 
+Producing a stripe of gene expression is a fundamental patterning function
+in development, and it has been shown that feed-forward motifs are minimal
+networks capable of performing this function \cite{munteanu_2014}. For example,
 Cotterell and Sharpe studied what kind of 3-node network topologies could
@@
 implemented at least 7 different mechanisms and had different levels of
-complexity \cite{Cotterell2010}.
+complexity, being most of them variations of feed-forward motifs \cite{Cotterell2010}.
 

### 2021-05-30 modules/DesignMorphogeneResponsiveGRCs.m
Changes
DesignMorphogeneResponsiveGRCs.m:

- Add list of public functions.
- Indent code inside functions to improve readability.
- Remove IntTime, sigmoidThreshold, alpha and NumNuclei from
  Module's protected variables, these are global vars now.
- Add End[] and EndPackage[].

EvolAlgorithm4PatternFormingGRCModel2MathPackageV9.m:

- Import global parameters with Needs["parameters`"].

VideosExpressionDynamics.m, parameters.m and fitness2.m:

- Add BeginPackage, End and EndPackage.
- Add Begin[ "`Private`"].
- Import modules with Needs.

@@ -1,12 +1,35 @@
-(* ::Package:: *)
+BeginPackage["DesignMorphogeneResponsiveGRCs`"]
 
+PatternFilterChecking::usage="";
+AssessExpPattern4SingleStripeFeat::usage="";
+FitnessScore4StripePattern::usage="";
+AssessFitnessScore4StripePattern4SSMorpGradient4SFGRM::usage="";
+AssessExpDistance::usage="";
+AssessTargetGeneExpVarianceAcrossLastTimeWindow::usage="";
+StripeFormingGRCs4SSMorpGradientSFGRM::usage="";
+AdditiveRegContributionF::usage="";
+AssessFS4GRC2GenerateStripedPatternInducedBySSMorpGradient4SFGRM::usage="";
+Sol4StripeFormingGRCs4SSMorpGradientSFGRM::usage="";
+AssessDegenerateStruc4EngineeredMorpResponsiveGRC::usage="";
+AssessFS4GRC2GenerateStripedPatternInducedBySSMorpGradient4SFGRM2::usage="";
+AssessFitnessScore4StripePattern4SSMorpGradient4SFGRM2::usage="";
+FitnessScore4StripePattern2::usage="";
+EngineerMorphInducibleGRCGenotypes4SFGRM::usage="";
+SolveGRCDynsBeforeAfterMorphInput::usage="";
+ReturnTimeSeriesStripeFormingGRCs4SSMorpGradientSFGRM::usage="";
+SumAndFilterF::usage="";
+AdditiveRegContributionF::usage="";
+InspectGRCExpressionDynamics4SSMorpGradientSFGRM::usage="";
+
+Begin[ "`Private`"]
 Needs["parameters`"]
 
 (*USE THE FOLLOWING METHODS TO PROBE THE DESIGN SPACE OF 3-NODE MORPHOGENE RESPONSIVE GRCs*)
 (************************************************************************************************************************)
 (************************************************************************************************************************)
 
-PatternFilterChecking[ExpProfiles_List]:=(1/Length[ExpProfiles])(EuclideanDistance[ExpProfiles,ConstantArray[Mean[ExpProfiles],Length[ExpProfiles]]])
+PatternFilterChecking[ExpProfiles_List]:=
[…]

### 2021-05-30 modules/EvolAlgorithm4PatternFormingGRCModel2MathPackageV9.m

@@ -1,6 +1,6 @@
 (* ::Package:: *)
 
-BeginPackage["EvolAlgorithm4PatternFormingGRCModel2MathPackageV9`",{"ErrorBarPlots`", "parameters`"}]
+BeginPackage["EvolAlgorithm4PatternFormingGRCModel2MathPackageV9`",{"ErrorBarPlots`"}]
 
 (*
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@ -72,6 +72,8 @@ calculateMorphDecay::usage="calculateMorphDecay[NumNuclei_:30, A0_:1] calculates
 (**************************************************************************************************************)
 
 Begin["`Private`"]
+
+Needs["parameters`"]
 
 (**************************************************************************************************************)
 

### 2021-05-30 modules/VideosExpressionDynamics.m

@@ -1,12 +1,17 @@
-(* ::Package:: *)
-
-Needs["parameters`"]
+BeginPackage["VideosExpressionDynamics`"]
 
 (*Package to analyze the expression dynamics of each genotype.
 This package depends on EvolAlgorithm4PatternFormingGRCModel2MathPackageV9*)
-<<EvolAlgorithm4PatternFormingGRCModel2MathPackageV9`
 
-videoExp::usage = "videoExp[ genotipo ] create a video of the expression dynamics of a GRN, it takes as argument 'gen', with options 'a','b' and 'c', image size and 'fps' (frames per second)."
+expEspacioTemporal::usage = "Create expression profile.";
+videoExp::usage = "videoExp[ genotipo ] create a video of the expression \
+dynamics of a GRN, it takes as argument 'gen', with options 'a','b' and 'c', \
+image size and 'fps' (frames per second).";
+
+Begin[ "`Private`"]
+Needs["parameters`"]
+Needs["EvolAlgorithm4PatternFormingGRCModel2MathPackageV9`"]
+Needs["DesignMorphogeneResponsiveGRCs`"]
 
 (*Import list of classified, ordered, non isomorphic motifs*)
 motivosClasificadosOrdenadosNoIsomorfos = Flatten[ToExpression[#],1] &/@
@@ -119,3 +124,6 @@ If[
 	FitnessF4SingleStripe[GRCPhenotReadout]
 	]
 		]
+
+End[]
+EndPackage[]

### 2021-05-30 modules/fitness2.m

@@ -1,12 +1,11 @@
 (* ::Package:: *)
 (* Author: Juan Camilo Arboleda Rivera based on code by Jayson Gutiérrez*)
+BeginPackage["fitness2`"]
 
+Begin[ "`Private`"]
 Needs["parameters`"]
-
-(*This module depends on DesignMorphogeneResponsiveGRCs.m and
-EvolAlgorithm4PatternFormingGRCModel2MathPackageV9.m*)
-<< EvolAlgorithm4PatternFormingGRCModel2MathPackageV9`
-<< DesignMorphogeneResponsiveGRCs`
+Needs["EvolAlgorithm4PatternFormingGRCModel2MathPackageV9`"]
+Needs["DesignMorphogeneResponsiveGRCs`"]
 
 (*The following is basically the function
 AssessFS4GRC2GenerateStripedPatternInducedBySSMorpGradient4SFGRM2*)
@@ -88,3 +87,6 @@ modifiedAssessFitnessScore4StripePattern4SSMorpGradient4SFGRM2[ICs_List,
 
           FS = MapIndexed[(AssessExpPattern4SingleStripeFeat[#]*
                   Q1[PF[[#2[[1]]]]]*SSCondition) &, EndPointExpPatterns]]
+
+End[]
+EndPackage[]

### 2021-05-30 modules/parameters.m

@@ -1,3 +1,14 @@
+BeginPackage["parameters`"]
+
+NumNuclei::usage="";
+alpha::usage="";
+sigmoidThreshold::usage="";
+IntTime::usage="";
+OptimalPattern::usage="";
+Dmax::usage="";
+
+Begin[ "`Private`"]
+
 (*Global parameters to be used in the algorithm*)
 NumNuclei = 30;
 alpha = 5.;
@@ -10,7 +21,10 @@ maximal level observed along the 1D field of cells for the output genes*)
 OptimalPattern =
  Flatten[{ConstantArray[1, Quotient[NumNuclei, 3]],
    ConstantArray[10, Quotient[NumNuclei, 3] + Mod[NumNuclei, 3]],
-   ConstantArray[1, Quotient[NumNuclei, 3]]}]
+   ConstantArray[1, Quotient[NumNuclei, 3]]}];
 
 (*Maximal discrepancy achievable for a given pattern w.r.t to the optimal pattern above*)
 Dmax=9*30;
+
+End[]
+EndPackage[]

### 2021-05-31 docs/manuscript/manuscript.tex
Mention most studies focus on single input

@@
 \cite{Schaerli2014,Jaeger2004}, in synthetic networks
-\cite{Basu2005,Reinitz1995,Schaerli2018}, or have constrained the design of the
-GRNs \cite{Elowitz2000}.
+\cite{Basu2005,Reinitz1995,Schaerli2018,Elowitz2000}, or have constrained the
+design of the GRNs to accept an input signal from only one of its nodes, which is
+not always biologically realistic as multiple input systems can exist~
+\cite{Cotterell2010, munteanu_2014, Schaerli2014}.
 

### 2021-05-31 docs/manuscript/manuscript.tex
Avoid line breaking of citations

--- 2021-05-14 docs/manuscript/manuscript.tex
@@
 as those in which a group of cells differentiate in response to a signalling
-molecule. Such molecules were referred to as morphogens by Turing in 1952
-\cite{Turing1952} and posterior theoretical studies on patterning led to the
+molecule. Such molecules were referred to as morphogens by Turing in
+1952~\cite{Turing1952} and posterior theoretical studies on patterning led to the
 conceptualization of the French Flag Problem by Wolpert, who also stated
 in this respect that a gradient of concentration of a morphogen could
-trigger cell differentiation in a one-dimensional field of cells
-\cite{Wolpert1969, sharpe_2019}.
+trigger cell differentiation in a one-dimensional field of cells~\cite{Wolpert1969,
+sharpe_2019}.
 Although the Bicoid protein (bcd), the first example of a molecule that acted
 as a morphogen, was only found in the 80's in the developing embryos of
-\textit{Drosophila melanogaster}
-\cite{driever_gradient_1988, driever_bicoid_1988}, now it is known that there
+\textit{Drosophila melanogaster}~\cite{driever_gradient_1988,
+driever_bicoid_1988}, now it is known that there
 are many other examples such as the Decapentaplegic (Dpp) protein in
-\textit{Drosophila} wing imaginal discs \cite{affolter_decapentaplegic_2007},
-as well as Sonic Hedgehog \cite{dessaud_dynamic_2010,cohen_theoretical_2014}
-and Wnt \cite{raspopovic_digit_2014}. However, information processing by GRNs
+\textit{Drosophila} wing imaginal discs~\cite{affolter_decapentaplegic_2007},
+as well as Sonic Hedgehog~\cite{dessaud_dynamic_2010,cohen_theoretical_2014}
+and Wnt~\cite{raspopovic_digit_2014}. However, information processing by GRNs
 has proven to be a complex process, and a major goal of developmental biology
@@
 Moreover, the design and implementation of GRNs through synthetic biology is
-emerging as a promising tool to study biological phenomena as pattern formation
-\cite{santos-moreno_using_2019} and as a novel therapeutic tool with amazing
-biomedical applications~\cite{karlsson_therapeutic_2012,
+emerging as a promising tool to study biological phenomena as pattern
+formation~\cite{santos-moreno_using_2019} and as a novel therapeutic tool with
+interesting biomedical applications~\cite{karlsson_therapeutic_2012,
 higashikuni_advancing_2017,abil_synthetic_2015,healy_genetic_2019,
@@
 
+Producing a stripe of gene expression is a fundamental patterning function
+in development, and it has been shown that feed-forward motifs are minimal
+networks capable of performing this function~\cite{munteanu_2014}. For example,
 Cotterell and Sharpe studied what kind of 3-node network topologies could
@@
 implemented at least 7 different mechanisms and had different levels of
-complexity \cite{Cotterell2010}.
+complexity, being most of them variations of feed-forward motifs~\cite{Cotterell2010}.
 
 However, although several studies have analyzed the properties of GRNs that
-process positional information, they have focused on particular natural networks
-\cite{Schaerli2014,Jaeger2004}, in synthetic networks
-\cite{Basu2005,Reinitz1995,Schaerli2018}, or have constrained the design of the
-GRNs \cite{Elowitz2000}.
+process positional information, they have focused on particular natural
+networks~\cite{Schaerli2014,Jaeger2004}, in synthetic networks~\cite{Basu2005,
+Reinitz1995,Schaerli2018,Elowitz2000}, or have constrained the
+design of the GRNs to accept an input signal from only one of its nodes, which is
+not always biologically realistic as multiple input systems can exist~\cite{Cotterell2010,
+munteanu_2014, Schaerli2014}.
 
@@
 
-Our model is a modification of the model used by Cotterell \& Sharpe
-\cite{Cotterell2010} and proposed by Reinitz et al. \cite{Reinitz1995}. The
+Our model is a modification of the model used by Cotterell \& Sharpe~\cite{Cotterell2010}
+and proposed by Reinitz et al.~\cite{Reinitz1995}. The
 model is a dynamical system that describes the change in concentration of gene
@@
 
-We adopted the same definition of neighborhood as Cotterell \& Sharpe
-\cite{Cotterell2010}, in which two topologies were neighbors if the gain or
+We adopted the same definition of neighborhood as Cotterell \&
+Sharpe~\cite{Cotterell2010}, in which two topologies were neighbors if the gain or
 removal of any one interaction can transform one of the topologies into the
@@
 As a measure of network complexity, we used the complexity index based on
-Shannon entropy proposed by Bonchev \& Rovray \cite{D.2005}. We calculated this
+Shannon entropy proposed by Bonchev \& Rovray~\cite{D.2005}. We calculated this
 index using the following equation:
@@
 (Pearson’s $\chi^2 = 12.8902$, $\text{p-value} = 0.115684$) characteristic of a
-Erdös-Rényi network \cite{Erdos1959}. This kind of network presents a great
+Erdös-Rényi network~\cite{Erdos1959}. This kind of network presents a great
 number of nodes with low degree and a few nodes with high degree
@@
 the most robust subgraphs is the ``Bistable'' motif reported by Cotterell \&
-Sharpe \cite{Cotterell2010}, the I3-FFL and the ``Overlapping domains'' motif
+Sharpe~\cite{Cotterell2010}, the I3-FFL and the ``Overlapping domains'' motif
 (which is a variation of the I3-FFL) also reported by Cotterell \& Sharpe (see
@@
  Average vertex degree  & 3.85714 \\ \hline
- Clustering coefficient \cite{Watts1998} & 0.117097 \\ \hline
+ Clustering coefficient~\cite{Watts1998} & 0.117097 \\ \hline
  Average path length    & 7.25261       \\ \hline
@@
 
-Robustness is an important feature of patterning mediated by morphogens
-\cite{lo_robust_2015}. In our study, the most abundant topologies were found to
-be part of highly connected clusters in the neutral network, and this fact
+Robustness is an important feature of patterning mediated by
+morphogens~\cite{lo_robust_2015}. In our study, the most abundant topologies were
+found to be part of highly connected clusters in the neutral network, and this fact
 indicates that these topologies are both robust to quantitative changes in the
@@
 qualitative changes in the topology itself. As robustness and evolvability are
-not mutually exclusive properties \cite{wagner_robustness_2008}, it is likely
+not mutually exclusive properties~\cite{wagner_robustness_2008}, it is likely
 that these topologies are also evolvable.
@@
 Other studies have shown that Incoherent Feed Forward Loops are able to form
-spatial stripes of gene expression \cite{ishihara_cross_2005}. In particular,
+spatial stripes of gene expression~\cite{ishihara_cross_2005}. In particular,
 we found that the I3-FFL is one of the network motifs that appears more
-frequently in these GRNs, as has also been reported by Cotterell \& Sharpe
-\cite{Cotterell2010}, and has been found in transcriptional networks involved
-in cell fate definition \cite{Li2019}. Moreover, our results indicate that this
+frequently in these GRNs, as has also been reported by Cotterell \&
+Sharpe~\cite{Cotterell2010}, and has been found in transcriptional networks involved
+in cell fate definition~\cite{Li2019}. Moreover, our results indicate that this
 is a robust motif that can be useful for construction of synthetic gene
@@
 Forward Loop type 2 (I2-FFL) is able to generate a striped gene expression
-pattern, \cite{Schaerli2014, Basu2005}, in our study this motif was present in
+pattern,~\cite{Schaerli2014, Basu2005}, in our study this motif was present in
 only 10.92\% of the topologies. This fact may be because genetic algorithms
@@
 We also found topologies with a similar topology to that of the “opposing
-gradients” reported by Schaerli et al. \cite{Schaerli2018,Schaerli2014}.
+gradients” reported by Schaerli et al.~\cite{Schaerli2018,Schaerli2014}.
 Although the “opposing gradients” mechanism requires the constitutive expression
@@
 Although GRNs that produce a band of gene expression with only two nodes have
-been described (e.g. the I-zero motif, \cite{Schaerli2014}) this is to our
+been described (e.g. the I-zero motif,~\cite{Schaerli2014}) this is to our
 knowledge the first time a GRN with a topology such as that of topology nine
@@
 These are subgraphs that have been reported in previous studies as networks
-involved in morphogenesis and development \cite{Cotterell2010, Schaerli2014,
+involved in morphogenesis and development~\cite{Cotterell2010, Schaerli2014,
 Schaerli2018}. These were used to calculate the subgraph profile and the results

### 2021-05-31 docs/manuscript/manuscript.tex
Emphasize role of multiple inputs

@@
 development. In this article we use an \textit{in silico} approach in order to find and study
-the design space of three node Gene Regulatory Networks that are able to produce a
-striped pattern of gene expression.
+the design space of multiple input three node Gene Regulatory Networks that are able
+to produce a striped pattern of gene expression.
 
@@
 
-In this study, we used a set of 2061 three-node GRNs selected by a genetic
-algorithm to interpret a morphogen signal and respond to it generating a
-striped pattern of gene expression in order to analyse the role of network
-motifs in this process and its dynamic properties.
+In this study, we used a genetic algorithm to generate a set of 2061 three-node GRNs
+including single and multiple input topologies that were able to interpret a morphogen
+signal and respond to it generating a striped pattern of gene expression in order to
+analyse the role of network motifs in this process and its dynamic properties.
 

### 2021-05-31 docs/manuscript/manuscript.tex
Explicitly mention S4_Fig and S1_Table
In response to Reviewer #1, Minor Points, 2, e.

@@
 Sharpe~\cite{Cotterell2010}, the I3-FFL and the ``Overlapping domains'' motif
-(which is a variation of the I3-FFL) also reported by Cotterell \& Sharpe (see
-Supplementary Material).
+(which is a variation of the I3-FFL) also reported by Cotterell \& Sharpe
+(see \nameref{S4_Fig} and \nameref{S1_Table}).
 

### 2021-06-01 modules/parameters.m
Change 30 for NumNuclei in Dmax

@@ -24,7 +24,7 @@ OptimalPattern =
    ConstantArray[1, Quotient[NumNuclei, 3]]}];
 
 (*Maximal discrepancy achievable for a given pattern w.r.t to the optimal pattern above*)
-Dmax=9*30;
+Dmax=9*NumNuclei;
 
 End[]
 EndPackage[]

### 2021-06-13 docs/manuscript/manuscript.tex
Refer to S2_Fig in methods
In response to Reviewer #1, Minor Points, 2, g.
Improve discussion about I4-FFL and constitutive expression

@@
 system. We chose that number of time steps because numerical experiments showed
-that GRNs reached the steady state in approximately 300 steps.
+that GRNs reached the steady state in approximately 300 steps (\nameref{S2_Fig}).
 
@@
 
-However, as our model was less stringent, we found a larger number of possible
-topologies with the same striped phenotype.\\
+However, as our model was less stringent and allowed the morphogen input to
+interact with any gene, we found a larger number of possible topologies with the
+same striped phenotype. For example, in a single input model the I4-FFL could
+not produce a stripe of gene expression~\cite{munteanu_2014}, whereas in our
+experimental setup 20.17\% of the topologies presented this motif as a
+subgraph~(\nameref{S1_Table}).
 
@@
 gradients” topology, had also commonly positive auto-regulation interactions on
-these two nodes, supplying in this way the lack of constitutive expression.
+these two nodes, supplying in this way the lack of constitutive
+expression as found in previous studies~\cite{munteanu_2014}.
 

### 2021-06-17 modules/DesignMorphogeneResponsiveGRCs.m
Add (* ::Package:: *)

@@ -1,3 +1,4 @@
+(* ::Package:: *)
 BeginPackage["DesignMorphogeneResponsiveGRCs`"]
 
 PatternFilterChecking::usage="";

### 2021-06-17 modules/VideosExpressionDynamics.m

@@ -1,3 +1,4 @@
+(* ::Package:: *)
 BeginPackage["VideosExpressionDynamics`"]
 
 (*Package to analyze the expression dynamics of each genotype.

### 2021-06-17 modules/parameters.m

@@ -1,3 +1,4 @@
+(* ::Package:: *)
 BeginPackage["parameters`"]
 
 NumNuclei::usage="";

### 2021-06-17 modules/DesignMorphogeneResponsiveGRCs.m
Import Evol... module in Design... module

@@ -24,6 +24,7 @@ InspectGRCExpressionDynamics4SSMorpGradientSFGRM::usage="";
 
 Begin[ "`Private`"]
 Needs["parameters`"]
+Needs["EvolAlgorithm4PatternFormingGRCModel2MathPackageV9`"]
 
 (*USE THE FOLLOWING METHODS TO PROBE THE DESIGN SPACE OF 3-NODE MORPHOGENE RESPONSIVE GRCs*)
 (************************************************************************************************************************)

### 2021-06-17 modules/DesignMorphogeneResponsiveGRCs.m
Pass EvolSteps as argument and change Block for Module

@@ -310,12 +310,12 @@ interactions can be effective in controlling the expression putput.*)
 (************************************************************************************************************************)
 
 (*Engineer morphogene responsive GRCs using a MCMC-like algorithm*)
-EngineerMorphInducibleGRCGenotypes4SFGRM[Repl_,PSR_,ThrFS_]:=Block[{EvolSteps,thresholdFitness,ParamSettingSamplingRate,
+EngineerMorphInducibleGRCGenotypes4SFGRM[Repl_,EvSteps_,PSR_,ThrFS_]:=Module[{EvolSteps,thresholdFitness,ParamSettingSamplingRate,
 MutaProbList,GRCWiring,MorphogenInput,DiffRates,ProtDegRates,GetInputParamSet,GRCDyns,RefInputParamSetting,Fb,
 Fa,DeltaF,ThisMutId,InputParamSetting,InputParamSettingAllin,OutputFileName,PatternFormingSeeds,DestinationFolder,
 ParamsSamplingCond,SamplingPointsList,BoolWMatrix,WContainer},
 
-EvolSteps=5000000;
+EvolSteps=EvSteps;(*5000000; *)
 ParamSettingSamplingRate=PSR;(*e.g. PSR = 500*)
 thresholdFitness=ThrFS;(*e.g. ThrFS = 0.925*)
 SamplingPointsList=Drop[Range[0,EvolSteps,ParamSettingSamplingRate],{1}];

### 2021-06-17 modules/EvolAlgorithm4PatternFormingGRCModel2MathPackageV9.m
Import Design... module into Evol... module

@@ -72,8 +72,8 @@ calculateMorphDecay::usage="calculateMorphDecay[NumNuclei_:30, A0_:1] calculates
 (**************************************************************************************************************)
 
 Begin["`Private`"]
-
 Needs["parameters`"]
+Needs["DesignMorphogeneResponsiveGRCs`"]
 
 (**************************************************************************************************************)
 

### 2021-06-17 modules/EvolAlgorithm4PatternFormingGRCModel2MathPackageV9.m
Change Block for Module in calculateMorphDecay

@@ -154,7 +154,7 @@ Method->{"Automatic","Rotation"->2 Pi},VertexRenderingFunction->({EdgeForm[Color
 
 (******************************************************************************)
 
-calculateMorphDecay[NumNuclei_:30, A0_:1] := Block[
+calculateMorphDecay[NumNuclei_:30, A0_:1] := Module[
     {rightEndConcentration, h},
 
   rightEndConcentration = A0*Exp[-29/(30*0.4)];

### 2021-06-17 example/DesignMorphogeneResponsiveGRCs.m
Delete duplicated files: DRY

@@ -1,470 +1,0 @@
-(* ::Package:: *)
-
-(*USE THE FOLLOWING METHODS TO PROBE THE DESIGN SPACE OF 3-NODE MORPHOGENE RESPONSIVE GRCs*)
-(************************************************************************************************************************)
-(************************************************************************************************************************)
-
-(************************************************************************************************************************)
-(************************************************************************************************************************)
-
-PatternFilterChecking[ExpProfiles_List]:=(1/Length[ExpProfiles])(EuclideanDistance[ExpProfiles,ConstantArray[Mean[ExpProfiles],Length[ExpProfiles]]])
-
-Q1[x_]:=(x^5)/( (x^5) + (0.1^5) )
-
-(************************************************************************************************************************)
-
-AssessExpPattern4SingleStripeFeat[ExpProfiles_List]:=Block[{},
-
-(*Optimal expression pattern represented in a scale between 1-10 in expression level, being 10 an expression level which is > 90% of the maximal level observed along the 1D field of cells for the output genes*)
-OptimalPattern=Flatten[{ConstantArray[1,10],ConstantArray[10,10],ConstantArray[1,10]}];
-(*Maximal discrepancy achievable for a given pattern w.r.t to the optimal pattern above*)
-Dmax=9*30;
-
-ThresholdedExpValues={{0.`,0.1`},{0.1`,0.2`},{0.2`,0.3`},{0.3`,0.4`},{0.4`,0.5`},{0.5`,0.6`},{0.6`,0.7`},{0.7`,0.8`},{0.8`,0.9`},{0.9`,1.001`}};
-
-testInt[Rng_,Value_]:= Rng[[1]]<=Value< Rng[[2]];
-
-NormEquilbPattern=ExpProfiles/Max[ExpProfiles];
-
-DiscretizedPattern=Table[Position[testInt[#,NormEquilbPattern[[x]]]&/@ThresholdedExpValues,True][[1,1]],{x,Length[NormEquilbPattern]}];
-
[…]

### 2021-06-17 example/EvolAlgorithm4PatternFormingGRCModel2MathPackageV9.m

@@ -1,790 +1,0 @@
-(* ::Package:: *)
-
-BeginPackage["EvolAlgorithm4PatternFormingGRCModel2MathPackageV9`",{"ErrorBarPlots`"}]
-
-(*
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Package developed by Jayson Guti\[EAcute]rrez on 15/9/2015%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-*)
-
-
-SSInputMorphogen::usage = "Generate input morphogen profile";
-
-GetRndComposedWiringMatrix::usage = "generate wiring";
-
-GetRndComposedWiringMatrix2::usage = "generate wiring";
-
-GetProteinDegrad::usage = "generate protein degradation";
-
-GetDiffRates::usage = "generate diffusion";
-
-DisplayNonWeightedHaploidGraphWithMorphInputs::usage = "Function for displaying non-weigthed regulatory wirings (GRC design templates), including the input Morphogen";
-
-DisplayWeightedHaploidGraphWithMorphInputs::usage = "Function for displaying weigthed regulatory wirings, including the input Morphogen";
-
-SpaceTimeFeats4StripeFormingGRCs4SSMorpGradientWithBiophyGroundedCRIF0::usage = "Function to assess spatio-temporal expression portraits of the GRCs implementing the thermodynamically-grounded CRIF";
-
-SpaceTimeFeats4StripeFormingGRCs4SSMorpGradientWithSumAndFilterModel0::usage = "Function to assess spatio-temporal expression portraits of the GRCs implementing the Sum&Filter CRIF";
-
-CreatSpaceTimePlots4GRC::usage="Function to create the spatio-temporal expression portraits of the GRCs";
-
[…]

### 2021-06-18 docs/manuscript/manuscript.tex
State that the interval for interactions was chosen arbitrarily
Add citation to Exelby et al. 2021

@@
 represents a regulator node of the GRN (Fig \ref{fig:model}A and
-\ref{fig:model}B) with $w_{\textit{ij}}~\in [-10,\ 10]$. Values in these
+\ref{fig:model}B) with $w_{\textit{ij}}~\in [-10,\ 10]$ (this interval was arbitrarily
+chosen and it is inherited from the model of Munteanu~et~al.~\cite{munteanu_2014}
+and Cotterell~\&~Sharpe\cite{Cotterell2010}). Values in these
 matrices can be negative, zero or positive and represent represion, no
@@
 
-Robustness is an important feature of patterning mediated by
-morphogens~\cite{lo_robust_2015}. In our study, the most abundant topologies were
+Robustness is an important feature of patterning mediated by morphogens, and
+the GRNs involved can be selected in an evolutionary process not only by its final
+steady state but also by its ability to buffer noise~\cite{lo_robust_2015, exelby_2021}.
+In our study, the most abundant topologies were
 found to be part of highly connected clusters in the neutral network, and this fact

### 2021-06-19 docs/manuscript/manuscript.tex
Add example of two-input network
Add range for diffusion parameters

@@
 not always biologically realistic as multiple input systems can exist~\cite{Cotterell2010,
-munteanu_2014, Schaerli2014}.
+munteanu_2014, Schaerli2014}. For example, the neural subtypes specification
+system in the vertebrate neural tube is a two-input network in which
+Sonic Hedgehog acts as a morphogen and Olig2 and Nkx2.2 act as the receiver
+nodes~\cite{dessaud_dynamic_2010, balaskas2012, exelby_2021}.
 
@@
 the three gene products; and the third of the degradation rates ($\delta$) of
-these same gene products, with $D$ and $\delta  \in [0, 0.05]$.\\
+these same gene products, with $D \in [0, 0.1]$ and $\delta \in [0, 0.05]$.\\
 

### 2021-06-20 docs/manuscript/manuscript.tex
Fix degree dist section

@@
 (Pearson’s $\chi^2 = 12.8902$, $\text{p-value} = 0.115684$) characteristic of a
-Erdös-Rényi network~\cite{Erdos1959}. This kind of network presents a great
+Erdös-Rényi network with many nodes and edges~\cite{Erdos1959}. This kind of
+network presents a great
 number of nodes with low degree and a few nodes with high degree

### 2021-06-21 modules/GraphUtils.m
refactor: Create module GraphUtils.m
- Functions generarIsomorfos and isSubgraphQ were moved from
VideosExpressionDynamics.m to GraphUtils.m

@@ -1,0 +1,83 @@
+(* ::Package:: *)
+
+BeginPackage["GraphUtils`"]
+
+getIsomorphs::usage =
+    "Generates the different isomorphs of a graph as adjacency matrices."
+
+isSubgraphQ::usage =
+    "Gives True if graph1 is "<>
+    "subgraph of graph2 and False in other case. Both lists must be "<>
+    "adjacency matrices."
+
+isGraphQ::usage =
+    "isGraphQ[ graph1_List, graph2_List ] Gives True if graph1 is the "<>
+    "same graph as graph2 (or an isomorph) and False in other case. "<>
+    "Both lists must be adjacency matrices."
+
+getSupergraphs::usage =
+    "getSupergraphs[topology_List, topologies_List] Returns a list of the "<>
+    "positions of topologies that are supergraphs of topology."
+
+findGraph::usage =
+    "findGraph[topology_List, topologies_List] Returns a list of the "<>
+    "positions of topologies that are isomorphic with topology."
+
+Begin["Private`"]
+
+getIsomorphs[genotipo_List] :=
+ Module[{rowOrder, columnOrder, isomorphs},
+  (*Define matrices indicating the order of rows and columns in the
+    matrices*)
+  rowOrder = Permutations[{1, 2, 3}];
+
+  (*columnOrder must be equal to rowOrder but with the morphogen
+    always as the first column*)
+  columnOrder =
+   Table[Prepend[rowOrder[[i]] + 1, 1], {i, Length[rowOrder]}];
+
+  (*Generate isomorphs for the matrix*)
+  isomorphs =
+   Table[Sign[genotipo][[rowOrder[[i]], columnOrder[[i]]]], {i,
+     Length[rowOrder]}]
+  ]
+
+isSubgraphQ[graph1_List, graph2_List] :=
+ Module[{isomorphs},
+  isomorphs = getIsomorphs[graph1];
+  Do[If[i > Length[isomorphs], Break[False]];
+   If[SubsetQ[Position[Sign[graph2], 1], Position[isomorphs[[i]], 1]] &&
+      SubsetQ[Position[Sign[graph2], -1],
+      Position[isomorphs[[i]], -1]], Break[True]];, {i,
+    Length[isomorphs] + 1}]
+  ]
+
+isGraphQ[graph1_List, graph2_List] :=
+ Module[{isomorphs},
[…]

### 2021-06-21 modules/VideosExpressionDynamics.m

@@ -82,29 +82,6 @@ videoExp[ genotipo_List, OptionsPattern[]] := Module[{patron},
         ImageSize -> OptionValue[tama\[NTilde]o] ], {n, 30}], OptionValue[fps] ]
     , {k, 3} ] ] ]
 
-generarIsomorfos::usage = "generarIsomorfos[ genotype(matrix) ] This function generates the different isomorphs of a matrix as adjacency matrices"
-
-generarIsomorfos[genotipo_List]:= Module[{ordfilas,ordcolumnas,isomorfos},
-  (*Define matrices indicating the order of rows and columns in the matrices*)
-  ordfilas = Permutations[{1, 2, 3}];
-  (*ordcolumnas must be equal to ordfilas but with the morphogen always as the first column*)
-  ordcolumnas = Table[ Prepend[ordfilas[[i]] + 1, 1] , {i, Length[ordfilas]}];
-  (*Generate isomorphs for the matrix*)
-  isomorfos = Table[
-    Sign[genotipo][[ ordfilas[[i]], ordcolumnas[[i]] ]]
-    , {i, Length[ordfilas]}] ]
-
-isSubgraphQ::usage = "isSubgraphQ[ list_1, list_2 ] Gives True if list_1 is subgraph of list_2 and False in other case. Both lists must be adjacency matrices."
-
-isSubgraphQ[subgraph_List,graph_List]:= Module[{isomorfos},
-  isomorfos = generarIsomorfos[subgraph];
-  Do[ If[ i > Length[isomorfos], Break[False] ];
-      If[
-    SubsetQ[ Position[ Sign[graph],1], Position[isomorfos[[i]],1] ] &&
-    SubsetQ[ Position[ Sign[graph],-1], Position[isomorfos[[i]],-1] ], Break[True] ];
-    ,{i, Length[isomorfos]+1}
-    ]]
-
 (********************************************************************************************************************)
 
 (*Redefine function "AssessFitnessScoreStripeFormingGRCs4SSMorpGradient"

### 2021-06-21 modules/hammingGraph.m
rename: ordfilas -> rowOrder, ordcolumnas -> columnOrder

@@ -9,10 +9,10 @@ importClassifiedTopologies[]:= {
 
 isomorfos::usage = "isomorfos[matriz] genera los isomorfos de una matriz dada."
 (*Definimos la función que presenta los isomorfos de cada topología*)
-isomorfos[matriz_] := Module[{ordfilas, ordcolumnas},
-  ordfilas = Permutations[{1, 2, 3}];
-  ordcolumnas = Table[Prepend[ordfilas[[i]] + 1, 1], {i, Length[ordfilas]}];
-  Table[ matriz[[ordfilas[[i]], ordcolumnas[[i]] ]], {i, Length[ordfilas]}] ]
+isomorfos[matriz_] := Module[{rowOrder, columnOrder},
+  rowOrder = Permutations[{1, 2, 3}];
+  columnOrder = Table[Prepend[rowOrder[[i]] + 1, 1], {i, Length[rowOrder]}];
+  Table[ matriz[[rowOrder[[i]], columnOrder[[i]] ]], {i, Length[rowOrder]}] ]
 
 (*La siguiente es la misma función de antes que calcula la menor distancia de Hamming entre dos matrices, pero con la nueva parte del código*)
 jaminnn::usage = "jaminnn[list1, list2] compara los isomorfos de las dos listas y arroja la menor distancia de Hamming entre ellas."

### 2021-06-22 docs/manuscript/manuscript.tex
Fix value of initial concentration of gene products
Correction made by Reviewer #2
State that morphogen dynamics is not considered

@@
 concentration gradient (Fig~\ref{fig:model}C). The initial concentration of
-each gen product in all cells was set to 0.1 in all simulations.\\
+each gen product in all cells was set to 1 in all simulations.\\
 
@@
 total number of cells, and $h$ is a decay parameter, whose value in our model
-was set to 0.4.
+was set to 0.4. We do not consider here how the morphogen gradient can be
+formed, i.e., we do not consider morphogen dynamics but we assume a static
+gradient.
 

### 2021-06-23 modules/DesignMorphogeneResponsiveGRCs.m
Don't create SampledMorphInducibleGRCGenotypes folder

@@ -332,7 +332,7 @@ NumMutants:= RandomSample[MutaProbList-> Range[10],1][[1]];
 
 (***********************************************************************)
 (***********************************************************************)
-DestinationFolder = "GRCModelWithSumAndFilterCRIF/SampledMorphInducibleGRCGenotypes";
+DestinationFolder = "GRCModelWithSumAndFilterCRIF";
 (*To ensure that the destination folders exist, if don't then create them!*)
 
 If[

### 2021-06-23 modules/DesignMorphogeneResponsiveGRCs.m
Add NumNuclei to the output filename

@@ -341,7 +341,7 @@ If[
 	];
 
 (*To read in within mathematica these files do: ReadList[FitnessFeaturesFile]*)
-OutputFileName = DestinationFolder<>"/ParamGenotypeGRC_Repl_"<>ToString[Repl]<>".dat";
+OutputFileName = DestinationFolder<>"/ParamGenotypeGRC_Repl_"<>ToString[Repl]<>"_NumNuclei_"<>ToString[NumNuclei]<>".dat";
 (***********************************************************************)
 (***********************************************************************)
 

### 2021-06-23 example/runLocalMCMClikeEvolGRCMathScriptFile.sh
Add for loop to run algorithm with different values of NumNuclei

@@ -7,5 +7,11 @@ arg2=5000
 arg3=500
 arg4=0.95
 
+for NUM_NUCLEI in 10 20 40 50; do
+
+echo "Running algorithm with NumNuclei = $NUM_NUCLEI"
+
 #/Applications/Mathematica2.app/Contents/MacOS/MathKernel math -noprompt -run "Repl=$arg1;EvolSteps=$arg2;ParamSettingSamplingRate=$arg3;thresholdFitness=$arg4;<<runMCMClikeEvolGRCMathScriptFile.m"
-math -noprompt -run "Repl=$arg1;EvolSteps=$arg2;ParamSettingSamplingRate=$arg3;thresholdFitness=$arg4;<<runMCMClikeEvolGRCMathScriptFile.m"
+math -noprompt -run "uNumNuclei=$NUM_NUCLEI;Repl=$arg1;EvolSteps=$arg2;ParamSettingSamplingRate=$arg3;thresholdFitness=$arg4;<<runMCMClikeEvolGRCMathScriptFile.m"
+
+done

### 2021-06-23 example/runMCMClikeEvolGRCMathScriptFile.m

@@ -4,6 +4,9 @@ SeedRandom[RandomInteger[10^7, 1][[1]]]
 
 << "../modules/EvolAlgorithm4PatternFormingGRCModel2MathPackageV9.m"
 << "../modules/DesignMorphogeneResponsiveGRCs.m"
+<< "../modules/parameters.m"
+
+NumNuclei = uNumNuclei
 
 EngineerMorphInducibleGRCGenotypes4SFGRM[ToExpression[Repl],ToExpression[EvolSteps],ToExpression[ParamSettingSamplingRate],ToExpression[thresholdFitness]]
 

### 2021-06-29 modules/VideosExpressionDynamics.m
Add function to plot expression dynamics

@@ -4,10 +4,16 @@ BeginPackage["VideosExpressionDynamics`"]
 (*Package to analyze the expression dynamics of each genotype.
 This package depends on EvolAlgorithm*)
 
-expEspacioTemporal::usage = "Create expression profile.";
-videoExp::usage = "videoExp[ genotipo ] create a video of the expression \
+expEspacioTemporal::usage =
+    "Create expression profile. Returns a list of lists.";
+
+videoExp::usage =
+    "videoExp[ genotipo ] create a video of the expression \
 dynamics of a GRN, it takes as argument 'gen', with options 'a','b' and 'c', \
 image size and 'fps' (frames per second).";
+
+expressionDynamics::usage =
+    "Returns a plot of the expression dynamics of the given GRN."
 
 Begin[ "`Private`"]
 Needs["parameters`"]
@@ -102,6 +108,12 @@ If[
 	FitnessF4SingleStripe[GRCPhenotReadout]
 	]
 		]
+
+(******************************************************************************)
+
+expressionDynamics[grn_List]:=Module[{},
+  CreatSpaceTimePlots4GRC[expEspacioTemporal[grn]]
+]
 
 End[]
 EndPackage[]

### 2021-07-02 docs/manuscript/manuscript.tex
Aknowledge computational services
change: ditribution -> abundance distribution

@@
 We calculated the Shannon entropy Eq~(\ref{ec:shannon}) of the network
-topologies distribution in order to test if this distribution could be obtained
+topologies abundance distribution in order to test if this distribution could be obtained
 by chance with a probability greater than 0.05. In order to obtain a 95\%
@@
 
+\section*{Aknowledgements}
+The computational resources (Stevin Supercomputer Infrastructure) and
+  services used in this work were provided by the VSC (Flemish Supercomputer
+  Center), funded by Ghent University, FWO and the Flemish Government –
+  department EWI.
+
 \nolinenumbers

### 2021-07-03 docs/manuscript/manuscript.tex
Add field size experiments to methods

@@
 
+\subsection*{Change in morphogenetic field size}
+We wanted to evaluate if the GRNs evolved in our model with a morphogenetic
+field size of 30 cells could produce the desired pattern in morphogenetic fields
+with a different number of cells. For this reason, we calculated the fitness of
+each GRN in morphogenetic fields with 10, 20, 40, and 50 cells. In these
+simulations we redefined the morphogen gradient in such a way that the initial
+and final concentration were constant in all the fields (\nameref{S5_Fig}).
+In order to know if topologies with the same topology had a similar fitness
+in the different fields, we performed a Principal Component Analysis in
+Scikit-learn~0.24.2~\cite{sklearn} using the fitness values of each GRN in
+the four different fields.
+
 % For placing video, \nameref{S1_Video} vel sagittis arcu lobortis.
@@
 
+\paragraph*{S5 Fig.}
+\label{S5_Fig}
+{\bf Response of GRNs to changes in morphogenetic field size.}
+(A) Morphogen gradients of fields with 10, 20, 40 and 50 cells. (B) Fitness of the
+GRNs in each one of the morphogenetic fields. (C) Principal Component Analysis
+of fitness by morphogenetic field size. Although principal components separate
+GRNs in clusters, not all the GRNs with the same topology are located in the
+same group. (D) Average fitness of GRNs evolved in different morphogenetic
+field sizes.
+
 \paragraph*{S1 Table.}

### 2021-07-03 docs/manuscript/manuscript.tex
Add field size to results

@@
 
-The average vertex degree indicates that in general all the network topologies
+The average vertex degree indicates that in general, all the network topologies
 display high robustness and that they could resist an average of approximately 4
@@
 path length indicates that the evolutionary landscape is easily accessible for
-most network topologies (Table~\ref{table1}).
+most network topologies (Table~\ref{table1}). Although the GRNs in this modelling
+framework are robust to changes in its topology, they are sensitive to changes in
+the morphogenetic field size, especially when this size is reduced, as only
+4.66, 56.82, 56.14 and 41.05 percent of the GRNs had a fitness greater than 0.9
+in fields with 10, 20, 40 and 50 cells, respectively (\nameref{S5_Fig}B).
 

### 2021-07-03 docs/manuscript/manuscript.tex
Discuss PCA results

@@
 simulations we redefined the morphogen gradient in such a way that the initial
-and final concentration were constant in all the fields (\nameref{S5_Fig}).
+and final concentration were constant in all the fields (\nameref{S5_Fig}A).
 In order to know if topologies with the same topology had a similar fitness
@@
 Robustness is an important feature of patterning mediated by morphogens, and
-the GRNs involved can be selected in an evolutionary process not only by its final
-steady state but also by its ability to buffer noise~\cite{lo_robust_2015, exelby_2021}.
+the GRNs involved can be selected in an evolutionary process not only by their final
+steady state but also by their ability to buffer noise~\cite{lo_robust_2015, exelby_2021}.
 In our study, the most abundant topologies were
@@
 not mutually exclusive properties~\cite{wagner_robustness_2008}, it is likely
-that these topologies are also evolvable.
+that these topologies are also evolvable.\\
+
+In contrast to this robustness to
+qualitative and quantitative changes, we found that many GRNs were susceptible
+to changes in morphogenetic field size, but interestingly, GRNs with the same
+topology behave in different ways for each field size (\nameref{S5_Fig}C),
+indicating that the specific values of network parameters are more important
+than the network topology when it comes to producing the same pattern at
+different scales.
 

### 2021-07-03 docs/manuscript/manuscript.tex
Add info about diffusion and degradation parameters

@@
 interact with any gene, we found a larger number of possible topologies with the
-same striped phenotype. For example, in a single input model the I4-FFL could
+same striped phenotype and a diverse combination of diffusion and degradation
+parameters (\nameref{S6_Fig}).
+For example, in a single input model the I4-FFL could
 not produce a stripe of gene expression~\cite{munteanu_2014}, whereas in our
@@
 
+\paragraph*{S6 Fig.}
+\label{S6_Fig}
+{\bf Diffusion and degradation parameters.}
+(A) Ternary plot showing the combination of diffusion parameters for genes A,
+B and C. (B) Ternary plot showing the combination of degradation parameters
+for genes A, B and C.
+
 \paragraph*{S1 Table.}

### 2021-07-03 docs/manuscript/manuscript.tex
Add link to source code
Add expression dynamics

@@
 
+The source code for the genetic algorithm and other parts of analysis is available at
+\url{https://github.com/Nesper94/gene-regulatory-networks}.
+
 \subsection*{Classification of GRNs}
@@
 very abundant network topologies were grouped together with GRNs displaying
-very infrequent network topologies.
+very infrequent network topologies, indicating that additional interactions
+in the networks have often the function of fine tunning gene expression, instead
+of drastically changing the expression mechanism. This can be seen in the
+expression dynamics of networks with different topologies (\nameref{S7_Fig}).
 
@@
 
+\paragraph*{S7 Fig.}
+\label{S7_Fig}
+{\bf Expression dynamics of main networks.}
+
 \paragraph*{S1 Table.}

### 2021-07-03 docs/manuscript/manuscript.tex
Extend discussion on robustness and connectivity

@@
 strength of interactions and diffusion and degradation parameters as well as to
-qualitative changes in the topology itself. As robustness and evolvability are
+qualitative changes in the topology itself. This, in principle, means that many
+mechanisms for producing striped patterns are available and accessible to each
+other, i.e., an organism can change between different GRNs without losing its
+fitness, which allows to fine-tune the regulatory system according to its specific
+needs. As robustness and evolvability are
 not mutually exclusive properties~\cite{wagner_robustness_2008}, it is likely
-that these topologies are also evolvable.\\
+that these topologies are also evolvable in the sense that they can produce
+heritable variation of the expression pattern.\\
 

### 2021-07-03 docs/manuscript/manuscript.tex
Add closing statement
Improve closing statement
Include Gierer-Meinhardt

@@
 
-Among the new network topologies we report here, perhaps the most interesting
-is the topology number 9, which is able to express the striped gene expression
-pattern with only two genes and indeed expresses this pattern in both genes
-(Fig~\ref{fig:topol9}).
+Among the new network topologies we report here, perhaps one of the most
+interesting ones is the topology number 9, which is able to express the striped
+gene expression pattern with only two genes (Fig~\ref{fig:topol9}).
+This topology is composed of a core that consists in the
+Gierer-Meinhardt network~\cite{giererMeinh1972, meinhardt2000} known to
+generate stripes and dots in bidimensional fields, and additional
+interactions such as the self inhibition of gene A and the inhibition of the genes
+by the morphogen. The Gierer-Meinhardt network consists in an activator of low
+diffusivity (gene C) and an inhibitor with greater diffusivity (gene A). The additional
+interactions of topology number 9 have the role of restricting the generation of
+the expression band to a precise location.
 
@@
 
+This work provides a catalog of GRNs that can be extended in future computational
+and theoretical analysis and provides a foundation to design and build synthetic
+regulatory networks or to identify new ones existing in nature.
+
 %\section*{Conclusion}

### 2021-07-03 docs/manuscript/manuscript.tex
Add example to discussion

@@
 mechanisms for producing striped patterns are available and accessible to each
-other, i.e., an organism can change between different GRNs without losing its
+other, i.e., an organism can change between different GRNs (for example between
+a Gierer-Meinhardt-like network to an Incoherent Feed-Forward Loop) without losing its
 fitness, which allows to fine-tune the regulatory system according to its specific

### 2021-07-03 docs/manuscript/manuscript.tex
Fix deg range description

@@
 the three gene products; and the third of the degradation rates ($\delta$) of
-these same gene products, with $D \in [0, 0.1]$ and $\delta \in [0, 0.05]$.\\
+these same gene products, with $D \in [0, 0.1]$ and $\delta = \log_{2}phl$, where
+$phl \in [5, 50]$ is the gene product half-life.\\
 

### 2021-07-03 docs/manuscript/manuscript.tex
Change order of supplementary figures

@@
 system. We chose that number of time steps because numerical experiments showed
-that GRNs reached the steady state in approximately 300 steps (\nameref{S2_Fig}).
+that GRNs reached the steady state in approximately 300 steps (\nameref{S1_Fig}).
 
@@
 a matrix  $T = (t_{ij})$ with each row corresponding to a different topology
-and 17 columns corresponding to subgraphs shown in \nameref{S4_Fig}.
+and 17 columns corresponding to subgraphs shown in \nameref{S2_Fig}.
 Each element $t_{ij}$ of
@@
 simulations we redefined the morphogen gradient in such a way that the initial
-and final concentration were constant in all the fields (\nameref{S5_Fig}A).
+and final concentration were constant in all the fields (\nameref{S3_Fig}A).
 In order to know if topologies with the same topology had a similar fitness
@@
 (which is a variation of the I3-FFL) also reported by Cotterell \& Sharpe
-(see \nameref{S4_Fig} and \nameref{S1_Table}).
+(see \nameref{S2_Fig} and \nameref{S1_Table}).
 
@@
 4.66, 56.82, 56.14 and 41.05 percent of the GRNs had a fitness greater than 0.9
-in fields with 10, 20, 40 and 50 cells, respectively (\nameref{S5_Fig}B).
+in fields with 10, 20, 40 and 50 cells, respectively (\nameref{S3_Fig}B).
 
@@
 of drastically changing the expression mechanism. This can be seen in the
-expression dynamics of networks with different topologies (\nameref{S7_Fig}).
+expression dynamics of networks with different topologies (\nameref{S4_Fig}).
 
@@
 then begins to activate the gene A also in the center of the field
-(\nameref{S1_Fig}).
+(\nameref{S5_Fig}).
 % (Fig~\ref{fig:topol9-dyn}).
@@
 mechanisms for producing striped patterns are available and accessible to each
-other, i.e., an organism can change between different GRNs (for example between
+other through bridge topologies (\nameref{S6_Fig}), i.e.,
+an organism can change between different GRNs (for example between
 a Gierer-Meinhardt-like network to an Incoherent Feed-Forward Loop) without losing its
@@
 to changes in morphogenetic field size, but interestingly, GRNs with the same
-topology behave in different ways for each field size (\nameref{S5_Fig}C),
+topology behave in different ways for each field size (\nameref{S3_Fig}C),
 indicating that the specific values of network parameters are more important
@@
 same striped phenotype and a diverse combination of diffusion and degradation
-parameters (\nameref{S6_Fig}).
+parameters (\nameref{S7_Fig}).
 For example, in a single input model the I4-FFL could
@@
 \label{S1_Fig}
-{\bf Dynamics of expression of topology number 9.}
-The dynamics of expression is shown from $t = 5$ to $t = 150$. The red line
-represents the expression level of the gene C along the morphogenetic field,
-whereas the dotted line represents the expression level of gene A. The
-striped pattern of gene expression can be seen for both genes since $t = 80$.
-
-\paragraph*{S2 Fig.}
-\label{S2_Fig}
 {\bf Most GRNs reach the steady state before 250 time steps.}
@@
 
-\paragraph*{S3 Fig.}
-\label{S3_Fig}
-{\bf Bridge topologies in the neutral network.}
-``Bridge topologies'' are those topologies that connect two clusters in the neutral
-network (Fig~\ref{fig:neutral-network}), and as such can be interpreted as
-intermediary steps in the evolution from a cluster of related topologies into
-another cluster. For example the topology 441 could be an initial step to reach the
-topology 9, as in this topology nodes B and C are not connected.
-
-\paragraph*{S4 Fig.}
-\label{S4_Fig}
+\paragraph*{S2 Fig.}
+\label{S2_Fig}
 {\bf Relevant subgraphs present in the set of GRNs.}
@@
 
-\paragraph*{S5 Fig.}
-\label{S5_Fig}
+\paragraph*{S3 Fig.}
+\label{S3_Fig}
 {\bf Response of GRNs to changes in morphogenetic field size.}
@@
 
+\paragraph*{S4 Fig.}
+\label{S4_Fig}
+{\bf Expression dynamics of main networks.}
+
+\paragraph*{S5 Fig.}
+\label{S5_Fig}
+{\bf Dynamics of expression of topology number 9.}
+The dynamics of expression is shown from $t = 5$ to $t = 150$. The red line
+represents the expression level of the gene C along the morphogenetic field,
+whereas the dotted line represents the expression level of gene A. The
+striped pattern of gene expression can be seen for both genes since $t = 80$.
+
 \paragraph*{S6 Fig.}
 \label{S6_Fig}
+{\bf Bridge topologies in the neutral network.}
+``Bridge topologies'' are those topologies that connect two clusters in the neutral
+network (Fig~\ref{fig:neutral-network}), and as such can be interpreted as
+intermediary steps in the evolution from a cluster of related topologies into
+another cluster. For example the topology 441 could be an initial step to reach the
+topology 9, as in this topology nodes B and C are not connected.
+
+\paragraph*{S7 Fig.}
+\label{S7_Fig}
 {\bf Diffusion and degradation parameters.}
@@
 
-\paragraph*{S7 Fig.}
-\label{S7_Fig}
-{\bf Expression dynamics of main networks.}
-
 \paragraph*{S1 Table.}
@@
 {\bf Proportion of relevant subgraphs in the set of GRNs.}
-The 19 subgraphs presented in this table are those presented in \nameref{S4_Fig}.
+The 19 subgraphs presented in this table are those presented in \nameref{S2_Fig}.
 Most of the topologies presented at least one negative feedback loop (70.73\%),

### 2021-07-03 docs/manuscript/manuscript.tex
Add number of multi-input topologies

@@
 genetic algorithm. However, we also found a great diversity of network
-topologies that could produce a striped gene expression pattern.\\
+topologies that could produce a striped gene expression pattern, where 615 out of
+714 were multi-input topologies, showing that considering multiple
+morphogen inputs allows to find different solution for the patterning
+problem.\\
 

### 2021-07-03 docs/manuscript/manuscript.tex
Mention sensibility to morphogenetic field size

@@
 is dynamically more complex than that of other topologies such as the simple
-I3-FFL. Initially the expression pattern of the gene A is uniform in the
+I3-FFL; however, and despite the fact that the Gierer-Meinhardt model form
+patterns that scale with tissue size~\cite{shoaf1984}, we found that topology 9
+was more sensible to changes in the morphogenetic field size than those
+topologies based on I3-FFL (\nameref{S2_Table}),
+indicating that the additional regulatory interactions need to be fine-tuned
+depending on the number of cells in the field.\\
+
+In the mechanism of expression of topology 9 the expression pattern of the
+gene A is initially uniform in the
 morphogenetic field, but later it is shaped by the morphogen gradient in such a
@@
 
-This is a complex mechanism described by the first time in this study, and it
-would be interesting to study its feasibility and possible implementation in
-synthetic gene regulatory networks.
+This is a complex mechanism remindful of the Gierer-Meinhardt model, but
+described by the first time in this study in the context of multiple morphogen
+inputs, and it would be interesting to study its feasibility and possible
+implementation in synthetic gene regulatory networks.
 
@@
 
+\paragraph*{S2 Table.}
+\label{S2_Table}
+{\bf Average fitness of each topology in different morphogenetic fields.}
+The average fitness by topology is reported for morphogenetic fields with 10, 20,
+40 and 50 cells.
+
 \section*{Aknowledgements}

### 2021-07-04 docs/manuscript/manuscript.tex
fix typo: gen -> gene
Cite Scikit-Bio

@@
 concentration gradient (Fig~\ref{fig:model}C). The initial concentration of
-each gen product in all cells was set to 1 in all simulations.\\
+each gene product in all cells was set to 1 in all simulations.\\
 
@@
 performed a Neighbor Joining clustering using the package Scikit Bio 0.5.5 from
-Python 3.
+Python 3~\cite{skbio2020}.
 
@@
 consisted in the Hamming distance between the row vectors in matrix $T$ between
-each pair of topologies. We then performed a clustering of these topologies
-using the Neighbor Joining method in the Scikit Bio package.
+each pair of topologies. We then clustered the topologies based on the
+Neighbor Joining method, as implemented in the Scikit Bio
+package~\cite{skbio2020}.
 

### 2021-07-04 docs/manuscript/manuscript.tex
Change title of morphogenetic field size section

@@
 
-\subsection*{Change in morphogenetic field size}
+\subsection*{Pattern formation under varying size of the morphogenetic field}
 We wanted to evaluate if the GRNs evolved in our model with a morphogenetic

### 2021-07-04 docs/manuscript/manuscript.tex
Rephrase text

@@
 
-Cells interact continuously with their environment and they must detect signals
-from it. In many cases Gene Regulatory Networks (GRNs) interpret these signals
-through gene interaction and produce a particular response. This information
+Cells interact continuously with their environment, which requires very specific
+regulatory strategies to avoid potential detrimental responses. Arguably, most
+cellular functions arise from the dynamic activity of Gene Regulatory Networks
+(GRNs), which play a central role in the interpretation of external and internal
+signals. This information
 processing function is of particular importance in developmental processes such
@@
 has proven to be a complex process, and a major goal of developmental biology
-is to understand mechanistically how positional information given by morphogens
-is translated in spatial differentiation.\\
+is to understand mechanistically how positional information conveyed by morphogens
+is translated into spatial differentiation.\\
 
@@
 
-The optimal pattern of gene expression defined by us, consisted in cells at the
-sides of the field ($n<11$ and $n>20$) displaying expression levels lower than
+The optimal pattern of gene expression defined in this study consists in cells at
+the border of the field ($n<11$ and $n>20$) displaying expression levels lower than
 10\% of the maximal level observed along the field of cells for the output gene,
@@
 (Fig~\ref{fig:deg-dist}A). Nodes corresponding to topologies number one and
-three showed the highest degree ($deg(1) = deg (3) = 14$). Curiously, nodes with
-higher degree resulted to be also those network topologies with higher abundance
+three showed the highest degree ($deg(1) = deg (3) = 14$). Curiously, we found
+that nodes with higher degree turned out to be also those network topologies
+with higher abundance
 (Fig~\ref{fig:deg-dist}B). This relationship between abundance and node degree
@@
 
-We wanted to know if connecting topologies separated by two mutational steps
-would produce a connected graph, for that reason we generated another neutral
-network joining neighbors at two mutational steps and we found that only
-topology number 420 remained unconnected from the rest of topologies in the
-graph (Fig~\ref{fig:2neut-net}).
+Furthermore, we examined the extent to which the metagraph of topologies
+remained highly connected when joining neighbors at two mutational steps.
+Interestingly, we found that only topology number 420 remained isolated from
+the rest of topologies in the metagraph (Fig~\ref{fig:2neut-net}).
 
@@
 
-Besides topological robustness, we wanted to know if network topologies were
-robust to perturbations in the values of the parameters. After performing the
-perturbations to a GRN from each topology, we plotted the abundance of the
-topology vs. its robustness. We did not find a significant correlation between
-these two variables ($r_S = −0.0389724$, $\text{p-value} = 0.298475$)
-(Fig~\ref{fig:ab-rob}A). However, when we grouped those topologies containing
+Next to topological robustness, we investigated the robustness of the sampled
+network topologies to parameter changes (see Methods). Fig~\ref{fig:ab-rob}A
+illustrates the association between the abundance of a topology class versus its
+robustness. We did not find a significant correlation between
+these two variables ($r_S = −0.0389724$, $\text{p-value} = 0.298475$).
+However, when we grouped those topologies containing
 the same subgraph and calculated their mean robustness, we found that between

### 2021-07-04 docs/manuscript/manuscript.tex
Rephrase Abstract::Methods

@@
 \subsection*{Methods}
-In this study we used a genetic algorithm in order to explore the genotypic
-space of three-node GRNs capable of producing a band expression pattern in a
-group of cells. In contrast to previous studies, we explore multiple input
-network designs that enable more morphogen interpretation mechanisms to arise.
+In this study we apply a Markov chain Monte Carlo (MCMC)-like algorithm to probe
+the design space of three-node GRNs with the ability to generate a band-like
+expression pattern (target phenotype) in the middle of an arrangement of 30
+cells, which resemble a simple (1-D) morphogenetic field in a developing embryo.
+Unlike the majority of modeling studies published so far, here we explore the
+space of GRN topologies with nodes having the potential to perceive differently
+the same input signal. This allows for a lot more flexibility during the search
+space process, and thus enables us to identify a larger set of potentially
+interesting and realizable morphogen interpretation mechanisms.
 

### 2021-07-04 docs/manuscript/manuscript.tex
Rephrase Abstract::Results

@@
 Out of 2061 GRNs selected using the genetic algorithm, we found 714 groups of
-topologies that were able to interpret a morphogen input. The main network motif
-involved in the formation of the band expression pattern was the type 3
-Incoherent Feed Forward Loop (I3-FFL); we also present here other network
-topologies that have not been previously reported and that can be useful as
-synthetic regulatory networks due to its simple design and robust morphogen
-interpretation mechanisms.
+network topologies that were able to correctly interpret the morphogen.
+Importantly, the main network motif found to generate the target phenotype in
+response to the input signal was the type 3 Incoherent Feed Forward Loop (I3-FFL),
+which is in agreement with previous theoretical expectations and experimental
+observations. Notably, compared to previously reported pattern forming GRN
+topologies, we have uncovered a great variety of novel network designs, some of
+which might be worth inquiring through synthetic biology methodologies to test
+for the ability of network design with minimal regulatory complexity to robustly
+interpret a developmental cue.
 

### 2021-07-05 docs/manuscript/manuscript.tex
Rephrase text

@@
 GRN was greater or equal than that of the previous GRN, mutated GRN was saved
-and the cycle continued from step 2. This cycle was repeated 5000000 times and
-GRNs with a fitness greater or equal than 0.95 were saved into a file.
+and the cycle continued from step 2. This cycle was repeated 50000, and a GRN
+sampled at a given step and with a fitness score ≥ 0.95 was saved to a file as
+long as its associated topology had not been sampled in previous steps.
 
@@
 
-In order to group all the isomorphic networks in the set of 2061 GRNs we used a
+In order to group all the isomorphic (i.e. topologically equivalent) networks from
+the totality (2061) of sampled GRNs we used a
 classification algorithm that selected a random GRN, then it generated all
@@
 
-Later we calculated the distance between all the pairs of tensors and we
-performed a Neighbor Joining clustering using the package Scikit Bio 0.5.5 from
+Next, we calculated the distance between all the pairs of tensors and performed
+a Neighbor Joining clustering using the package Scikit Bio 0.5.5 from
 Python 3~\cite{skbio2020}.
@@
 
-The average vertex degree indicates that in general, all the network topologies
-display high robustness and that they could resist an average of approximately 4
-topological mutations without losing their fitness. Additionally, the average
-path length indicates that the evolutionary landscape is easily accessible for
-most network topologies (Table~\ref{table1}). Although the GRNs in this modelling
-framework are robust to changes in its topology, they are sensitive to changes in
-the morphogenetic field size, especially when this size is reduced, as only
-4.66, 56.82, 56.14 and 41.05 percent of the GRNs had a fitness greater than 0.9
-in fields with 10, 20, 40 and 50 cells, respectively (\nameref{S3_Fig}B).
+On the other hand, analysis of the average vertex degree seems to be indicative
+that network topologies are likely to withstand numerous topological
+modifications (approximately 4) without significant loss in their ability to
+generate a striped pattern of gene expression, representing a high fitness
+solution to the prescribed task. Additionally, the average path length seems
+to suggest that the evolutionary landscape could be easily traversed by
+most network topologies without incurring in significant fitness loss as long
+as mutational changes results in moves between adjacent neighbors across
+the metagraph (Table~\ref{table1}). It is interesting to note that,
+although the examined GRNs seem to exhibit, on average, a high robustness to
+topological/parameter changes, they generally show a rather high sensitivity to
+variation in the size of the morphogenetic field. For instance, we found that
+when the size of the morphogenetic field was fixed at 10, 20, 40 and 50 cells
+only 4.66\%, 56.82\%, 56.14\% and 41.05\%, respectively, of the examined
+GRNs showed a fitness greater than 0.9 (\nameref{S3_Fig}B).
 

### 2021-07-05 docs/manuscript/manuscript.tex
Rephrase text

@@
 \noindent
-being $a_i$ the degree of \emph{i}-th node and $V$ the number of nodes in the
-network. The network descriptors were calculated using built-in functions from
-Wolfram Mathematica software.\\
-
-We performed a Pearson’s chi-square goodness of fit test for the node degree
-distribution using Wolfram Mathematica.
+where $a_i$ denotes the degree of the \emph{i}-th node and $V$ the number of
+nodes in the network. Both the network descriptors and Pearson’s chi-square
+goodness of fit tests were performed for the node degree distribution using
+the built-in functions from the Wolfram Mathematica software.\\
 
 \subsection*{Pattern formation under varying size of the morphogenetic field}
-We wanted to evaluate if the GRNs evolved in our model with a morphogenetic
-field size of 30 cells could produce the desired pattern in morphogenetic fields
-with a different number of cells. For this reason, we calculated the fitness of
-each GRN in morphogenetic fields with 10, 20, 40, and 50 cells. In these
-simulations we redefined the morphogen gradient in such a way that the initial
-and final concentration were constant in all the fields (\nameref{S3_Fig}A).
-In order to know if topologies with the same topology had a similar fitness
-in the different fields, we performed a Principal Component Analysis in
-Scikit-learn~0.24.2~\cite{sklearn} using the fitness values of each GRN in
-the four different fields.
+We evaluated the ability of each of the sampled GRNs to generate a striped
+pattern of gene expression under varying sizes of the morphogenetic field.
+Specifically, we calculated the fitness of each GRN across morphogenetic fields
+composed of 10, 20, 40, and 50 cells. In these simulations we redefined the
+morphogen gradient in such a way that the initial and final concentration were
+constant in all the fields (\nameref{S3_Fig}A). In order to know if topologies
+with the same topology had a similar fitness in the different fields, we performed
+a Principal Component Analysis in Scikit-learn~0.24.2~\cite{sklearn} using the
+fitness values of each in the four different fields.
 
@@
 The classification of GRNs showed that our initial set of 2061 GRNs can be
-grouped in a number of 714 distinct network topologies, each one of them
-containing a different number of GRNs (Fig~\ref{fig:distopol}). This abundances
+grouped into 714 distinct network topology classes, with each containing a
+variable number of GRNs (Fig~\ref{fig:distopol}). This abundances
 distribution was not produced by chance (Shannon entropy = 8.52963, 95\%
@@
 found nodes inside the neutral network that had more connections between them
-than with other nodes, generating clusters of nodes, interestingly in
-almost every cluster is present one or more of the 15 most abundant network
+than with other nodes, which leads to further clustering of nodes. Interestingly in
+almost every cluster one can find one or more of the 15 most abundant network
 topologies. In addition, we found that the first eight most abundant network
@@
 the same subgraph and calculated their mean robustness, we found that between
-the most robust subgraphs is the ``Bistable'' motif reported by Cotterell \&
-Sharpe~\cite{Cotterell2010}, the I3-FFL and the ``Overlapping domains'' motif
-(which is a variation of the I3-FFL) also reported by Cotterell \& Sharpe
+the most robust subgraphs is the ``Bistable'' motif, the I3-FFL, as well as
+the ``Overlapping domains'' motif (which is a variation of the I3-FFL), which
+have been previously reported in Cotterell \& Sharpe~\cite{Cotterell2010}
 (see \nameref{S2_Fig} and \nameref{S1_Table}).
@@
 gene expression pattern with only two genes (Fig~\ref{fig:topol9}).
-This topology is composed of a core that consists in the
-Gierer-Meinhardt network~\cite{giererMeinh1972, meinhardt2000} known to
-generate stripes and dots in bidimensional fields, and additional
-interactions such as the self inhibition of gene A and the inhibition of the genes
-by the morphogen. The Gierer-Meinhardt network consists in an activator of low
-diffusivity (gene C) and an inhibitor with greater diffusivity (gene A). The additional
-interactions of topology number 9 have the role of restricting the generation of
-the expression band to a precise location.
+This topology is composed of a core quite similar to the Gierer-Meinhardt
+mechanism~\cite{giererMeinh1972, meinhardt2000} known to generate
+stripes and dots in bidimensional fields, with the only differences being
+the self inhibition of gene A and the inhibition of the genes by the morphogen.
+Such additional interactions in the topology number 9 play a role in restricting
+the striped expression pattern to a precise location along the morphogenetic field.
 

### 2021-07-05 docs/manuscript/manuscript.tex
Rephrase Author summary

@@
 \section*{Author summary}
-Systems biology is a very fast growing field thanks to the advances in computing and
-mathematical modeling of biological systems, and has proven to be a powerful tool that
-can help us to understand complex biological processes such as pattern formation in
-development. In this article we use an \textit{in silico} approach in order to find and study
-the design space of multiple input three node Gene Regulatory Networks that are able
-to produce a striped pattern of gene expression.
+Systems biology is a fast growing field largely powered by advances in
+high-performance computing and sophisticated mathematical modeling of
+biological systems. Based on these advances, we are now in a position to not
+only mechanistically understand but also to accurately predict the behavior of
+complex biological processes, including cell differentiation and spatial pattern
+formation during embryogenesis. In this article we use an \textit{in silico}
+approach to probe the design space of multi-input, three node Gene Regulatory
+Networks (GRNs) capable of generating a striped pattern of gene expression in
+the context of a simplified 1-D morphogenetic field.
 

### 2021-07-05 docs/manuscript/manuscript.tex
Rephrase introduction

@@
 
-Producing a stripe of gene expression is a fundamental patterning function
-in development, and it has been shown that feed-forward motifs are minimal
-networks capable of performing this function~\cite{munteanu_2014}. For example,
-Cotterell and Sharpe studied what kind of 3-node network topologies could
-interact with a morphogen gradient and produce a stripe of gene expression
-in a one-dimensional field of cells and found a variety of networks that
-implemented at least 7 different mechanisms and had different levels of
-complexity, being most of them variations of feed-forward motifs~\cite{Cotterell2010}.
-
-However, although several studies have analyzed the properties of GRNs that
-process positional information, they have focused on particular natural
-networks~\cite{Schaerli2014,Jaeger2004}, in synthetic networks~\cite{Basu2005,
-Reinitz1995,Schaerli2018,Elowitz2000}, or have constrained the
-design of the GRNs to accept an input signal from only one of its nodes, which is
-not always biologically realistic as multiple input systems can exist~\cite{Cotterell2010,
-munteanu_2014, Schaerli2014}. For example, the neural subtypes specification
-system in the vertebrate neural tube is a two-input network in which
-Sonic Hedgehog acts as a morphogen and Olig2 and Nkx2.2 act as the receiver
-nodes~\cite{dessaud_dynamic_2010, balaskas2012, exelby_2021}.
+In a tissue context, the generation of a stripe of gene expression is a fundamental
+patterning function in development, and it has been shown that simple feed-forward
+motifs can robustly achieve such a patterning task~\cite{munteanu_2014}. For
+example, Cotterell and Sharpe studied  what kind of 3-node network topologies
+could effectively translate a morphogen gradient into a striped gene expression
+pattern in a one-dimensional field of cells, and found a variety of networks that
+implemented at least 7 different mechanisms with varying levels of complexity,
+being most of them variations of feed-forward motifs~\cite{Cotterell2010}.
+
+Although GRNs with positional information processing capacities have been
+extensively analyzed, most studies have so far placed emphasis on particular
+regulatory systems that respond to a single input only. These studies have
+typically focused on either regulatory systems observed in
+nature~\cite{vonDassow2000,Reinitz1995,Jaeger2004}, synthetic
+implementations~\cite{Basu2005,Schaerli2018,Elowitz2000,barbier2020} with
+predefined topologies, or have deliberately constrained the study of GRNs to just
+a handful of alternative designs~\cite{munteanu_2014,Schaerli2014}. However,
+previous work has demonstrated the necessity to expand computational analysis
+of pattern forming GRNs to multi-input settings. For instance, it has been shown
+that the neural subtypes specification system in the vertebrate neural tube involves
+a  two-input network in which Sonic Hedgehog acts as a morphogen and the Olig2
+and Nkx2.2 genes can act as the receiver nodes~\cite{dessaud_dynamic_2010,
+balaskas2012, exelby_2021}.
 

### 2021-07-06 docs/manuscript/manuscript.tex
Rephrase study description

@@
 
-In this study, we used a genetic algorithm to generate a set of 2061 three-node GRNs
-including single and multiple input topologies that were able to interpret a morphogen
-signal and respond to it generating a striped pattern of gene expression in order to
-analyse the role of network motifs in this process and its dynamic properties.
+In this study, we apply a Markov chain Monte Carlo (MCMC)-like algorithm to
+probe the design space of three-node GRNs looking for topologies capable of
+translating a morphogen gradient  (input signal) into a striped pattern of gene
+expression (phenotype). Importantly, unlike most previous computational
+studies, here we allowed any node (e.g. gene) in a GRN to perceive the input
+signal in varying ways, and selected the best performing GRNs based on a
+fitness criterion used to assess the quality of the phenotype with respect to
+a prescribed optimal pattern. Based on this computational strategy, we were
+able to uncover a great variety of distinct classes of network topologies that
+tend to form a complex interconnected meta-graph that could be easily
+traversed throughout evolution via single changes in the wiring of the
+different GRNs.
 

### 2021-07-06 docs/manuscript/manuscript.tex
Add ref to S3 Fig in Methods

@@
 a Principal Component Analysis in Scikit-learn~0.24.2~\cite{sklearn} using the
-fitness values of each in the four different fields.
+fitness values of each in the four different fields (see \nameref{S3_Fig}).
 

### 2021-07-08 docs/manuscript/manuscript.tex
Rephrase text

@@
 together but instead we found GRNs with different topologies grouped as close
-neighbors (Fig~\ref{fig:clustering}A). More interestingly, some GRNs displaying
+neighbors (Fig~\ref{fig:clustering}A). Intriguingly, some GRNs displaying
 very abundant network topologies were grouped together with GRNs displaying
-very infrequent network topologies, indicating that additional interactions
-in the networks have often the function of fine tunning gene expression, instead
-of drastically changing the expression mechanism. This can be seen in the
-expression dynamics of networks with different topologies (\nameref{S4_Fig}).
+very infrequent network topologies. This seems to indicate that additional
+interactions in the networks are likely to represent alternative regulatory
+mechanisms for fine tuning gene expression, instead of drastically changing
+the expression mechanism, which can be seen in the expression dynamics of
+networks with different topologies (\nameref{S4_Fig}).
 

### 2021-07-08 docs/manuscript/manuscript.tex
Rephrase text

@@
 
-This study can help us to understand the mechanisms by which the GRNs work in
-nature as modules that make part of genomes, and it also can provide guidance
-in the construction of new synthetic GRNs in the field of synthetic biology.
-
-We found the presence of network motifs in the set of GRNs produced by the
-genetic algorithm. However, we also found a great diversity of network
+In this work, we have conducted extensive computational explorations of small
+GRNs with the capacity to generate a simple spatial pattern in a 1-D
+morphogenetic field. Our findings shed new light on regulatory rules and
+dynamical mechanisms that  GRNs observed in nature could potentially
+implement to achieve fundamental spatio-temporal tasks, which might
+serve as guidance for future synthetic biology implementations.
+
+Overall, based on our extensive search space exploration we were able to identify
+a great variety of pattern-forming GRN topologies  with distinctive and often
+widely shared building blocks (network motifs), which are likely to confer on
+each topology the ability to achieve specific regulatory tasks in a semi-autonomous
+manner. However, we also found a great diversity of network
 topologies that could produce a striped gene expression pattern, where 615 out of
@@
 morphogen inputs allows to find different solution for the patterning
-problem.\\
-
-Robustness is an important feature of patterning mediated by morphogens, and
-the GRNs involved can be selected in an evolutionary process not only by their final
-steady state but also by their ability to buffer noise~\cite{lo_robust_2015, exelby_2021}.
-In our study, the most abundant topologies were
-found to be part of highly connected clusters in the neutral network, and this fact
-indicates that these topologies are both robust to quantitative changes in the
-strength of interactions and diffusion and degradation parameters as well as to
-qualitative changes in the topology itself. This, in principle, means that many
-mechanisms for producing striped patterns are available and accessible to each
-other through bridge topologies (\nameref{S6_Fig}), i.e.,
-an organism can change between different GRNs (for example between
-a Gierer-Meinhardt-like network to an Incoherent Feed-Forward Loop) without losing its
-fitness, which allows to fine-tune the regulatory system according to its specific
+problem. Most importantly, we found that this ensemble of pattern-forming GRNs
+tend to form a highly connected meta-graph, which highlights at a very high level
+emergent properties, such as robustness and evolvability, of complex biological
+systems \cite{kitano2004}.\\
+
+Robustness is a defining feature of any developmental patterning process,
+and our results suggests that the underlying morphogen interpretation
+mechanisms (GRNs) are highly accessible through exploration of the
+topology space constrained not only by selection for their final steady state,
+but also by their ability to buffer noise~\cite{lo_robust_2015, exelby_2021}.
+
+In our study, the most abundant topologies were found forming highly
+connected clusters in a so-called neutral network of genotypes,
+implying that the corresponding GRN topologies would tend to be robust
+to quantitative changes in the strength of regulatory interactions, diffusion and
+degradation parameters, as well as to qualitative changes in the topology
+itself. In essence, such organization could be seen as an intrinsic evolutionary
+property of GRNs that facilitates the coexistence of robustness and
+evolvability~\cite{wagner_robustness_2008}. For instance, bridge topologies
+in the neutral network connect distinct classes of pattern-forming
+GRNs~(\nameref{S6_Fig}), which suggests that an organism could switch
+between different GRNs (for example between a Gierer-Meinhardt-like network
+to an Incoherent Feed-Forward Loop) with minimal fitness costs. This high-level
+property of the neutral network of topologies could thus facilitate the fine-tuning
+of the regulatory systems through genetic modifications in order to meet specific functional requirements under selective pressures.
+
+ which allows to fine-tune the regulatory system according to its specific
 needs. As robustness and evolvability are

### 2021-07-09 docs/manuscript/manuscript.tex
Rephrase text

@@
 property of the neutral network of topologies could thus facilitate the fine-tuning
-of the regulatory systems through genetic modifications in order to meet specific functional requirements under selective pressures.
-
- which allows to fine-tune the regulatory system according to its specific
-needs. As robustness and evolvability are
-not mutually exclusive properties~\cite{wagner_robustness_2008}, it is likely
-that these topologies are also evolvable in the sense that they can produce
-heritable variation of the expression pattern.\\
+of the regulatory systems through genetic modifications in order to meet
+specific functional requirements under selective pressures.
 

### 2021-07-09 docs/manuscript/manuscript.tex
change genetic for search space

@@
 \subsection*{Results}
-Out of 2061 GRNs selected using the genetic algorithm, we found 714 groups of
+Out of 2061 GRNs selected using the search space algorithm, we found 714 groups of
 network topologies that were able to correctly interpret the morphogen.
@@
 
-\subsection*{Genetic algorithm}
-
-The genetic algorithm used to produce the set of gene regulatory networks (GRNs)
-consisted in the following steps:
+\subsection*{Search space algorithm}
+
+The Markov chain Monte Carlo (MCMC)-like algorithm used to produce the set of
+gene regulatory networks (GRNs) consisted in the following steps:
 
@@
 
-The source code for the genetic algorithm and other parts of analysis is available at
-\url{https://github.com/Nesper94/gene-regulatory-networks}.
+The source code for the search space algorithm and other parts of analysis is
+available at \url{https://github.com/Nesper94/gene-regulatory-networks}.
 
@@
 generate a striped pattern of gene expression, representing a high fitness
-solution to the prescribed task. Additionally, the average path length seems
+solution to the prescribed task. The average complexity index of network
+topologies was 39.2435 ($\pm 11.6768$) and was found to be negatively
+correlated with the abundance ($r_S = -0.316653$, $\text{p-value} =
+2.25161\times10^{-18}$) (Fig~\ref{fig:ab-rob}B).
+Additionally, the average path length seems
 to suggest that the evolutionary landscape could be easily traversed by
@@
 
-The average complexity index of network topologies was 39.2435 ($\pm 11.6768$)
-and was found to be negatively correlated with the abundance ($r_S = -0.316653$,
-$\text{p-value} = 2.25161\times10^{-18}$) (Fig~\ref{fig:ab-rob}B).
-
 Hierarchical clustering of spatiotemporal expression profiles showed that GRNs
@@
 the expression mechanism, which can be seen in the expression dynamics of
-networks with different topologies (\nameref{S4_Fig}).
+networks with different topologies (\nameref{S4_Fig}). Additionally, clustering
+of topologies revealed that there are six main groups of topologies in our set
+of 2061 GRNs (Fig~\ref{fig:clustering}B).
 
@@
 \end{figure}
-
-Clustering of topologies revealed that there are six main groups of topologies
-in our set of 2061 GRNs (Fig~\ref{fig:clustering}B).
 
@@
  with topology 9. Gene \textbf{B} is not shown as it was lost in this topology
- in the evolutionary process simulated by the genetic algorithm.
+ in the evolutionary process simulated by the search space algorithm.
  \label{fig:topol9}
@@
 
-Although the topology 9 consists of only two genes, its mechanism of expression
-is dynamically more complex than that of other topologies such as the simple
+Despite having only two genes, the regulatory mechanism underlying
+the topology 9 generates dynamically more complex expression patterns
+compared to other topologies, such as the
 I3-FFL; however, and despite the fact that the Gierer-Meinhardt model form
@@
 
-In the mechanism of expression of topology 9 the expression pattern of the
-gene A is initially uniform in the
-morphogenetic field, but later it is shaped by the morphogen gradient in such a
-way that two important events occur, in first place the overall expression
-level of the gene A is decreased by the inhibition of the morphogen, releasing
-gene C from its repression by gene A; in second place, it appears a
-non-homogeneous spatial distribution in which gene A is expressed at higher
-levels in the cells at the rightmost part of the morphogenetic field.\\
+In the mechanism of expression of topology 9 the initial expression
+pattern of gene A is uniform across the morphogenetic field, but later
+on the morphogen gradient shapes its dynamics in such a way that
+two important events occur: firstly, the overall expression level of the
+gene A is decreased by the inhibition of the morphogen, releasing
+gene C from its repression by gene A; next, gene A tend to display a
+non-homogeneous spatial distribution with higher expression levels
+being shifted towards the cells at the rightmost part of the morphogenetic
+field.\\
 
@@
 but also by their ability to buffer noise~\cite{lo_robust_2015, exelby_2021}.
-
 In our study, the most abundant topologies were found forming highly
@@
 pattern,~\cite{Schaerli2014, Basu2005}, in our study this motif was present in
-only 10.92\% of the topologies. This fact may be because genetic algorithms
+only 10.92\% of the topologies. This fact may be because evolutionary algorithms
 can not always find global minima but local minima or its surroundings, however,
@@
 
-The cladogram of spatiotemporal expression profiles shows that the genetic
+The cladogram of spatiotemporal expression profiles shows that the search space
 algorithm selected similar expression mechanisms that can be produced by very

### 2021-07-09 docs/manuscript/manuscript.tex
State that quantities are dimensionless
Remove redundant step

@@
 interaction, or activation respectively. The magnitude of the value is
-proportional to the strength of the interaction. In this study we define
+proportional to the strength of the interaction; however, this and all other
+magnitudes in our model are dimensionless. In this study we define
 a 'genotype' as the set of interaction, diffusion and degradation values that
@@
 
- \item{\bf Mutate GRN network parameters or gene product diffusion or
-degradation rates randomly and evaluate fitness again.} If fitness of the
-mutated GRN is greater than that of the previous GRN, save GRN and continue
-mutating.
-
  \item {\bf Mutate GRNs randomly and evaluate fitness again.} Mutations

### 2021-07-09 docs/manuscript/manuscript.tex
fix error

@@
 GRN was greater or equal than that of the previous GRN, mutated GRN was saved
-and the cycle continued from step 2. This cycle was repeated 50000, and a GRN
+and the cycle continued from step 2. This cycle was repeated 50000 times, and a GRN
 sampled at a given step and with a fitness score ≥ 0.95 was saved to a file as

### 2021-07-10 docs/manuscript/manuscript.tex
Fix typo in Acknowledgements

@@
 
-\section*{Aknowledgements}
+\section*{Acknowledgements}
 The computational resources (Stevin Supercomputer Infrastructure) and

### 2021-07-10 docs/manuscript/manuscript.tex
Improve description of supp. figs

@@
 (A) Morphogen gradients of fields with 10, 20, 40 and 50 cells. (B) Fitness of the
-GRNs in each one of the morphogenetic fields. (C) Principal Component Analysis
-of fitness by morphogenetic field size. Although principal components separate
-GRNs in clusters, not all the GRNs with the same topology are located in the
+GRNs by topology in each one of the morphogenetic fields. (C) Principal Component
+Analysis of fitness by morphogenetic field size. Although principal components
+separate GRNs in clusters, not all the GRNs with the same topology are located in the
 same group. (D) Average fitness of GRNs evolved in different morphogenetic
@@
 {\bf Expression dynamics of main networks.}
+The dynamics of expression is shown from $t = 1$ to $t = 30$ for the eight most
+abundant topologies. The red line
+represents the expression level of the gene C along the morphogenetic field,
+the blue line represents the expression level of gene B and
+the dotted line represents the expression level of gene A.
 
@@
 B and C. (B) Ternary plot showing the combination of degradation parameters
-for genes A, B and C.
+for genes A, B and C. The color gradient that goes from topology 1 in white to
+topology 714 in black shows no distinguishable pattern.
 
@@
 The average fitness by topology is reported for morphogenetic fields with 10, 20,
-40 and 50 cells.
+40 and 50 cells. Data from this table are shown in \nameref{S3_Fig}B.
 

### 2021-07-10 docs/manuscript/manuscript.tex
improve methods: sigm

@@
 The input function is a sigmoid function described by Eq~\ref{ec:input-func}.
+Sigmoid functions are often used to approximate Hill functions describing gene
+activation/inhibition, and this approach has been used extensively to model the
+response of signaling pathways and gene interactions in GRNs \cite{mendoza2006,
+vu2007, zhang2013, armao2016}. Moreover, differential equation models using
+sigmoid functions have shown to fit gene expression data \cite{chen2005,
+dahlquist2015}.
+
 \begin{equation}

### 2021-07-10 docs/manuscript/manuscript.tex
Rephrase sigmoid section

--- 2021-07-10 docs/manuscript/manuscript.tex
@@
 
-The input function is a sigmoid function described by Eq~\ref{ec:input-func}.
+Sigmoid functions are often used to approximate Hill functions describing gene
+activation/inhibition in which steepness can be modulated by mechanisms as
+molecular cooperativity and target sequestration \cite{ricci2011}. This approach
+has been used extensively to model the response of signaling pathways and gene
+interactions in GRNs \cite{mendoza2006, vu2007, zhang2013, armao2016}.
+Moreover, differential equation models using sigmoid functions have shown to
+fit gene expression data \cite{chen2005, dahlquist2015}. In this study, the
+input function representing regulatory interactions is a sigmoid function
+described by Eq~\ref{ec:input-func}.
+
 \begin{equation}
@@
 the totality (2061) of sampled GRNs we used a
-classification algorithm that selected a random GRN, then it generated all
-isomorphs of that GRN by performing permutations in rows and columns. The
+classification algorithm that selected a random GRN and generated all
+isomorphs of that GRN by interchanging rows of the adjacency matrix
+($i \leftrightarrow j$) while interchanging the corresponding columns
+($i+1 \leftrightarrow j+1$). The
 algorithm then compared all six isomorphs with other GRN and classified the

### 2021-07-10 docs/manuscript/manuscript.tex
Update manuscript.tex

@@
 Gloria Machado-Rodríguez\textsuperscript{1},
-Boris Rodríguez-Rey\textsuperscript{2},
+Boris A. Rodríguez\textsuperscript{2},
 Jayson Gutiérrez\textsuperscript{3},
@@
 \subsection*{Background}
-A central problem in developmental and synthetic biology is to understand the
-mechanisms by which cells in a tissue or in a Petri dish process external cues
-and transform such information into a coherent response, e.g. a terminal
+A central problem in developmental and synthetic biology is understanding the
+mechanisms by which cells in a tissue or a Petri dish process external cues
+and transform such information into a coherent response, e.g., a terminal
 differentiation state. It was long believed that this type of positional
 information could be entirely attributed to a gradient of concentration of a
-certain signaling molecule (i.e. a morphogen). However, advances in experimental
+specific signaling molecule (i.e., a morphogen). However, advances in experimental
 methodologies and computer modeling have demonstrated the crucial role of the
@@
 carried by the morphogen, which is eventually translated into a spatial
-pattern. This morphogen interpretation mechanism has gained a lot of attention
+pattern. This morphogen interpretation mechanism has gained much attention
 in systems biology as a tractable system to investigate the emergent properties

### 2021-07-10 docs/manuscript/manuscript.tex
Update manuscript.tex

@@
 \subsection*{Methods}
-In this study we apply a Markov chain Monte Carlo (MCMC)-like algorithm to probe
+In this study, we apply a Markov chain Monte Carlo (MCMC)-like algorithm to probe
 the design space of three-node GRNs with the ability to generate a band-like
@@
 cells, which resemble a simple (1-D) morphogenetic field in a developing embryo.
-Unlike the majority of modeling studies published so far, here we explore the
+Unlike most modeling studies published so far, here we explore the
 space of GRN topologies with nodes having the potential to perceive differently
@@
 \subsection*{Results}
-Out of 2061 GRNs selected using the search space algorithm, we found 714 groups of
-network topologies that were able to correctly interpret the morphogen.
-Importantly, the main network motif found to generate the target phenotype in
-response to the input signal was the type 3 Incoherent Feed Forward Loop (I3-FFL),
-which is in agreement with previous theoretical expectations and experimental
-observations. Notably, compared to previously reported pattern forming GRN
+Out of 2061 GRNs selected using the search space algorithm, we found 714 classes of
+network topologies that could correctly interpret the morphogen.
+Notably, the main network motif that generated the target phenotype in
+response to the input signal was the type 3 Incoherent Feed-Forward Loop (I3-FFL),
+which agrees with previous theoretical expectations and experimental
+observations. Particularly, compared to a previously reported pattern forming GRN
 topologies, we have uncovered a great variety of novel network designs, some of
 which might be worth inquiring through synthetic biology methodologies to test
-for the ability of network design with minimal regulatory complexity to robustly
-interpret a developmental cue.
+for the ability of network design with minimal regulatory complexity to 
+interpret a developmental cue robustly.
 
@@
 high-performance computing and sophisticated mathematical modeling of
-biological systems. Based on these advances, we are now in a position to not
-only mechanistically understand but also to accurately predict the behavior of
+biological systems. Based on these advances, we are now in a position to
+mechanistically understand and accurately predict the behavior of
 complex biological processes, including cell differentiation and spatial pattern
-formation during embryogenesis. In this article we use an \textit{in silico}
-approach to probe the design space of multi-input, three node Gene Regulatory
-Networks (GRNs) capable of generating a striped pattern of gene expression in
+formation during embryogenesis. In this article, we use an \textit{in silico}
+approach to probe the design space of multi-input, three-node Gene Regulatory
+Networks (GRNs) capable of generating a striped gene expression pattern in
 the context of a simplified 1-D morphogenetic field.
@@
 
-Cells interact continuously with their environment, which requires very specific
+Cells interact continuously with their environment, which requires precise
 regulatory strategies to avoid potential detrimental responses. Arguably, most
 cellular functions arise from the dynamic activity of Gene Regulatory Networks
-(GRNs), which play a central role in the interpretation of external and internal
-signals. This information
-processing function is of particular importance in developmental processes such
-as those in which a group of cells differentiate in response to a signalling
-molecule. Such molecules were referred to as morphogens by Turing in
-1952~\cite{Turing1952} and posterior theoretical studies on patterning led to the
+(GRNs), which play a central role in interpreting external and internal
+signals. This information processing function is critical in 
+developmental processes such as those in which a group of cells differentiates in
+response to a signaling molecule. Such molecules were referred to as morphogens by Turing in
+1952~\cite{Turing1952}, and posterior theoretical studies on patterning led to the
 conceptualization of the French Flag Problem by Wolpert, who also stated
@@
 Although the Bicoid protein (bcd), the first example of a molecule that acted
-as a morphogen, was only found in the 80's in the developing embryos of
+as a morphogen, was only found in the 80s in the developing embryos of
 \textit{Drosophila melanogaster}~\cite{driever_gradient_1988,
@@
 Our understanding of how GRNs process information has increased thanks to the
-concept and theory of network motifs, which are defined as patterns of
-interconnections occurring in networks at numbers that are significantly higher
+concept and theory of network motifs, defined as patterns of
+interconnections occurring in networks at numbers significantly higher
 than those in randomized networks~\cite{Milo2002}. This novel approach, along
-with mathematical modeling and computational systems biology methods have been
-shown to be a powerful tool to study GRNs~\cite{kitano_computational_2002}, and
-the predictions of studies using these methods have been validated by
-experimental studies~\cite{Kalir2005,Kalir2004,Mangan2003a,ODonnell2005}, and
-some of them have even used synthetic GRNs to produce artificial cell
+with mathematical modeling and computational systems biology methods, is a 
+powerful tool to study GRNs~\cite{kitano_computational_2002}, and
+experimental studies have validated the predictions of studies using these
+methods~\cite{Kalir2005,Kalir2004,Mangan2003a,ODonnell2005}.
+Some have even used synthetic GRNs to produce artificial cell
 differentiation~\cite{Basu2005}, circadian gene
@@
 could effectively translate a morphogen gradient into a striped gene expression
-pattern in a one-dimensional field of cells, and found a variety of networks that
-implemented at least 7 different mechanisms with varying levels of complexity,
-being most of them variations of feed-forward motifs~\cite{Cotterell2010}.
+pattern in a one-dimensional field of cells. They found a variety of networks that
+implemented at least 7 different mechanisms with varying complexity levels,
+most of them variations of feed-forward motifs~\cite{Cotterell2010}.
 
 Although GRNs with positional information processing capacities have been
-extensively analyzed, most studies have so far placed emphasis on particular
+extensively analyzed, most studies have so far emphasized particular
 regulatory systems that respond to a single input only. These studies have
@@
 expression (phenotype). Importantly, unlike most previous computational
-studies, here we allowed any node (e.g. gene) in a GRN to perceive the input
-signal in varying ways, and selected the best performing GRNs based on a
-fitness criterion used to assess the quality of the phenotype with respect to
-a prescribed optimal pattern. Based on this computational strategy, we were
-able to uncover a great variety of distinct classes of network topologies that
+studies, here we allowed any node (e.g., gene) in a GRN to perceive the input
+signal in varying ways and selected the best-performing GRNs based on a
+fitness criterion used to assess the quality of the phenotype to
+a prescribed optimal pattern. Based on this computational strategy, we uncovered
+a great variety of distinct classes of network topologies that
 tend to form a complex interconnected meta-graph that could be easily
@@
 different GRNs.
-
-% Lorem ipsum dolor sit~\cite{bib1} amet, consectetur adipiscing elit.
-Nullam Eq~(\ref{eq:schemeP}) sapien tellus, commodo id velit id, eleifend volutpat quam.
-Nunc pellentesque pretium diam, quis maximus dolor faucibus id.~\cite{bib2} Nunc convallis sodales ante, ut ullamcorper est egestas vitae.
 

### 2021-07-10 docs/manuscript/manuscript.tex
Update manuscript.tex

@@
 implement to achieve fundamental spatio-temporal tasks, which might
-serve as guidance for future synthetic biology implementations.
-
-Overall, based on our extensive search space exploration we were able to identify
+guide for future synthetic biology implementations.
+
+Overall, based on our extensive search space exploration we identified
 a great variety of pattern-forming GRN topologies  with distinctive and often
-widely shared building blocks (network motifs), which are likely to confer on
+widely shared building blocks (network motifs), which are likely to confer
 each topology the ability to achieve specific regulatory tasks in a semi-autonomous
@@
 714 were multi-input topologies, showing that considering multiple
-morphogen inputs allows to find different solution for the patterning
+morphogen inputs allows finding a different solution for the patterning
 problem. Most importantly, we found that this ensemble of pattern-forming GRNs
-tend to form a highly connected meta-graph, which highlights at a very high level
-emergent properties, such as robustness and evolvability, of complex biological
-systems \cite{kitano2004}.\\
-
-Robustness is a defining feature of any developmental patterning process,
-and our results suggests that the underlying morphogen interpretation
-mechanisms (GRNs) are highly accessible through exploration of the
-topology space constrained not only by selection for their final steady state,
-but also by their ability to buffer noise~\cite{lo_robust_2015, exelby_2021}.
+tends to form a highly connected meta-graph, which highlights emergent
+properties, such as robustness and evolvability, of complex biological
+systems at a very high level~\cite{kitano2004}.\\
+
+Robustness is a defining feature of any developmental patterning process.
+Our results suggest that the underlying morphogen interpretation
+mechanisms (GRNs) are highly accessible by exploring the
+topology space constrained by selection for their final steady-state and
+by their ability to buffer noise~\cite{lo_robust_2015, exelby_2021}.
 In our study, the most abundant topologies were found forming highly
@@
 implying that the corresponding GRN topologies would tend to be robust
-to quantitative changes in the strength of regulatory interactions, diffusion and
+to quantitative changes in the strength of regulatory interactions, diffusion, and
 degradation parameters, as well as to qualitative changes in the topology
@@
 in the neutral network connect distinct classes of pattern-forming
-GRNs~(\nameref{S6_Fig}), which suggests that an organism could switch
-between different GRNs (for example between a Gierer-Meinhardt-like network
+GRNs~(\nameref{S6_Fig}), suggesting that an organism could switch
+between different GRNs (for example, between a Gierer-Meinhardt-like network
 to an Incoherent Feed-Forward Loop) with minimal fitness costs. This high-level
 property of the neutral network of topologies could thus facilitate the fine-tuning
-of the regulatory systems through genetic modifications in order to meet
+of the regulatory systems through genetic modifications to meet
 specific functional requirements under selective pressures.
@@
 qualitative and quantitative changes, we found that many GRNs were susceptible
-to changes in morphogenetic field size, but interestingly, GRNs with the same
-topology behave in different ways for each field size (\nameref{S3_Fig}C),
+to changes in morphogenetic field size. However, interestingly, GRNs with the same
+topology behave in different ways for each field size~(\nameref{S3_Fig}C),
 indicating that the specific values of network parameters are more important
@@
 
-Other studies have shown that Incoherent Feed Forward Loops are able to form
+Other studies have shown that Incoherent Feed-Forward Loops can form
 spatial stripes of gene expression~\cite{ishihara_cross_2005}. In particular,
@@
 in cell fate definition~\cite{Li2019}. Moreover, our results indicate that this
-is a robust motif that can be useful for construction of synthetic gene
-regulatory networks that are exposed to stochastic fluctuations in the
+is a robust motif that can be useful for constructing synthetic gene
+regulatory networks exposed to stochastic fluctuations in the
 environment and mutations.
@@
 However, as our model was less stringent and allowed the morphogen input to
-interact with any gene, we found a larger number of possible topologies with the
+interact with any gene, we found a more significant number of possible topologies with the
 same striped phenotype and a diverse combination of diffusion and degradation
-parameters (\nameref{S7_Fig}).
-For example, in a single input model the I4-FFL could
-not produce a stripe of gene expression~\cite{munteanu_2014}, whereas in our
-experimental setup 20.17\% of the topologies presented this motif as a
+parameters~(\nameref{S7_Fig}). For example, in a single input model, the I4-FFL could
+not produce a stripe of gene expression~\cite{munteanu_2014}, whereas, in our
+experimental setup, 20.17\% of the topologies presented this motif as a
 subgraph~(\nameref{S1_Table}).
 
-Although it has been experimentally demonstrated that the Incoherent Feed
-Forward Loop type 2 (I2-FFL) is able to generate a striped gene expression
-pattern,~\cite{Schaerli2014, Basu2005}, in our study this motif was present in
+Although it has been experimentally demonstrated that the Incoherent Feed-Forward
+Loop type 2 (I2-FFL) can generate a striped gene expression
+pattern~\cite{Schaerli2014, Basu2005}, in our study, this motif was present in
 only 10.92\% of the topologies. This fact may be because evolutionary algorithms
-can not always find global minima but local minima or its surroundings, however,
+can not always find global minima but local minima or their surroundings. However,
 it is also possible that the I2-FFL motif is not as robust as other motifs and
@@
 indicates that networks with simpler designs are more robust to changes in
-interaction parameters and that more complex networks are less robusts. For
-example, topology number 420 (Fig~\ref{fig:2neut-net}), which
+interaction parameters and that more complex networks are less robust. For
+example, topology number 420~(Fig~\ref{fig:2neut-net}), which
 presents high complexity, was one of the
@@
 The cladogram of spatiotemporal expression profiles shows that the search space
-algorithm selected similar expression mechanisms that can be produced by very
-different topologies.
+algorithm selected similar expression mechanisms that very
+different topologies can produce.
 
 Although GRNs that produce a band of gene expression with only two nodes have
-been described (e.g. the I-zero motif,~\cite{Schaerli2014}) this is to our
-knowledge the first time a GRN with a topology such as that of topology nine
-(Fig \ref{fig:topol9})
-and its expression mechanism are reported, increasing our knowledge of the
-genotype-phenotype map in this kind of system, and it would be very interesting
+been described (e.g., the I-zero motif~\cite{Schaerli2014}), this is to our
+knowledge the first time a GRN with a topology such as that of topology
+nine~(Fig \ref{fig:topol9})
+and its expression mechanism is reported, increasing our knowledge of the
+genotype-phenotype map in this kind of system. It would be very interesting
 to confirm experimentally the ability of this topology to generate the described
@@
 and theoretical analysis and provides a foundation to design and build synthetic
-regulatory networks or to identify new ones existing in nature.
+regulatory networks or identify new ones existing in nature.
 
 %\section*{Conclusion}
-
-%Aliquam tincidunt, ex in facilisis elementum, libero lectus luctus est, non vulputate nisl augue at dolor. For more information, see \nameref{S1_Appendix}.
 
@@
 services used in this work were provided by the VSC (Flemish Supercomputer
-  Center), funded by Ghent University, FWO and the Flemish Government –
-  department EWI.
+Center), funded by Ghent University, FWO and the Flemish Government–department EWI.
+All authors acknowledge partial support from CoDI-Universidad de Antioquia under
+project "Estudio de la Evolución de Circuitos de Regulación del Desarollo Embrionario a través de
+la Biología Evolutiva de Sistemas" code 2017-14367. All authors acknowledge the Colombian
+youth for their defense of a better country.
 

### 2021-07-10 docs/manuscript/manuscript.tex
Update manuscript.tex

@@
 Unlike most modeling studies published so far, here we explore the
-space of GRN topologies with nodes having the potential to perceive differently
-the same input signal. This allows for a lot more flexibility during the search
+space of GRN topologies with nodes having the potential to perceive
+the same input signal differently. This allows for a lot more flexibility during the search
 space process, and thus enables us to identify a larger set of potentially

### 2021-07-11 docs/manuscript/manuscript.tex
remove trailing whitespaces

@@
  The filters mentioned above are integrated in the following functions that
- evaluates the  quality of a given phenotype:
+ evaluate the quality of a given phenotype:
 
@@
 
+\paragraph*{S1 Appendix.}
+\label{S1_Appendix}
+{\bf Diffusion independent networks and robustness to changes in the signal input.}
+
 \paragraph*{S1 Table.}

### 2021-07-11 docs/manuscript/manuscript.tex
Reformat and add citations

@@
 expression~\cite{atkinson_development_2003}, counting
-devices~\cite{friedland_synthetic_2009} and systems that respond to light.
+devices~\cite{friedland_synthetic_2009} and systems that respond to
+light~\cite{gardner2012,levskaya2005}.
 Moreover, the design and implementation of GRNs through synthetic biology is
@@
 nature~\cite{vonDassow2000,Reinitz1995,Jaeger2004}, synthetic
-implementations~\cite{Basu2005,Schaerli2018,Elowitz2000,barbier2020} with
-predefined topologies, or have deliberately constrained the study of GRNs to just
+implementations with predefined topologies~\cite{Basu2005, Schaerli2018,
+Elowitz2000, barbier2020}, or have deliberately constrained the study of GRNs to just
 a handful of alternative designs~\cite{munteanu_2014,Schaerli2014}. However,
@@
 signal in varying ways and selected the best-performing GRNs based on a
-fitness criterion used to assess the quality of the phenotype to
+fitness criterion used to assess the quality of the phenotype with respect to
 a prescribed optimal pattern. Based on this computational strategy, we uncovered
@@
 concentration gradient (Fig~\ref{fig:model}C). The initial concentration of
-each gene product in all cells was set to 1 in all simulations.\\
+each gene product in all cells was set to 0.1 in all simulations.\\
 
@@
 initial concentrations randomly from a Normal Distribution with $\mu = 1$ for
-$A_0$, $\mu = 0.4$ for $h$ and $\mu = 1$ for initial concentrations and a
+$A_0$, $\mu = 0.4$ for $h$ and $\mu = 0.1$ for initial concentrations and a
 Coefficient of Variation of 30\% in each case.

### 2021-07-11 docs/manuscript/manuscript.tex
Rephrase text

@@
 compared to other topologies, such as the
-I3-FFL; however, and despite the fact that the Gierer-Meinhardt model form
+I3-FFL; however, and despite the fact that the Gierer-Meinhardt model forms
 patterns that scale with tissue size~\cite{shoaf1984}, we found that topology 9
@@
 
-These two events conduce to the repression of the gene C by the gene A in the
+These two events lead to the repression of the gene C by the gene A in the
 rightmost part of the morphogenetic field and by the morphogen in the leftmost
@@
 
-This is a complex mechanism remindful of the Gierer-Meinhardt model, but
-described by the first time in this study in the context of multiple morphogen
-inputs, and it would be interesting to study its feasibility and possible
-implementation in synthetic gene regulatory networks.
+This is a complex mechanism reminiscent of the Gierer-Meinhardt model, which
+to the best of our knowledge is reported for the first time in the context of
+multiple morphogen inputs, and could thus be of particular interest to for
+synthetic biology implementations to assess the feasibility and robustness
+conferred by this mechanism to GRNs.
 
@@
 morphogenetic field. Our findings shed new light on regulatory rules and
-dynamical mechanisms that  GRNs observed in nature could potentially
-implement to achieve fundamental spatio-temporal tasks, which might
-guide for future synthetic biology implementations.
+dynamical mechanisms that GRNs in nature could potentially implement to
+achieve stereotyped spatio-temporal tasks.
 
@@
 each topology the ability to achieve specific regulatory tasks in a semi-autonomous
-manner. However, we also found a great diversity of network
+manner. In particular, we uncovered a great variety of network
 topologies that could produce a striped gene expression pattern, where 615 out of
-714 were multi-input topologies, showing that considering multiple
-morphogen inputs allows finding a different solution for the patterning
-problem. Most importantly, we found that this ensemble of pattern-forming GRNs
+714 were found to be multi-input topologies, which highlights the importance of
+often disregarded pleiotropic signaling events as a potentially critical design
+specification rule for successful synthetic biology implementations of GRNs
+with various types of pattern-forming abilities. Most importantly, we found that
+this ensemble of pattern-forming GRNs
 tends to form a highly connected meta-graph, which highlights emergent
@@
 in the neutral network connect distinct classes of pattern-forming
-GRNs~(\nameref{S6_Fig}), suggesting that an organism could switch
-between different GRNs (for example, between a Gierer-Meinhardt-like network
-to an Incoherent Feed-Forward Loop) with minimal fitness costs. This high-level
+GRNs~(\nameref{S6_Fig}), suggesting that evolutionary transitions between
+different GRNs (for example, between a Gierer-Meinhardt-like network to an
+Incoherent Feed-Forward Loop) could be achieved with minimal fitness costs. This high-level
 property of the neutral network of topologies could thus facilitate the fine-tuning
@@
 we found that the I3-FFL is one of the network motifs that appears more
-frequently in these GRNs, as has also been reported by Cotterell \&
+frequently in these GRNs, which is in agreement with the study by Cotterell \&
 Sharpe~\cite{Cotterell2010}, and has been found in transcriptional networks involved

### 2021-07-11 docs/manuscript/manuscript.tex
Rephrase text

@@
 However, as our model was less stringent and allowed the morphogen input to
-interact with any gene, we found a more significant number of possible topologies with the
-same striped phenotype and a diverse combination of diffusion and degradation
-parameters~(\nameref{S7_Fig}). For example, in a single input model, the I4-FFL could
-not produce a stripe of gene expression~\cite{munteanu_2014}, whereas, in our
-experimental setup, 20.17\% of the topologies presented this motif as a
-subgraph~(\nameref{S1_Table}).
-
-Although it has been experimentally demonstrated that the Incoherent Feed-Forward
-Loop type 2 (I2-FFL) can generate a striped gene expression
-pattern~\cite{Schaerli2014, Basu2005}, in our study, this motif was present in
-only 10.92\% of the topologies. This fact may be because evolutionary algorithms
-can not always find global minima but local minima or their surroundings. However,
-it is also possible that the I2-FFL motif is not as robust as other motifs and
-therefore be less represented.
+interact with any gene, we found a significantly larger number of pattern-forming
+topologies, which can operate under disparate combinations of diffusion and degradation
+parameters~(\nameref{S7_Fig}). For example, in a single input GRN model, the I4-FFL
+has not been found capable of generating a striped gene expression
+pattern~\cite{munteanu_2014}, whereas, in our experimental setup, 20.17\% of
+the topologies were found  to implement such a regulatory motif~(\nameref{S1_Table}).
+
+Although recent experimental evidence highlights the importance of the Incoherent
+Feed-Forward Loop type 2 (I2-FFL) for generating a striped gene expression
+pattern~\cite{Schaerli2014, Basu2005}, it is intriguing to find that such motif is
+underrepresented in our study, with only 10.92\% of the topologies implementing
+such regulatory mechanism. This might be due to the difficulty to find optimal
+solutions when probing the design space of multi-input GRNs, which are likely to
+contain multiple locally optimal solutions. In addition, it is important to mention
+here that the application of a MCMC-like algorithm as a search space strategy for
+such complex tasks cannot be expected to allow for extensive exploration of the
+design space of GRNs nor for the accessibility of every single locally optimal
+solution. However, it is also possible that the I2-FFL motif is not as robust as
+other motifs and therefore be less represented.
 
@@
 of two of the genes in the GRN, and our model does not take constitutive
-expression into account, we realized that the GRNs that presented the “opposing
-gradients” topology, had also commonly positive auto-regulation interactions on
-these two nodes, supplying in this way the lack of constitutive
+expression into account, we observed that the GRNs that implement the “opposing
+gradients” mechanism also tend to implement auto-regulation interactions on
+these two nodes, bypassing in this way the lack of constitutive
 expression as found in previous studies~\cite{munteanu_2014}.
@@
 been described (e.g., the I-zero motif~\cite{Schaerli2014}), this is to our
-knowledge the first time a GRN with a topology such as that of topology
-nine~(Fig \ref{fig:topol9})
-and its expression mechanism is reported, increasing our knowledge of the
-genotype-phenotype map in this kind of system. It would be very interesting
-to confirm experimentally the ability of this topology to generate the described
-pattern and it would be interesting to study its feasibility and possible
-implementation in synthetic gene regulatory networks.
-
-This work provides a catalog of GRNs that can be extended in future computational
-and theoretical analysis and provides a foundation to design and build synthetic
-regulatory networks or identify new ones existing in nature.
+knowledge the first time that a novel GRN topology such as number
+9~(Fig \ref{fig:topol9}) has been reported, which adds considerably to existing
+knowledge on the genotype-phenotype mapping problems studied in the
+context of developmental pattern formation. It would be interesting to assess
+experimentally the ability of this particular topology to generate striped gene
+expression patterns, as well as to assess its robustness in the face of mutational
+changes and noisy morphogen input profiles.
+
+Overall, we believe our work provides a set of enticing hypotheses on GRN designs,
+represented as a large catalog of distinct GRN topologies that could provide
+valuable information as starting points for future computational and theoretical
+analysis of the genotype-phenotype map, as well as for experimental validation and
+future discovery of potentially interesting GRN designs.
 

### 2021-07-11 docs/manuscript/manuscript.tex
update manuscript
fix quotation marks

@@
 
-The cladogram of spatiotemporal expression profiles shows that the search space
-algorithm selected similar expression mechanisms that very
-different topologies can produce.
-
 Although GRNs that produce a band of gene expression with only two nodes have
@@
 All authors acknowledge partial support from CoDI-Universidad de Antioquia under
-project "Estudio de la Evolución de Circuitos de Regulación del Desarollo Embrionario a través de
-la Biología Evolutiva de Sistemas" code 2017-14367. All authors acknowledge the Colombian
+project ``Estudio de la Evolución de Circuitos de Regulación del Desarollo Embrionario a través de
+la Biología Evolutiva de Sistemas'' code 2017-14367. All authors acknowledge the Colombian
 youth for their defense of a better country.

### 2021-10-10 docs/manuscript/manuscript.tex
Remove comments from .tex file
Comment \includegraphics to remove figures
Change index from 3 to 1
Response to Reviewer 3 point 6
Clarify Shannon entropy.

@@
-% Template for PLoS
-% Version 3.4 January 2017
-%
-% % % % % % % % % % % % % % % % % % % % % %
-%
-% -- IMPORTANT NOTE
-%
-% This template contains comments intended
-% to minimize problems and delays during our production
-% process. Please follow the template instructions
-% whenever possible.
-%
-% % % % % % % % % % % % % % % % % % % % % % %
-%
-% Once your paper is accepted for publication,
-% PLEASE REMOVE ALL TRACKED CHANGES in this file
-% and leave only the final text of your manuscript.
-% PLOS recommends the use of latexdiff to track changes during review, as this will help to maintain a clean tex file.
-% Visit https://www.ctan.org/pkg/latexdiff?lang=en for info or contact us at latex@plos.org.
-%
-%
-% There are no restrictions on package use within the LaTeX files except that
-% no packages listed in the template may be deleted.
-%
-% Please do not include colors or graphics in the text.
-%
-% The manuscript LaTeX source should be contained within a single file (do not use \input, \externaldocument, or similar commands).
-%
-% % % % % % % % % % % % % % % % % % % % % % %
-%
-% -- FIGURES AND TABLES
-%
-% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
-%
-% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
-% - Figures should be uploaded separately from your manuscript file.
-% - Figures generated using LaTeX should be extracted and removed from the PDF before submission.
-% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
-% For figure citations, please use "Fig" instead of "Figure".
-% See http://journals.plos.org/plosone/s/figures for PLOS figure guidelines.
-%
-% Tables should be cell-based and may not contain:
-% - spacing/line breaks within cells to alter layout or alignment
-% - do not nest tabular environments (no tabular environments within tabular environments)
-% - no graphics or colored text (cell background color/shading OK)
-% See http://journals.plos.org/plosone/s/tables for table guidelines.
-%
-% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
-%
-% % % % % % % % % % % % % % % % % % % % % % % %
-%
-% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
-%
-% IMPORTANT
-% Below are a few tips to help format your equations and other special characters according to our specifications.
-For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://journals.plos.org/plosone/s/latex
-%
-% For inline equations, please be sure to include all portions of an equation in the math environment.
-For example, x$^2$ is incorrect; this should be formatted as $x^2$ (or $\mathrm{x}^2$ if the romanized font is desired).
-%
-% Do not include text that is not math in the math environment. For example, CO2 should be written as CO\textsubscript{2} instead of CO$_2$.
-%
-% Please add line breaks to long display equations when possible in order to fit size of the column.
-%
-% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
-%
-% When adding superscript or subscripts outside of brackets/braces, please group using {}.  For example, change "[U(D,E,\gamma)]^2" to "{[U(D,E,\gamma)]}^2".
-%
-% Do not use \cal for caligraphic font.  Instead, use \mathcal{}
-%
-% % % % % % % % % % % % % % % % % % % % % % % %
-%
-% Please contact latex@plos.org with any questions.
-%
-% % % % % % % % % % % % % % % % % % % % % % % %
-
 \documentclass[10pt,letterpaper]{article}
@@
 Jayson Gutiérrez\textsuperscript{3},
-% Name5 Surname\textsuperscript{2\ddag},
-% Name6 Surname\textsuperscript{2\ddag},
-% Name7 Surname\textsuperscript{1,2,3*},
-% with the Lorem Ipsum Consortium\textsuperscript{\textpilcrow}
 \\
@@
 Belgium
-% \textbf{3} Affiliation Dept/Program/Center, Institution Name, City, State, Country
 \\
 \bigskip
-
-% Insert additional author notes using the symbols described below. Insert symbol callouts after author names as necessary.
-%
-% Remove or comment out the author notes below if they aren't used.
-%
-% Primary Equal Contribution Note
-% \Yinyang These authors contributed equally to this work.
-
-% Additional Equal Contribution Note
-% Also use this double-dagger symbol for special authorship notes, such as senior authorship.
-% \ddag These authors also contributed equally to this work.
-
-% Current address notes
-% \textcurrency Current Address: Dept/Program/Center, Universidad de Antioquia,
-% Medellín, Antioquia, Colombia % change symbol to "\textcurrency a" if more than one current address note
-% \textcurrency b Insert second current address
-% \textcurrency c Insert third current address
-
-% Deceased author note
-% \dag Deceased
-
-% Group/Consortium Author Note
-% \textpilcrow Membership list can be found in the Acknowledgments section.
 
@@
 
-% Equation provided by the LaTeX template.
-% \begin{eqnarray}
-% \label{eq:schemeP}
-% 	\mathrm{P_Y} = \underbrace{H(Y_n) - H(Y_n|\mathbf{V}^{Y}_{n})}_{S_Y} + \underbrace{H(Y_n|\mathbf{V}^{Y}_{n})- H(Y_n|\mathbf{V}^{X,Y}_{n})}_{T_{X\rightarrow Y}},
-% \end{eqnarray}
-
 \section*{Materials and methods}
@@
 %  \centering
- \includegraphics[width=\textwidth]{figures/metodos/Fig1}
+%  \includegraphics[width=\textwidth]{figures/metodos/Fig1}
     \caption{\bf Modelling of GRNs.}
@@
  \begin{equation}
-  P_{filter} = \left( \frac{1}{I \cdot N} \right) \sqrt{ \sum_{i=3}^{I=3}
+  P_{filter} = \left( \frac{1}{I \cdot N} \right) \sqrt{ \sum_{i=1}^{I=3}
   \sum_{n=1}^{N=30} \left( [G]^i_n - \langle [G]^i_n \rangle \right)^2 },
@@
 
-% For placing video, \nameref{S1_Video} vel sagittis arcu lobortis.
-
-% Place figure captions after the first paragraph in which they are cited.
-% \begin{figure}[!h]
-% \caption{{\bf Bold the figure title.}
-% Figure caption text here, please use this space for the figure panel descriptions instead of using subfigure commands. A: Lorem ipsum dolor sit amet. B: Consectetur adipiscing elit.}
-% \label{fig1}
-% \end{figure}
-
 % Results and Discussion can be combined.
@@
 variable number of GRNs (Fig~\ref{fig:distopol}). This abundances
-distribution was not produced by chance (Shannon entropy = 8.52963, 95\%
-confidence interval = [10.8915, 10.9275]), and instead it displays some level
-of structure, indicating the existence of network motifs.
+distribution was not produced by chance as its Shannon entropy was
+significantly lower than that of random networks produced without
+selection involved (Shannon entropy = 8.52963, 95\%
+confidence interval for random networks = [10.8915, 10.9275]), and
+instead it displays some level of structure, indicating the existence
+of network motifs.
 
 \begin{figure}[!h]
- \includegraphics[width=\textwidth]{figures/results/Fig2}
+%  \includegraphics[width=\textwidth]{figures/results/Fig2}
  \caption{\bf Distribution of abundances of network topologies.}
@@
 \begin{figure}[!h]
- \includegraphics[width=\textwidth]{figures/results/Fig3}
+%  \includegraphics[width=\textwidth]{figures/results/Fig3}
  \caption{\bf Neutral network.}
@@
 \begin{figure}[!h]
- \includegraphics[width=\textwidth]{figures/results/Fig4}
+%  \includegraphics[width=\textwidth]{figures/results/Fig4}
  \caption{\bf Node degree in the neutral network.}
@@
 \begin{figure}
- \includegraphics[width=\textwidth]{figures/results/Fig5}
+%  \includegraphics[width=\textwidth]{figures/results/Fig5}
  \caption{\bf Network of topologies separated by two mutational steps.}
@@
 \begin{figure}[!h]
- \includegraphics[width=\textwidth]{figures/results/Fig6}
+%  \includegraphics[width=\textwidth]{figures/results/Fig6}
  \caption{\bf Relationship between Abundance and robustness and complexity
@@
 \begin{figure}[!h]
- \includegraphics[width=\textwidth]{figures/results/Fig7}
+%  \includegraphics[width=\textwidth]{figures/results/Fig7}
  \caption{\bf Clustering of expression profiles and topologies.}
@@
 \begin{figure}[!h]
- \includegraphics[width=\textwidth]{figures/results/Fig8}
+%  \includegraphics[width=\textwidth]{figures/results/Fig8}
  \caption{\bf Expression dynamics of topology 9.}
@@
 (\nameref{S5_Fig}).
-% (Fig~\ref{fig:topol9-dyn}).
-
-% \begin{figure}[!h]
-%  \includegraphics[width=\textwidth]{figures/results/topol-9-dynamics}
-%  \caption{\bf Dynamics of expression in topology number 9.}
-%  \label{fig:topol9-dyn}
-% \end{figure}
 
@@
 different GRNs (for example, between a Gierer-Meinhardt-like network to an
-Incoherent Feed-Forward Loop) could be achieved with minimal fitness costs. This high-level
-property of the neutral network of topologies could thus facilitate the fine-tuning
-of the regulatory systems through genetic modifications to meet
-specific functional requirements under selective pressures.
+Incoherent Feed-Forward Loop) could be achieved with minimal fitness
+costs. This high-level property of the neutral network of topologies could thus
+facilitate the fine-tuning of the regulatory systems through genetic modifications
+to meet specific functional requirements under selective pressures.
 
@@
 All authors acknowledge partial support from CoDI-Universidad de Antioquia under
-project ``Estudio de la Evolución de Circuitos de Regulación del Desarollo Embrionario a través de
-la Biología Evolutiva de Sistemas'' code 2017-14367. All authors acknowledge the Colombian
-youth for their defense of a better country.
+project ``Estudio de la Evolución de Circuitos de Regulación del Desarollo
+Embrionario a través de la Biología Evolutiva de Sistemas'' code 2017-14367. All
+authors acknowledge the Colombian youth for their defense of a better country.
 
@@
 
-% Either type in your references using
-% \begin{thebibliography}{}
-% \bibitem{}
-% Text
-% \end{thebibliography}
-%
-% or
-%
-% Compile your BiBTeX database using our plos2015.bst
-% style file and paste the contents of your .bbl file
-% here. See http://journals.plos.org/plosone/s/latex for
-% step-by-step instructions.
-%
-% \begin{thebibliography}{10}
-%
-% \bibitem{bib1}
-% Conant GC, Wolfe KH.
-% \newblock {{T}urning a hobby into a job: how duplicated genes find new
-%   functions}.
-% \newblock Nat Rev Genet. 2008 Dec;9(12):938--950.
-%
-% \bibitem{bib2}
-% Ohno S.
-% \newblock Evolution by gene duplication.
-% \newblock London: George Alien \& Unwin Ltd. Berlin, Heidelberg and New York:
-%   Springer-Verlag.; 1970.
-%
-% \bibitem{bib3}
-% Magwire MM, Bayer F, Webster CL, Cao C, Jiggins FM.
-% \newblock {{S}uccessive increases in the resistance of {D}rosophila to viral
-%   infection through a transposon insertion followed by a {D}uplication}.
-% \newblock PLoS Genet. 2011 Oct;7(10):e1002337.
-%
-% \end{thebibliography}
-
 \bibliography{referencias-tesis}{}

### 2021-10-17 docs/manuscript/manuscript.tex
Improve Fig 1
Improve description of isomorph identification

@@
     colors represent lower expression and red colors represent higher
-    expression.
+    expression. Time axis goes from 0 to 500 integration steps. Horizontal
+    axis represents the morphogenetic field, i. e. cells from 1 to 30 for
+    each gene.
  \label{fig:model}
@@
 the totality (2061) of sampled GRNs we used a
-classification algorithm that selected a random GRN and generated all
+classification algorithm that selected a GRN and generated all
 isomorphs of that GRN by interchanging rows of the adjacency matrix
 ($i \leftrightarrow j$) while interchanging the corresponding columns
-($i+1 \leftrightarrow j+1$). The
-algorithm then compared all six isomorphs with other GRN and classified the
-topologies in the same group if they were isomorphic networks.
+($i+1 \leftrightarrow j+1$). This kind of transformation of the adjacency
+matrix results in isospectral graphs that are also isomorphic to the
+original graph and is comparable to generate a permutation matrix $P$ and
+multiplying $PMP^{T}$, where $M$ is the adjacency matrix and $P^{T}$ the
+transpose of $P$. The algorithm then compared these isomorphs with another
+GRN and classified the topologies in the same group if at least one of the
+transformed adjacency matrices (or the original one) was identical to that
+of the other GRN, i. e., if they were isomorphic networks.
 

### 2021-10-17 docs/manuscript/manuscript.tex
Improve S5 Fig. caption

@@
 {\bf Dynamics of expression of topology number 9.}
-The dynamics of expression is shown from $t = 5$ to $t = 150$. The red line
+The dynamics of expression is shown from $t = 5$ to $t = 150$, where this
+topology reaches its steady state expression. The red line
 represents the expression level of the gene C along the morphogenetic field,

### 2021-10-18 docs/manuscript/manuscript.tex
Correct statement of scaling of Turing mechanisms

@@
 I3-FFL; however, and despite the fact that the Gierer-Meinhardt model forms
-patterns that scale with tissue size~\cite{shoaf1984}, we found that topology 9
+patterns that can scale with tissue size in the presence of external
+forcing~\cite{shoaf1984, ishihara_2006}, we found that topology 9
 was more sensible to changes in the morphogenetic field size than those

### 2021-10-20 docs/manuscript/manuscript.tex
Explain multi-input morphogen concentration
Improve mutation description

@@
 In our model the morphogen could interact with any of the three genes on the
-GRN, but could not be affected by them. The morphogenetic field was defined by
-an unidimensional array of 30 isogenic cells exposed to the morphogen
-concentration gradient (Fig~\ref{fig:model}C). The initial concentration of
-each gene product in all cells was set to 0.1 in all simulations.\\
+GRN, but could not be affected by them. The morphogenetic field was defined
+by an unidimensional array of 30 isogenic cells exposed to the morphogen
+concentration gradient (Fig~\ref{fig:model}C). In multiple-input networks,
+all the genes of a given cell are exposed to the same morphogen
+concentration, given by an exponential decay function (see Morphogen
+spatial distribution). The initial concentration of each gene product in
+all cells was set to 0.1 in all simulations.\\
 
@@
 
- \item {\bf Mutate GRNs randomly and evaluate fitness again.} Mutations
-included changes in the  numerical values of degradation and diffusion
-parameters and interaction strengths, and addition and removal of interactions
-among genes and between the genes and the morphogen. If fitness of the mutated
-GRN was greater or equal than that of the previous GRN, mutated GRN was saved
-and the cycle continued from step 2. This cycle was repeated 50000 times, and a GRN
-sampled at a given step and with a fitness score ≥ 0.95 was saved to a file as
+ \item {\bf Mutate GRNs randomly and evaluate fitness again.} Three types
+ of mutations were possible in our algorithm:
+1) changes in the numerical value of the degradation parameter, 2) changes
+in the numerical value of the diffusion parameter and 3) changes in the
+numerical values of the interaction strengths. The type of mutation to be
+performed was selected randomly with a probability of 0.15 for type 1 and
+type 2 mutations and 0.7 for type 3 mutations. If type 3 mutation was
+selected, either the interaction strength was set to zero (with a
+probability of 0.2) or was changed randomly (with a probability of 0.8).
+If fitness of the mutated GRN was greater or equal than that of the
+previous GRN, mutated GRN was saved and the cycle continued from step 2.
+This cycle was repeated 50000 times, and a GRN sampled at a given step
+and with a fitness score ≥ 0.95 was saved to a file as
 long as its associated topology had not been sampled in previous steps.

### 2021-10-23 docs/manuscript/manuscript.tex
Improve captions
Improve Fig. 6 caption
Include S1 Appendix in main text
Improve S3 Fig caption

@@
 %  \includegraphics[width=\textwidth]{figures/results/Fig6}
- \caption{\bf Relationship between Abundance and robustness and complexity
- index.} (A) Abundance vs. Robustness. The robustness of each topology was
+ \caption{\bf Relationship between the abundance of topologies with
+ robustness and complexity index.}
+ (A) Abundance vs. Robustness. The robustness of each topology was
  calculated as the ratio of perturbations in which the fitness was greater or
@@
 only 4.66\%, 56.82\%, 56.14\% and 41.05\%, respectively, of the examined
-GRNs showed a fitness greater than 0.9 (\nameref{S3_Fig}B).
+GRNs showed a fitness greater than 0.9 (\nameref{S3_Fig}B). In addition to
+robustness, we wanted to investigate which of the GRNs could still produce
+a striped pattern in absence of diffusion, as it has been shown that some
+networks can work without diffusion involved \cite{Schaerli2018}. We found
+that out of 2061 GRNs, 404 were diffusion independent, many of them
+presenting the I3-FFL (see \nameref{S1_Appendix}).
 
@@
 {\bf Response of GRNs to changes in morphogenetic field size.}
-(A) Morphogen gradients of fields with 10, 20, 40 and 50 cells. (B) Fitness of the
+(A) Morphogen gradients of fields with 10, 20, 40 and 50 cells. These
+morphogen gradients were set up so that the morphogen concentration in the
+first cell and in the last cell were constant and only one
+morphogen gradient was considered at a time depending on the number of
+cells in the morphogenetic field. (B) Fitness of the
 GRNs by topology in each one of the morphogenetic fields. (C) Principal Component

### 2021-11-01 docs/manuscript/manuscript.tex
Small improvements and fix typos

@@
 If fitness of the mutated GRN was greater or equal than that of the
-previous GRN, mutated GRN was saved and the cycle continued from step 2.
+previous GRN, the mutated GRN was saved and the cycle continued from step 2.
 This cycle was repeated 50000 times, and a GRN sampled at a given step
@@
 matrix results in isospectral graphs that are also isomorphic to the
-original graph and is comparable to generate a permutation matrix $P$ and
+original graph and is equivalent to generate a permutation matrix $P$ and
 multiplying $PMP^{T}$, where $M$ is the adjacency matrix and $P^{T}$ the
@@
 morphogen gradient in such a way that the initial and final concentration were
-constant in all the fields (\nameref{S3_Fig}A). In order to know if topologies
+constant in all the fields (\nameref{S3_Fig}A). In order to know if GRNs
 with the same topology had a similar fitness in the different fields, we performed
@@
  Topology 1 the most abundant with a number of 95 genotypes. The number of
- genotypes belonging to each network topology is presented below
+ genotypes belonging to each network topology is presented below the
  topology number.
@@
 found nodes inside the neutral network that had more connections between them
-than with other nodes, which leads to further clustering of nodes. Interestingly in
+than with other nodes, which leads to further clustering of nodes. Interestingly, in
 almost every cluster one can find one or more of the 15 most abundant network
@@
 
-The degree distribution of this network seems to follow a Poisson distribution
+The degree distribution of this neutral network seems to follow a Poisson distribution
 (Pearson’s $\chi^2 = 12.8902$, $\text{p-value} = 0.115684$) characteristic of a
@@
 %  \includegraphics[width=\textwidth]{figures/results/Fig5}
- \caption{\bf Network of topologies separated by two mutational steps.}
+ \caption{\bf Neutral network of topologies separated by two mutational
+ steps.}
  The network shows that almost all the topologies are connected by two or less
@@
 to the best of our knowledge is reported for the first time in the context of
-multiple morphogen inputs, and could thus be of particular interest to for
+multiple morphogen inputs, and could thus be of particular interest for
 synthetic biology implementations to assess the feasibility and robustness
@@
 qualitative and quantitative changes, we found that many GRNs were susceptible
-to changes in morphogenetic field size. However, interestingly, GRNs with the same
+to changes in morphogenetic field size. However, GRNs with the same
 topology behave in different ways for each field size~(\nameref{S3_Fig}C),
@@
 
-However, as our model was less stringent and allowed the morphogen input to
+Moreover, as our model was less stringent and allowed the morphogen input to
 interact with any gene, we found a significantly larger number of pattern-forming
@@
 
-We also found topologies with a similar topology to that of the “opposing
+We also found GRNs with a similar topology to that of the “opposing
 gradients” reported by Schaerli et al.~\cite{Schaerli2018,Schaerli2014}.
@@
 Center), funded by Ghent University, FWO and the Flemish Government–department EWI.
-All authors acknowledge partial support from CoDI-Universidad de Antioquia under
+All authors acknowledge partial support from CODI-Universidad de Antioquia under
 project ``Estudio de la Evolución de Circuitos de Regulación del Desarollo

### 2021-12-03 docs/manuscript/manuscript.tex
Include references in the main .tex

@@
 
-\bibliography{referencias-tesis}{}
+\begin{thebibliography}{10}
+
+\bibitem{Turing1952}
+Turing AM.
+\newblock {The chemical basis of morphogenesis}.
+\newblock Bulletin of Mathematical Biology. 1952;52(1-2):153--197.
+\newblock doi:{10.1007/BF02459572}.
+
+\bibitem{Wolpert1969}
+Wolpert L.
+\newblock {Positional information and the spatial pattern of cellular
+  differentiation}.
+\newblock Journal of Theoretical Biology. 1969;25(1):1--47.
+\newblock doi:{https://doi.org/10.1016/S0022-5193(69)80016-0}.
+
+\bibitem{sharpe_2019}
+Sharpe J.
+\newblock Wolpert's {French} {Flag}: what's the problem?
+\newblock Development (Cambridge, England). 2019;146(24).
+\newblock doi:{10.1242/dev.185967}.
+
+\bibitem{driever_gradient_1988}
+Driever W, Nüsslein-Volhard C.
+\newblock A gradient of bicoid protein in {Drosophila} embryos.
+\newblock Cell. 1988;54(1):83--93.
+\newblock doi:{10.1016/0092-8674(88)90182-1}.
+
+\bibitem{driever_bicoid_1988}
+Driever W, Nüsslein-Volhard C.
+\newblock The bicoid protein determines position in the {Drosophila} embryo in
+  a concentration-dependent manner.
+\newblock Cell. 1988;54(1):95--104.
+\newblock doi:{10.1016/0092-8674(88)90183-3}.
+
+\bibitem{affolter_decapentaplegic_2007}
+Affolter M, Basler K.
+\newblock The {Decapentaplegic} morphogen gradient: from pattern formation to
+  growth regulation.
+\newblock Nature Reviews Genetics. 2007;8(9):663--674.
+\newblock doi:{10.1038/nrg2166}.
+
+\bibitem{dessaud_dynamic_2010}
+Dessaud E, Ribes V, Balaskas N, Yang LL, Pierani A, Kicheva A, et~al.
+\newblock Dynamic {Assignment} and {Maintenance} of {Positional} {Identity} in
+  the {Ventral} {Neural} {Tube} by the {Morphogen} {Sonic} {Hedgehog}.
+\newblock PLOS Biology. 2010;8(6):e1000382.
+\newblock doi:{10.1371/journal.pbio.1000382}.
+
+\bibitem{cohen_theoretical_2014}
+Cohen M, Page KM, Perez-Carrasco R, Barnes CP, Briscoe J.
+\newblock A theoretical framework for the regulation of {Shh}
+  morphogen-controlled gene expression.
+\newblock Development. 2014;141(20):3868--3878.
+\newblock doi:{10.1242/dev.112573}.
+
+\bibitem{raspopovic_digit_2014}
+Raspopovic J, Marcon L, Russo L, Sharpe J.
+\newblock Digit patterning is controlled by a {Bmp}-{Sox9}-{Wnt} {Turing}
+  network modulated by morphogen gradients.
+\newblock Science. 2014;345(6196):566.
+\newblock doi:{10.1126/science.1252960}.
+
+\bibitem{Milo2002}
+Milo R, Shen-Orr SS, Itzkovitz S, Kashtan N, Chklovskii D, Alon U.
+\newblock {Network Motifs: Simple Building Blocks of Complex Networks}.
+\newblock Science. 2002;298(5594):824--827.
+\newblock doi:{10.1126/science.298.5594.824}.
+
+\bibitem{kitano_computational_2002}
+Kitano H.
+\newblock Computational systems biology.
+\newblock Nature. 2002;420(6912):206--210.
+\newblock doi:{10.1038/nature01254}.
+
+\bibitem{Kalir2005}
+Kalir S, Mangan S, Alon U. {A coherent feed-forward loop with a SUM input
+  function prolongs flagella expression in Escherichia coli}; 2005.
+
+\bibitem{Kalir2004}
+Kalir S, Alon U.
+\newblock {Using a quantitative blueprint to reprogram the dynamics of the
+  flagella gene network}.
+\newblock Cell. 2004;117(6):713--720.
+\newblock doi:{10.1016/j.cell.2004.05.010}.
+
+\bibitem{Mangan2003a}
+Mangan S, Zaslaver A, Alon U.
+\newblock {The coherent feedforward loop serves as a sign-sensitive delay
+  element in transcription networks.}
+\newblock Journal of molecular biology. 2003;334(2):197--204.
+
+\bibitem{ODonnell2005}
+O'Donnell KA, Wentzel EA, Zeller KI, Dang CV, Mendell JT.
+\newblock {c-Myc-regulated microRNAs modulate E2F1 expression.}
+\newblock Nature. 2005;435(7043):839--843.
+\newblock doi:{10.1038/nature03677}.
+
+\bibitem{Basu2005}
+Basu S, Gerchman Y, Collins CH, Arnold FH, Weiss R.
+\newblock {A synthetic multicellular system for programmed pattern formation}.
+\newblock Nature. 2005;434(7037):1130--1134.
+\newblock doi:{10.1038/nature03461}.
+
+\bibitem{atkinson_development_2003}
+Atkinson MR, Savageau MA, Myers JT, Ninfa AJ.
+\newblock Development of genetic circuitry exhibiting toggle switch or
+  oscillatory behavior in {Escherichia} coli.
+\newblock Cell. 2003;113(5):597--607.
+\newblock doi:{10.1016/s0092-8674(03)00346-5}.
+
+\bibitem{friedland_synthetic_2009}
+Friedland AE, Lu TK, Wang X, Shi D, Church G, Collins JJ.
+\newblock Synthetic gene networks that count.
+\newblock Science (New York, NY). 2009;324(5931):1199--1202.
+\newblock doi:{10.1126/science.1172005}.
+
+\bibitem{gardner2012}
+Gardner L, Deiters A.
+\newblock Light-controlled synthetic gene circuits.
+\newblock Current Opinion in Chemical Biology. 2012;16(3-4):292--299.
+\newblock doi:{10.1016/j.cbpa.2012.04.010}.
+
+\bibitem{levskaya2005}
+Levskaya A, Chevalier AA, Tabor JJ, Simpson ZB, Lavery LA, Levy M, et~al.
+\newblock Engineering {Escherichia} coli to see light.
+\newblock Nature. 2005;438(7067):441--442.
+\newblock doi:{10.1038/nature04405}.
+
+\bibitem{santos-moreno_using_2019}
+Santos-Moreno J, Schaerli Y.
+\newblock Using {Synthetic} {Biology} to {Engineer} {Spatial} {Patterns}.
+\newblock Advanced Biosystems. 2019;3(4):1800280.
+\newblock doi:{https://doi.org/10.1002/adbi.201800280}.
+
+\bibitem{karlsson_therapeutic_2012}
+Karlsson M, Weber W.
+\newblock Therapeutic synthetic gene networks.
+\newblock Current Opinion in Biotechnology. 2012;23(5):703--711.
+\newblock doi:{10.1016/j.copbio.2012.01.003}.
+
+\bibitem{higashikuni_advancing_2017}
+Higashikuni Y, Chen WC, Lu TK.
+\newblock Advancing therapeutic applications of synthetic gene circuits.
+\newblock Current Opinion in Biotechnology. 2017;47:133--141.
+\newblock doi:{10.1016/j.copbio.2017.06.011}.
+
+\bibitem{abil_synthetic_2015}
+Abil Z, Xiong X, Zhao H.
+\newblock Synthetic biology for therapeutic applications.
+\newblock Molecular Pharmaceutics. 2015;12(2):322--331.
+\newblock doi:{10.1021/mp500392q}.
+
+\bibitem{healy_genetic_2019}
+Healy CP, Deans TL.
+\newblock Genetic circuits to engineer tissues with alternative functions.
+\newblock Journal of Biological Engineering. 2019;13:39.
+\newblock doi:{10.1186/s13036-019-0170-7}.
+
+\bibitem{kitada_programming_2018}
+Kitada T, DiAndreth B, Teague B, Weiss R.
+\newblock Programming gene and engineered-cell therapies with synthetic
+  biology.
+\newblock Science (New York, NY). 2018;359(6376).
+\newblock doi:{10.1126/science.aad1067}.
+
+\bibitem{munteanu_2014}
+Munteanu A, Cotterell J, Solé RV, Sharpe J.
+\newblock Design principles of stripe-forming motifs: the role of positive
+  feedback.
+\newblock Scientific Reports. 2014;4(1):5003.
+\newblock doi:{10.1038/srep05003}.
+
+\bibitem{Cotterell2010}
+Cotterell J, Sharpe J.
+\newblock {An atlas of gene regulatory networks reveals multiple three-gene
+  mechanisms for interpreting morphogen gradients}.
+\newblock Molecular Systems Biology. 2010;6(425):1--14.
+\newblock doi:{10.1038/msb.2010.74}.
+
+\bibitem{vonDassow2000}
+von Dassow G, Meir E, Munro EM, Odell GM.
+\newblock The segment polarity network is a robust developmental module.
+\newblock Nature. 2000;406(6792):188--192.
+\newblock doi:{10.1038/35018085}.
+
+\bibitem{Reinitz1995}
+Reinitz J, Mjolsness E, Sharp DH.
+\newblock {Model for cooperative control of positional information in
+  Drosophila by bicoid and maternal hunchback}.
+\newblock Journal of Experimental Zoology. 1995;271(1):47--56.
+\newblock doi:{10.1002/jez.1402710106}.
+
+\bibitem{Jaeger2004}
+Jaeger J, Blagov M, Kosman D, Kozlov KN, Manu, Myasnikova E, et~al.
+\newblock {Dynamical analysis of regulatory interactions in the gap gene system
+  of Drosophila melanogaster}.
+\newblock Genetics. 2004;167(4):1721--1737.
+\newblock doi:{10.1534/genetics.104.027334}.
+
+\bibitem{Schaerli2018}
+Schaerli Y, Jim{\'{e}}nez A, Duarte JM, Mihajlovic L, Renggli J, Isalan M,
+  et~al.
+\newblock {Synthetic circuits reveal how mechanisms of gene regulatory networks
+  constrain evolution}.
+\newblock Molecular Systems Biology. 2018;14(9):e8102.
+\newblock doi:{10.15252/msb.20178102}.
+
+\bibitem{Elowitz2000}
+Elowitz MB, Leibier S.
+\newblock {A synthetic oscillatory network of transcriptional regulators}.
+\newblock Nature. 2000;403(6767):335--338.
+\newblock doi:{10.1038/35002125}.
+
+\bibitem{barbier2020}
+Barbier I, Perez-Carrasco R, Schaerli Y.
+\newblock Controlling spatiotemporal pattern formation in a concentration
+  gradient with a synthetic toggle switch.
+\newblock Molecular Systems Biology. 2020;16(6):e9361.
+\newblock doi:{10.15252/msb.20199361}.
+
+\bibitem{Schaerli2014}
+Schaerli Y, Munteanu A, Gili M, Cotterell J, Sharpe J, Isalan M.
+\newblock {A unified design space of synthetic stripe-forming networks}.
+\newblock Nature Communications. 2014;5(May):4905.
+\newblock doi:{10.1038/ncomms5905}.
+
+\bibitem{balaskas2012}
+Balaskas N, Ribeiro A, Panovska J, Dessaud E, Sasai N, Page KM, et~al.
+\newblock Gene {Regulatory} {Logic} for {Reading} the {Sonic} {Hedgehog}
+  {Signaling} {Gradient} in the {Vertebrate} {Neural} {Tube}.
+\newblock Cell. 2012;148(1):273--284.
+\newblock doi:{10.1016/j.cell.2011.10.047}.
+
+\bibitem{exelby_2021}
+Exelby K, Herrera-Delgado E, Perez LG, Perez-Carrasco R, Sagner A, Metzis V,
+  et~al.
+\newblock Precision of tissue patterning is controlled by dynamical properties
+  of gene regulatory networks.
+\newblock Development. 2021;148(dev197566).
+\newblock doi:{10.1242/dev.197566}.
+
+\bibitem{ricci2011}
+Ricci F, Vallée-Bélisle A, Plaxco KW.
+\newblock High-precision, in vitro validation of the sequestration mechanism
+  for generating ultrasensitive dose-response curves in regulatory networks.
+\newblock PLoS computational biology. 2011;7(10):e1002171.
+\newblock doi:{10.1371/journal.pcbi.1002171}.
+
+\bibitem{mendoza2006}
+Mendoza L, Xenarios I.
+\newblock A method for the generation of standardized qualitative dynamical
+  systems of regulatory networks.
+\newblock Theoretical Biology \& Medical Modelling. 2006;3:13.
+\newblock doi:{10.1186/1742-4682-3-13}.
+
+\bibitem{vu2007}
+Vu TT, Vohradsky J.
+\newblock Nonlinear differential equation model for quantification of
+  transcriptional regulation applied to microarray data of {Saccharomyces}
+  cerevisiae.
+\newblock Nucleic Acids Research. 2007;35(1):279--287.
+\newblock doi:{10.1093/nar/gkl1001}.
+
+\bibitem{zhang2013}
+Zhang Q, Bhattacharya S, Andersen ME.
+\newblock Ultrasensitive response motifs: basic amplifiers in molecular
+  signalling networks.
+\newblock Open Biology. 2013;3(4):130031.
+\newblock doi:{10.1098/rsob.130031}.
+
+\bibitem{armao2016}
+Armao JJ, Lehn JM.
+\newblock Nonlinear {Kinetic} {Behavior} in {Constitutional} {Dynamic}
+  {Reaction} {Networks}.
+\newblock Journal of the American Chemical Society. 2016;138(51):16809--16814.
+\newblock doi:{10.1021/jacs.6b11107}.
+
+\bibitem{chen2005}
+Chen KC, Wang TY, Tseng HH, Huang CYF, Kao CY.
+\newblock A stochastic differential equation model for quantifying
+  transcriptional regulatory network in {Saccharomyces} cerevisiae.
+\newblock Bioinformatics (Oxford, England). 2005;21(12):2883--2890.
+\newblock doi:{10.1093/bioinformatics/bti415}.
+
+\bibitem{dahlquist2015}
+Dahlquist KD, Fitzpatrick BG, Camacho ET, Entzminger SD, Wanner NC.
+\newblock Parameter {Estimation} for {Gene} {Regulatory} {Networks} from
+  {Microarray} {Data}: {Cold} {Shock} {Response} in {Saccharomyces} cerevisiae.
+\newblock Bulletin of Mathematical Biology. 2015;77(8):1457--1492.
+\newblock doi:{10.1007/s11538-015-0092-6}.
+
+\bibitem{skbio2020}
+scikit-bio~development team T. scikit-bio: A Bioinformatics Library for Data
+  Scientists, Students, and Developers; 2020.
+\newblock Available from: \url{http://scikit-bio.org}.
+
+\bibitem{D.2005}
+Bonchev DG, Rouvray DH.
+\newblock {Complexity in Chemistry, Biology, and Ecology, ser. Mathematical and
+  Computational Chemistry.}
+\newblock Springer; 2005.
+
+\bibitem{sklearn}
+Pedregosa F, Varoquaux G, Gramfort A, Michel V, Thirion B, Grisel O, et~al.
+\newblock Scikit-learn: Machine Learning in {P}ython.
+\newblock Journal of Machine Learning Research. 2011;12:2825--2830.
+
+\bibitem{Erdos1959}
+Erd{\"{o}}s P, R{\'{e}}nyi a.
+\newblock {On random graphs}.
+\newblock Publicationes Mathematicae. 1959;6:290--297.
+\newblock doi:{10.2307/1999405}.
+
+\bibitem{Watts1998}
+Watts DJ, Strogatz SH.
+\newblock {Collective dynamics of 'small-world' networks.}
+\newblock Nature. 1998;393(6684):440--2.
+\newblock doi:{10.1038/30918}.
+
+\bibitem{giererMeinh1972}
+Gierer A, Meinhardt H.
+\newblock A theory of biological pattern formation.
+\newblock Kybernetik. 1972;12(1):30--39.
+\newblock doi:{10.1007/BF00289234}.
+
+\bibitem{meinhardt2000}
+Meinhardt H, Gierer A.
+\newblock Pattern formation by local self-activation and lateral inhibition.
+\newblock BioEssays: News and Reviews in Molecular, Cellular and Developmental
+  Biology. 2000;22(8):753--760.
+\newblock doi:{10.1002/1521-1878(200008)22:8<753::AID-BIES9>3.0.CO;2-Z}.
+
+\bibitem{shoaf1984}
+Shoaf SA, Conway K, Hunt RK.
+\newblock Application of reaction-diffusion models to cell patterning in
+  {Xenopus} retina. {Initiation} of patterns and their biological stability.
+\newblock Journal of Theoretical Biology. 1984;109(3):299--329.
+\newblock doi:{10.1016/s0022-5193(84)80085-5}.
+
+\bibitem{ishihara_2006}
+Ishihara S, Kaneko K.
+\newblock Turing pattern with proportion preservation.
+\newblock Journal of Theoretical Biology. 2006;238(3):683--693.
+\newblock doi:{10.1016/j.jtbi.2005.06.016}.
+
+\bibitem{kitano2004}
+Kitano H.
+\newblock Biological robustness.
+\newblock Nature Reviews Genetics. 2004;5(11):826--837.
+\newblock doi:{10.1038/nrg1471}.
+
+\bibitem{lo_robust_2015}
+Lo WC, Zhou S, Wan FYM, Lander AD, Nie Q.
+\newblock Robust and precise morphogen-mediated patterning: trade-offs,
+  constraints and mechanisms.
+\newblock Journal of The Royal Society Interface. 2015;12(102):20141041.
+\newblock doi:{10.1098/rsif.2014.1041}.
+
+\bibitem{wagner_robustness_2008}
+Wagner A.
+\newblock Robustness and evolvability: a paradox resolved.
+\newblock Proceedings of the Royal Society B: Biological Sciences.
+  2008;275(1630):91--100.
+\newblock doi:{10.1098/rspb.2007.1137}.
+
+\bibitem{ishihara_cross_2005}
+Ishihara S, Fujimoto K, Shibata T.
+\newblock Cross talking of network motifs in gene regulation that generates
+  temporal pulses and spatial stripes.
+\newblock Genes to Cells. 2005;10(11):1025--1038.
+\newblock doi:{https://doi.org/10.1111/j.1365-2443.2005.00897.x}.
+
+\bibitem{Li2019}
+Li L, Rispoli R, Patient R, Ciau-Uitz A, Porcher C.
+\newblock {Etv6 activates vegfa expression through positive and negative
+  transcriptional regulatory networks in Xenopus embryos}.
+\newblock Nature Communications. 2019;10(1).
+\newblock doi:{10.1038/s41467-019-09050-y}.
+
+\end{thebibliography}
 

### 2021-12-03 docs/manuscript/manuscript.tex
Write abstract in a single paragraph and cite "Materials and methods"

@@
 
-\subsection*{Background}
 A central problem in developmental and synthetic biology is understanding the
@@
 in systems biology as a tractable system to investigate the emergent properties
-of complex genotype-phenotype maps.
-
-\subsection*{Methods}
-In this study, we apply a Markov chain Monte Carlo (MCMC)-like algorithm to probe
+of complex genotype-phenotype maps. In this study, we apply a Markov chain
+Monte Carlo (MCMC)-like algorithm to probe
 the design space of three-node GRNs with the ability to generate a band-like
@@
 interesting and realizable morphogen interpretation mechanisms.
-
-\subsection*{Results}
 Out of 2061 GRNs selected using the search space algorithm, we found 714 classes of
@@
 Next to topological robustness, we investigated the robustness of the sampled
-network topologies to parameter changes (see Methods). Fig~\ref{fig:ab-rob}A
-illustrates the association between the abundance of a topology class versus its
+network topologies to parameter changes (see Materials and methods).
+Fig~\ref{fig:ab-rob}A illustrates the association
+between the abundance of a topology class versus its
 robustness. We did not find a significant correlation between

### 2021-12-04 docs/manuscript/manuscript.tex
Change S3 Fig A to S3A Fig and so on...

@@
 morphogen gradient in such a way that the initial and final concentration were
-constant in all the fields (\nameref{S3_Fig}A). In order to know if GRNs
+constant in all the fields (\hyperref[S3_Fig]{S3A Fig}). In order to know if GRNs
 with the same topology had a similar fitness in the different fields, we performed
@@
 only 4.66\%, 56.82\%, 56.14\% and 41.05\%, respectively, of the examined
-GRNs showed a fitness greater than 0.9 (\nameref{S3_Fig}B). In addition to
+GRNs showed a fitness greater than 0.9 (\hyperref[S3_Fig]{S3B Fig}). In addition to
 robustness, we wanted to investigate which of the GRNs could still produce
@@
 to changes in morphogenetic field size. However, GRNs with the same
-topology behave in different ways for each field size~(\nameref{S3_Fig}C),
+topology behave in different ways for each field size~(\hyperref[S3_Fig]{S3C Fig}),
 indicating that the specific values of network parameters are more important
@@
 The average fitness by topology is reported for morphogenetic fields with 10, 20,
-40 and 50 cells. Data from this table are shown in \nameref{S3_Fig}B.
+40 and 50 cells. Data from this table are shown in \hyperref[S3_Fig]{S3B Fig}.
 

### 2021-12-10 docs/manuscript/manuscript.tex
Move list of legends for supporting info files after references list
Remove financial information from Acknow...
Remove address from Affiliations

@@
 Boris A. Rodríguez\textsuperscript{2},
-Jayson Gutiérrez\textsuperscript{3},
+Jayson Gutiérrez\textsuperscript{3}
 \\
@@
 Instituto de Biología, Facultad de Ciencias Exactas y Naturales, Universidad de Antioquia
-UdeA, Calle 70 No. 52-21, Medellín, Colombia.
+UdeA, Medellín, Colombia.
 \\
@@
 Instituto de Física, Facultad de Ciencias Exactas y Naturales, Universidad de Antioquia
-UdeA, Calle 70 No. 52-21, Medellín, Colombia.
+UdeA, Medellín, Colombia.
 \\
@@
 
-%\section*{Conclusion}
+\section*{Acknowledgements}
+The computational resources (Stevin Supercomputer Infrastructure) and
+services used in this work were provided by the VSC (Flemish Supercomputer
+Center), funded by Ghent University, FWO and the Flemish Government–department EWI.
+All authors acknowledge the Colombian youth for their defense of a better
+country.
+
+\nolinenumbers
+
+\begin{thebibliography}{10}
+
+\bibitem{Turing1952}
+Turing AM.
+\newblock {The chemical basis of morphogenesis}.
+\newblock Bulletin of Mathematical Biology. 1952;52(1-2):153--197.
+\newblock doi:{10.1007/BF02459572}.
+
+\bibitem{Wolpert1969}
+Wolpert L.
+\newblock {Positional information and the spatial pattern of cellular
+  differentiation}.
+\newblock Journal of Theoretical Biology. 1969;25(1):1--47.
+\newblock doi:{https://doi.org/10.1016/S0022-5193(69)80016-0}.
+
+\bibitem{sharpe_2019}
+Sharpe J.
+\newblock Wolpert's {French} {Flag}: what's the problem?
+\newblock Development (Cambridge, England). 2019;146(24).
+\newblock doi:{10.1242/dev.185967}.
+
+\bibitem{driever_gradient_1988}
+Driever W, Nüsslein-Volhard C.
+\newblock A gradient of bicoid protein in {Drosophila} embryos.
+\newblock Cell. 1988;54(1):83--93.
+\newblock doi:{10.1016/0092-8674(88)90182-1}.
+
+\bibitem{driever_bicoid_1988}
+Driever W, Nüsslein-Volhard C.
+\newblock The bicoid protein determines position in the {Drosophila} embryo in
+  a concentration-dependent manner.
+\newblock Cell. 1988;54(1):95--104.
+\newblock doi:{10.1016/0092-8674(88)90183-3}.
+
+\bibitem{affolter_decapentaplegic_2007}
+Affolter M, Basler K.
+\newblock The {Decapentaplegic} morphogen gradient: from pattern formation to
+  growth regulation.
+\newblock Nature Reviews Genetics. 2007;8(9):663--674.
+\newblock doi:{10.1038/nrg2166}.
+
+\bibitem{dessaud_dynamic_2010}
+Dessaud E, Ribes V, Balaskas N, Yang LL, Pierani A, Kicheva A, et~al.
+\newblock Dynamic {Assignment} and {Maintenance} of {Positional} {Identity} in
+  the {Ventral} {Neural} {Tube} by the {Morphogen} {Sonic} {Hedgehog}.
+\newblock PLOS Biology. 2010;8(6):e1000382.
+\newblock doi:{10.1371/journal.pbio.1000382}.
+
+\bibitem{cohen_theoretical_2014}
+Cohen M, Page KM, Perez-Carrasco R, Barnes CP, Briscoe J.
+\newblock A theoretical framework for the regulation of {Shh}
+  morphogen-controlled gene expression.
+\newblock Development. 2014;141(20):3868--3878.
+\newblock doi:{10.1242/dev.112573}.
+
+\bibitem{raspopovic_digit_2014}
+Raspopovic J, Marcon L, Russo L, Sharpe J.
+\newblock Digit patterning is controlled by a {Bmp}-{Sox9}-{Wnt} {Turing}
+  network modulated by morphogen gradients.
+\newblock Science. 2014;345(6196):566.
+\newblock doi:{10.1126/science.1252960}.
+
+\bibitem{Milo2002}
+Milo R, Shen-Orr SS, Itzkovitz S, Kashtan N, Chklovskii D, Alon U.
+\newblock {Network Motifs: Simple Building Blocks of Complex Networks}.
+\newblock Science. 2002;298(5594):824--827.
+\newblock doi:{10.1126/science.298.5594.824}.
+
+\bibitem{kitano_computational_2002}
+Kitano H.
+\newblock Computational systems biology.
+\newblock Nature. 2002;420(6912):206--210.
+\newblock doi:{10.1038/nature01254}.
+
+\bibitem{Kalir2005}
+Kalir S, Mangan S, Alon U. {A coherent feed-forward loop with a SUM input
+  function prolongs flagella expression in Escherichia coli}; 2005.
+
+\bibitem{Kalir2004}
+Kalir S, Alon U.
+\newblock {Using a quantitative blueprint to reprogram the dynamics of the
+  flagella gene network}.
+\newblock Cell. 2004;117(6):713--720.
+\newblock doi:{10.1016/j.cell.2004.05.010}.
+
+\bibitem{Mangan2003a}
+Mangan S, Zaslaver A, Alon U.
+\newblock {The coherent feedforward loop serves as a sign-sensitive delay
+  element in transcription networks.}
+\newblock Journal of molecular biology. 2003;334(2):197--204.
+
+\bibitem{ODonnell2005}
+O'Donnell KA, Wentzel EA, Zeller KI, Dang CV, Mendell JT.
+\newblock {c-Myc-regulated microRNAs modulate E2F1 expression.}
+\newblock Nature. 2005;435(7043):839--843.
+\newblock doi:{10.1038/nature03677}.
+
+\bibitem{Basu2005}
+Basu S, Gerchman Y, Collins CH, Arnold FH, Weiss R.
+\newblock {A synthetic multicellular system for programmed pattern formation}.
+\newblock Nature. 2005;434(7037):1130--1134.
+\newblock doi:{10.1038/nature03461}.
+
+\bibitem{atkinson_development_2003}
+Atkinson MR, Savageau MA, Myers JT, Ninfa AJ.
+\newblock Development of genetic circuitry exhibiting toggle switch or
+  oscillatory behavior in {Escherichia} coli.
+\newblock Cell. 2003;113(5):597--607.
+\newblock doi:{10.1016/s0092-8674(03)00346-5}.
+
+\bibitem{friedland_synthetic_2009}
+Friedland AE, Lu TK, Wang X, Shi D, Church G, Collins JJ.
+\newblock Synthetic gene networks that count.
+\newblock Science (New York, NY). 2009;324(5931):1199--1202.
+\newblock doi:{10.1126/science.1172005}.
+
+\bibitem{gardner2012}
+Gardner L, Deiters A.
+\newblock Light-controlled synthetic gene circuits.
+\newblock Current Opinion in Chemical Biology. 2012;16(3-4):292--299.
+\newblock doi:{10.1016/j.cbpa.2012.04.010}.
+
+\bibitem{levskaya2005}
+Levskaya A, Chevalier AA, Tabor JJ, Simpson ZB, Lavery LA, Levy M, et~al.
+\newblock Engineering {Escherichia} coli to see light.
+\newblock Nature. 2005;438(7067):441--442.
+\newblock doi:{10.1038/nature04405}.
+
+\bibitem{santos-moreno_using_2019}
+Santos-Moreno J, Schaerli Y.
+\newblock Using {Synthetic} {Biology} to {Engineer} {Spatial} {Patterns}.
+\newblock Advanced Biosystems. 2019;3(4):1800280.
+\newblock doi:{https://doi.org/10.1002/adbi.201800280}.
+
+\bibitem{karlsson_therapeutic_2012}
+Karlsson M, Weber W.
+\newblock Therapeutic synthetic gene networks.
+\newblock Current Opinion in Biotechnology. 2012;23(5):703--711.
+\newblock doi:{10.1016/j.copbio.2012.01.003}.
+
+\bibitem{higashikuni_advancing_2017}
+Higashikuni Y, Chen WC, Lu TK.
+\newblock Advancing therapeutic applications of synthetic gene circuits.
+\newblock Current Opinion in Biotechnology. 2017;47:133--141.
+\newblock doi:{10.1016/j.copbio.2017.06.011}.
+
+\bibitem{abil_synthetic_2015}
+Abil Z, Xiong X, Zhao H.
+\newblock Synthetic biology for therapeutic applications.
+\newblock Molecular Pharmaceutics. 2015;12(2):322--331.
+\newblock doi:{10.1021/mp500392q}.
+
+\bibitem{healy_genetic_2019}
+Healy CP, Deans TL.
+\newblock Genetic circuits to engineer tissues with alternative functions.
+\newblock Journal of Biological Engineering. 2019;13:39.
+\newblock doi:{10.1186/s13036-019-0170-7}.
+
+\bibitem{kitada_programming_2018}
+Kitada T, DiAndreth B, Teague B, Weiss R.
+\newblock Programming gene and engineered-cell therapies with synthetic
+  biology.
+\newblock Science (New York, NY). 2018;359(6376).
+\newblock doi:{10.1126/science.aad1067}.
+
+\bibitem{munteanu_2014}
+Munteanu A, Cotterell J, Solé RV, Sharpe J.
+\newblock Design principles of stripe-forming motifs: the role of positive
+  feedback.
+\newblock Scientific Reports. 2014;4(1):5003.
+\newblock doi:{10.1038/srep05003}.
+
+\bibitem{Cotterell2010}
+Cotterell J, Sharpe J.
+\newblock {An atlas of gene regulatory networks reveals multiple three-gene
+  mechanisms for interpreting morphogen gradients}.
+\newblock Molecular Systems Biology. 2010;6(425):1--14.
+\newblock doi:{10.1038/msb.2010.74}.
+
+\bibitem{vonDassow2000}
+von Dassow G, Meir E, Munro EM, Odell GM.
+\newblock The segment polarity network is a robust developmental module.
+\newblock Nature. 2000;406(6792):188--192.
+\newblock doi:{10.1038/35018085}.
+
+\bibitem{Reinitz1995}
+Reinitz J, Mjolsness E, Sharp DH.
+\newblock {Model for cooperative control of positional information in
+  Drosophila by bicoid and maternal hunchback}.
+\newblock Journal of Experimental Zoology. 1995;271(1):47--56.
+\newblock doi:{10.1002/jez.1402710106}.
+
+\bibitem{Jaeger2004}
+Jaeger J, Blagov M, Kosman D, Kozlov KN, Manu, Myasnikova E, et~al.
+\newblock {Dynamical analysis of regulatory interactions in the gap gene system
+  of Drosophila melanogaster}.
+\newblock Genetics. 2004;167(4):1721--1737.
+\newblock doi:{10.1534/genetics.104.027334}.
+
+\bibitem{Schaerli2018}
+Schaerli Y, Jim{\'{e}}nez A, Duarte JM, Mihajlovic L, Renggli J, Isalan M,
+  et~al.
+\newblock {Synthetic circuits reveal how mechanisms of gene regulatory networks
+  constrain evolution}.
+\newblock Molecular Systems Biology. 2018;14(9):e8102.
+\newblock doi:{10.15252/msb.20178102}.
+
+\bibitem{Elowitz2000}
+Elowitz MB, Leibier S.
+\newblock {A synthetic oscillatory network of transcriptional regulators}.
+\newblock Nature. 2000;403(6767):335--338.
+\newblock doi:{10.1038/35002125}.
+
+\bibitem{barbier2020}
+Barbier I, Perez-Carrasco R, Schaerli Y.
+\newblock Controlling spatiotemporal pattern formation in a concentration
+  gradient with a synthetic toggle switch.
+\newblock Molecular Systems Biology. 2020;16(6):e9361.
+\newblock doi:{10.15252/msb.20199361}.
+
+\bibitem{Schaerli2014}
+Schaerli Y, Munteanu A, Gili M, Cotterell J, Sharpe J, Isalan M.
+\newblock {A unified design space of synthetic stripe-forming networks}.
+\newblock Nature Communications. 2014;5(May):4905.
+\newblock doi:{10.1038/ncomms5905}.
+
+\bibitem{balaskas2012}
+Balaskas N, Ribeiro A, Panovska J, Dessaud E, Sasai N, Page KM, et~al.
+\newblock Gene {Regulatory} {Logic} for {Reading} the {Sonic} {Hedgehog}
+  {Signaling} {Gradient} in the {Vertebrate} {Neural} {Tube}.
+\newblock Cell. 2012;148(1):273--284.
+\newblock doi:{10.1016/j.cell.2011.10.047}.
+
+\bibitem{exelby_2021}
+Exelby K, Herrera-Delgado E, Perez LG, Perez-Carrasco R, Sagner A, Metzis V,
+  et~al.
+\newblock Precision of tissue patterning is controlled by dynamical properties
+  of gene regulatory networks.
+\newblock Development. 2021;148(dev197566).
+\newblock doi:{10.1242/dev.197566}.
+
+\bibitem{ricci2011}
+Ricci F, Vallée-Bélisle A, Plaxco KW.
+\newblock High-precision, in vitro validation of the sequestration mechanism
+  for generating ultrasensitive dose-response curves in regulatory networks.
+\newblock PLoS computational biology. 2011;7(10):e1002171.
+\newblock doi:{10.1371/journal.pcbi.1002171}.
+
+\bibitem{mendoza2006}
+Mendoza L, Xenarios I.
+\newblock A method for the generation of standardized qualitative dynamical
+  systems of regulatory networks.
+\newblock Theoretical Biology \& Medical Modelling. 2006;3:13.
+\newblock doi:{10.1186/1742-4682-3-13}.
+
+\bibitem{vu2007}
+Vu TT, Vohradsky J.
+\newblock Nonlinear differential equation model for quantification of
+  transcriptional regulation applied to microarray data of {Saccharomyces}
+  cerevisiae.
+\newblock Nucleic Acids Research. 2007;35(1):279--287.
+\newblock doi:{10.1093/nar/gkl1001}.
+
+\bibitem{zhang2013}
+Zhang Q, Bhattacharya S, Andersen ME.
+\newblock Ultrasensitive response motifs: basic amplifiers in molecular
+  signalling networks.
+\newblock Open Biology. 2013;3(4):130031.
+\newblock doi:{10.1098/rsob.130031}.
+
+\bibitem{armao2016}
+Armao JJ, Lehn JM.
+\newblock Nonlinear {Kinetic} {Behavior} in {Constitutional} {Dynamic}
+  {Reaction} {Networks}.
+\newblock Journal of the American Chemical Society. 2016;138(51):16809--16814.
+\newblock doi:{10.1021/jacs.6b11107}.
+
+\bibitem{chen2005}
+Chen KC, Wang TY, Tseng HH, Huang CYF, Kao CY.
+\newblock A stochastic differential equation model for quantifying
+  transcriptional regulatory network in {Saccharomyces} cerevisiae.
+\newblock Bioinformatics (Oxford, England). 2005;21(12):2883--2890.
+\newblock doi:{10.1093/bioinformatics/bti415}.
+
+\bibitem{dahlquist2015}
+Dahlquist KD, Fitzpatrick BG, Camacho ET, Entzminger SD, Wanner NC.
+\newblock Parameter {Estimation} for {Gene} {Regulatory} {Networks} from
+  {Microarray} {Data}: {Cold} {Shock} {Response} in {Saccharomyces} cerevisiae.
+\newblock Bulletin of Mathematical Biology. 2015;77(8):1457--1492.
+\newblock doi:{10.1007/s11538-015-0092-6}.
+
+\bibitem{skbio2020}
+scikit-bio~development team T. scikit-bio: A Bioinformatics Library for Data
+  Scientists, Students, and Developers; 2020.
+\newblock Available from: \url{http://scikit-bio.org}.
+
+\bibitem{D.2005}
+Bonchev DG, Rouvray DH.
+\newblock {Complexity in Chemistry, Biology, and Ecology, ser. Mathematical and
+  Computational Chemistry.}
+\newblock Springer; 2005.
+
+\bibitem{sklearn}
+Pedregosa F, Varoquaux G, Gramfort A, Michel V, Thirion B, Grisel O, et~al.
+\newblock Scikit-learn: Machine Learning in {P}ython.
+\newblock Journal of Machine Learning Research. 2011;12:2825--2830.
+
+\bibitem{Erdos1959}
+Erd{\"{o}}s P, R{\'{e}}nyi a.
+\newblock {On random graphs}.
+\newblock Publicationes Mathematicae. 1959;6:290--297.
+\newblock doi:{10.2307/1999405}.
+
+\bibitem{Watts1998}
+Watts DJ, Strogatz SH.
+\newblock {Collective dynamics of 'small-world' networks.}
+\newblock Nature. 1998;393(6684):440--2.
+\newblock doi:{10.1038/30918}.
+
+\bibitem{giererMeinh1972}
+Gierer A, Meinhardt H.
+\newblock A theory of biological pattern formation.
+\newblock Kybernetik. 1972;12(1):30--39.
+\newblock doi:{10.1007/BF00289234}.
+
+\bibitem{meinhardt2000}
+Meinhardt H, Gierer A.
+\newblock Pattern formation by local self-activation and lateral inhibition.
+\newblock BioEssays: News and Reviews in Molecular, Cellular and Developmental
+  Biology. 2000;22(8):753--760.
+\newblock doi:{10.1002/1521-1878(200008)22:8<753::AID-BIES9>3.0.CO;2-Z}.
+
+\bibitem{shoaf1984}
+Shoaf SA, Conway K, Hunt RK.
+\newblock Application of reaction-diffusion models to cell patterning in
+  {Xenopus} retina. {Initiation} of patterns and their biological stability.
+\newblock Journal of Theoretical Biology. 1984;109(3):299--329.
+\newblock doi:{10.1016/s0022-5193(84)80085-5}.
+
+\bibitem{ishihara_2006}
+Ishihara S, Kaneko K.
+\newblock Turing pattern with proportion preservation.
+\newblock Journal of Theoretical Biology. 2006;238(3):683--693.
+\newblock doi:{10.1016/j.jtbi.2005.06.016}.
+
+\bibitem{kitano2004}
+Kitano H.
+\newblock Biological robustness.
+\newblock Nature Reviews Genetics. 2004;5(11):826--837.
+\newblock doi:{10.1038/nrg1471}.
+
+\bibitem{lo_robust_2015}
+Lo WC, Zhou S, Wan FYM, Lander AD, Nie Q.
+\newblock Robust and precise morphogen-mediated patterning: trade-offs,
+  constraints and mechanisms.
+\newblock Journal of The Royal Society Interface. 2015;12(102):20141041.
+\newblock doi:{10.1098/rsif.2014.1041}.
+
+\bibitem{wagner_robustness_2008}
+Wagner A.
+\newblock Robustness and evolvability: a paradox resolved.
+\newblock Proceedings of the Royal Society B: Biological Sciences.
+  2008;275(1630):91--100.
+\newblock doi:{10.1098/rspb.2007.1137}.
+
+\bibitem{ishihara_cross_2005}
+Ishihara S, Fujimoto K, Shibata T.
+\newblock Cross talking of network motifs in gene regulation that generates
+  temporal pulses and spatial stripes.
+\newblock Genes to Cells. 2005;10(11):1025--1038.
+\newblock doi:{https://doi.org/10.1111/j.1365-2443.2005.00897.x}.
+
+\bibitem{Li2019}
+Li L, Rispoli R, Patient R, Ciau-Uitz A, Porcher C.
+\newblock {Etv6 activates vegfa expression through positive and negative
+  transcriptional regulatory networks in Xenopus embryos}.
+\newblock Nature Communications. 2019;10(1).
+\newblock doi:{10.1038/s41467-019-09050-y}.
+
+\end{thebibliography}
 
@@
 
-\section*{Acknowledgements}
-The computational resources (Stevin Supercomputer Infrastructure) and
-services used in this work were provided by the VSC (Flemish Supercomputer
-Center), funded by Ghent University, FWO and the Flemish Government–department EWI.
-All authors acknowledge partial support from CODI-Universidad de Antioquia under
-project ``Estudio de la Evolución de Circuitos de Regulación del Desarollo
-Embrionario a través de la Biología Evolutiva de Sistemas'' code 2017-14367. All
-authors acknowledge the Colombian youth for their defense of a better country.
-
-\nolinenumbers
-
-\begin{thebibliography}{10}
-
-\bibitem{Turing1952}
-Turing AM.
-\newblock {The chemical basis of morphogenesis}.
-\newblock Bulletin of Mathematical Biology. 1952;52(1-2):153--197.
-\newblock doi:{10.1007/BF02459572}.
-
-\bibitem{Wolpert1969}
-Wolpert L.
-\newblock {Positional information and the spatial pattern of cellular
-  differentiation}.
-\newblock Journal of Theoretical Biology. 1969;25(1):1--47.
-\newblock doi:{https://doi.org/10.1016/S0022-5193(69)80016-0}.
-
-\bibitem{sharpe_2019}
-Sharpe J.
-\newblock Wolpert's {French} {Flag}: what's the problem?
-\newblock Development (Cambridge, England). 2019;146(24).
-\newblock doi:{10.1242/dev.185967}.
-
-\bibitem{driever_gradient_1988}
-Driever W, Nüsslein-Volhard C.
-\newblock A gradient of bicoid protein in {Drosophila} embryos.
-\newblock Cell. 1988;54(1):83--93.
-\newblock doi:{10.1016/0092-8674(88)90182-1}.
-
-\bibitem{driever_bicoid_1988}
-Driever W, Nüsslein-Volhard C.
-\newblock The bicoid protein determines position in the {Drosophila} embryo in
-  a concentration-dependent manner.
-\newblock Cell. 1988;54(1):95--104.
-\newblock doi:{10.1016/0092-8674(88)90183-3}.
-
-\bibitem{affolter_decapentaplegic_2007}
-Affolter M, Basler K.
-\newblock The {Decapentaplegic} morphogen gradient: from pattern formation to
-  growth regulation.
-\newblock Nature Reviews Genetics. 2007;8(9):663--674.
-\newblock doi:{10.1038/nrg2166}.
-
-\bibitem{dessaud_dynamic_2010}
-Dessaud E, Ribes V, Balaskas N, Yang LL, Pierani A, Kicheva A, et~al.
-\newblock Dynamic {Assignment} and {Maintenance} of {Positional} {Identity} in
-  the {Ventral} {Neural} {Tube} by the {Morphogen} {Sonic} {Hedgehog}.
-\newblock PLOS Biology. 2010;8(6):e1000382.
-\newblock doi:{10.1371/journal.pbio.1000382}.
-
-\bibitem{cohen_theoretical_2014}
-Cohen M, Page KM, Perez-Carrasco R, Barnes CP, Briscoe J.
-\newblock A theoretical framework for the regulation of {Shh}
-  morphogen-controlled gene expression.
-\newblock Development. 2014;141(20):3868--3878.
-\newblock doi:{10.1242/dev.112573}.
-
-\bibitem{raspopovic_digit_2014}
-Raspopovic J, Marcon L, Russo L, Sharpe J.
-\newblock Digit patterning is controlled by a {Bmp}-{Sox9}-{Wnt} {Turing}
-  network modulated by morphogen gradients.
-\newblock Science. 2014;345(6196):566.
-\newblock doi:{10.1126/science.1252960}.
-
-\bibitem{Milo2002}
-Milo R, Shen-Orr SS, Itzkovitz S, Kashtan N, Chklovskii D, Alon U.
-\newblock {Network Motifs: Simple Building Blocks of Complex Networks}.
-\newblock Science. 2002;298(5594):824--827.
-\newblock doi:{10.1126/science.298.5594.824}.
-
-\bibitem{kitano_computational_2002}
-Kitano H.
-\newblock Computational systems biology.
-\newblock Nature. 2002;420(6912):206--210.
-\newblock doi:{10.1038/nature01254}.
-
-\bibitem{Kalir2005}
-Kalir S, Mangan S, Alon U. {A coherent feed-forward loop with a SUM input
-  function prolongs flagella expression in Escherichia coli}; 2005.
-
-\bibitem{Kalir2004}
-Kalir S, Alon U.
-\newblock {Using a quantitative blueprint to reprogram the dynamics of the
-  flagella gene network}.
-\newblock Cell. 2004;117(6):713--720.
-\newblock doi:{10.1016/j.cell.2004.05.010}.
-
-\bibitem{Mangan2003a}
-Mangan S, Zaslaver A, Alon U.
-\newblock {The coherent feedforward loop serves as a sign-sensitive delay
-  element in transcription networks.}
-\newblock Journal of molecular biology. 2003;334(2):197--204.
-
-\bibitem{ODonnell2005}
-O'Donnell KA, Wentzel EA, Zeller KI, Dang CV, Mendell JT.
-\newblock {c-Myc-regulated microRNAs modulate E2F1 expression.}
-\newblock Nature. 2005;435(7043):839--843.
-\newblock doi:{10.1038/nature03677}.
-
-\bibitem{Basu2005}
-Basu S, Gerchman Y, Collins CH, Arnold FH, Weiss R.
-\newblock {A synthetic multicellular system for programmed pattern formation}.
-\newblock Nature. 2005;434(7037):1130--1134.
-\newblock doi:{10.1038/nature03461}.
-
-\bibitem{atkinson_development_2003}
-Atkinson MR, Savageau MA, Myers JT, Ninfa AJ.
-\newblock Development of genetic circuitry exhibiting toggle switch or
-  oscillatory behavior in {Escherichia} coli.
-\newblock Cell. 2003;113(5):597--607.
-\newblock doi:{10.1016/s0092-8674(03)00346-5}.
-
-\bibitem{friedland_synthetic_2009}
-Friedland AE, Lu TK, Wang X, Shi D, Church G, Collins JJ.
-\newblock Synthetic gene networks that count.
-\newblock Science (New York, NY). 2009;324(5931):1199--1202.
-\newblock doi:{10.1126/science.1172005}.
-
-\bibitem{gardner2012}
-Gardner L, Deiters A.
-\newblock Light-controlled synthetic gene circuits.
-\newblock Current Opinion in Chemical Biology. 2012;16(3-4):292--299.
-\newblock doi:{10.1016/j.cbpa.2012.04.010}.
-
-\bibitem{levskaya2005}
-Levskaya A, Chevalier AA, Tabor JJ, Simpson ZB, Lavery LA, Levy M, et~al.
-\newblock Engineering {Escherichia} coli to see light.
-\newblock Nature. 2005;438(7067):441--442.
-\newblock doi:{10.1038/nature04405}.
-
-\bibitem{santos-moreno_using_2019}
-Santos-Moreno J, Schaerli Y.
-\newblock Using {Synthetic} {Biology} to {Engineer} {Spatial} {Patterns}.
-\newblock Advanced Biosystems. 2019;3(4):1800280.
-\newblock doi:{https://doi.org/10.1002/adbi.201800280}.
-
-\bibitem{karlsson_therapeutic_2012}
-Karlsson M, Weber W.
-\newblock Therapeutic synthetic gene networks.
-\newblock Current Opinion in Biotechnology. 2012;23(5):703--711.
-\newblock doi:{10.1016/j.copbio.2012.01.003}.
-
-\bibitem{higashikuni_advancing_2017}
-Higashikuni Y, Chen WC, Lu TK.
-\newblock Advancing therapeutic applications of synthetic gene circuits.
-\newblock Current Opinion in Biotechnology. 2017;47:133--141.
-\newblock doi:{10.1016/j.copbio.2017.06.011}.
-
-\bibitem{abil_synthetic_2015}
-Abil Z, Xiong X, Zhao H.
-\newblock Synthetic biology for therapeutic applications.
-\newblock Molecular Pharmaceutics. 2015;12(2):322--331.
-\newblock doi:{10.1021/mp500392q}.
-
-\bibitem{healy_genetic_2019}
-Healy CP, Deans TL.
-\newblock Genetic circuits to engineer tissues with alternative functions.
-\newblock Journal of Biological Engineering. 2019;13:39.
-\newblock doi:{10.1186/s13036-019-0170-7}.
-
-\bibitem{kitada_programming_2018}
-Kitada T, DiAndreth B, Teague B, Weiss R.
-\newblock Programming gene and engineered-cell therapies with synthetic
-  biology.
-\newblock Science (New York, NY). 2018;359(6376).
-\newblock doi:{10.1126/science.aad1067}.
-
-\bibitem{munteanu_2014}
-Munteanu A, Cotterell J, Solé RV, Sharpe J.
-\newblock Design principles of stripe-forming motifs: the role of positive
-  feedback.
-\newblock Scientific Reports. 2014;4(1):5003.
-\newblock doi:{10.1038/srep05003}.
-
-\bibitem{Cotterell2010}
-Cotterell J, Sharpe J.
-\newblock {An atlas of gene regulatory networks reveals multiple three-gene
-  mechanisms for interpreting morphogen gradients}.
-\newblock Molecular Systems Biology. 2010;6(425):1--14.
-\newblock doi:{10.1038/msb.2010.74}.
-
-\bibitem{vonDassow2000}
-von Dassow G, Meir E, Munro EM, Odell GM.
-\newblock The segment polarity network is a robust developmental module.
-\newblock Nature. 2000;406(6792):188--192.
-\newblock doi:{10.1038/35018085}.
-
-\bibitem{Reinitz1995}
-Reinitz J, Mjolsness E, Sharp DH.
-\newblock {Model for cooperative control of positional information in
-  Drosophila by bicoid and maternal hunchback}.
-\newblock Journal of Experimental Zoology. 1995;271(1):47--56.
-\newblock doi:{10.1002/jez.1402710106}.
-
-\bibitem{Jaeger2004}
-Jaeger J, Blagov M, Kosman D, Kozlov KN, Manu, Myasnikova E, et~al.
-\newblock {Dynamical analysis of regulatory interactions in the gap gene system
-  of Drosophila melanogaster}.
-\newblock Genetics. 2004;167(4):1721--1737.
-\newblock doi:{10.1534/genetics.104.027334}.
-
-\bibitem{Schaerli2018}
-Schaerli Y, Jim{\'{e}}nez A, Duarte JM, Mihajlovic L, Renggli J, Isalan M,
-  et~al.
-\newblock {Synthetic circuits reveal how mechanisms of gene regulatory networks
-  constrain evolution}.
-\newblock Molecular Systems Biology. 2018;14(9):e8102.
-\newblock doi:{10.15252/msb.20178102}.
-
-\bibitem{Elowitz2000}
-Elowitz MB, Leibier S.
-\newblock {A synthetic oscillatory network of transcriptional regulators}.
-\newblock Nature. 2000;403(6767):335--338.
-\newblock doi:{10.1038/35002125}.
-
-\bibitem{barbier2020}
-Barbier I, Perez-Carrasco R, Schaerli Y.
-\newblock Controlling spatiotemporal pattern formation in a concentration
-  gradient with a synthetic toggle switch.
-\newblock Molecular Systems Biology. 2020;16(6):e9361.
-\newblock doi:{10.15252/msb.20199361}.
-
-\bibitem{Schaerli2014}
-Schaerli Y, Munteanu A, Gili M, Cotterell J, Sharpe J, Isalan M.
-\newblock {A unified design space of synthetic stripe-forming networks}.
-\newblock Nature Communications. 2014;5(May):4905.
-\newblock doi:{10.1038/ncomms5905}.
-
-\bibitem{balaskas2012}
-Balaskas N, Ribeiro A, Panovska J, Dessaud E, Sasai N, Page KM, et~al.
-\newblock Gene {Regulatory} {Logic} for {Reading} the {Sonic} {Hedgehog}
-  {Signaling} {Gradient} in the {Vertebrate} {Neural} {Tube}.
-\newblock Cell. 2012;148(1):273--284.
-\newblock doi:{10.1016/j.cell.2011.10.047}.
-
-\bibitem{exelby_2021}
-Exelby K, Herrera-Delgado E, Perez LG, Perez-Carrasco R, Sagner A, Metzis V,
-  et~al.
-\newblock Precision of tissue patterning is controlled by dynamical properties
-  of gene regulatory networks.
-\newblock Development. 2021;148(dev197566).
-\newblock doi:{10.1242/dev.197566}.
-
-\bibitem{ricci2011}
-Ricci F, Vallée-Bélisle A, Plaxco KW.
-\newblock High-precision, in vitro validation of the sequestration mechanism
-  for generating ultrasensitive dose-response curves in regulatory networks.
-\newblock PLoS computational biology. 2011;7(10):e1002171.
-\newblock doi:{10.1371/journal.pcbi.1002171}.
-
-\bibitem{mendoza2006}
-Mendoza L, Xenarios I.
-\newblock A method for the generation of standardized qualitative dynamical
-  systems of regulatory networks.
-\newblock Theoretical Biology \& Medical Modelling. 2006;3:13.
-\newblock doi:{10.1186/1742-4682-3-13}.
-
-\bibitem{vu2007}
-Vu TT, Vohradsky J.
-\newblock Nonlinear differential equation model for quantification of
-  transcriptional regulation applied to microarray data of {Saccharomyces}
-  cerevisiae.
-\newblock Nucleic Acids Research. 2007;35(1):279--287.
-\newblock doi:{10.1093/nar/gkl1001}.
-
-\bibitem{zhang2013}
-Zhang Q, Bhattacharya S, Andersen ME.
-\newblock Ultrasensitive response motifs: basic amplifiers in molecular
-  signalling networks.
-\newblock Open Biology. 2013;3(4):130031.
-\newblock doi:{10.1098/rsob.130031}.
-
-\bibitem{armao2016}
-Armao JJ, Lehn JM.
-\newblock Nonlinear {Kinetic} {Behavior} in {Constitutional} {Dynamic}
-  {Reaction} {Networks}.
-\newblock Journal of the American Chemical Society. 2016;138(51):16809--16814.
-\newblock doi:{10.1021/jacs.6b11107}.
-
-\bibitem{chen2005}
-Chen KC, Wang TY, Tseng HH, Huang CYF, Kao CY.
-\newblock A stochastic differential equation model for quantifying
-  transcriptional regulatory network in {Saccharomyces} cerevisiae.
-\newblock Bioinformatics (Oxford, England). 2005;21(12):2883--2890.
-\newblock doi:{10.1093/bioinformatics/bti415}.
-
-\bibitem{dahlquist2015}
-Dahlquist KD, Fitzpatrick BG, Camacho ET, Entzminger SD, Wanner NC.
-\newblock Parameter {Estimation} for {Gene} {Regulatory} {Networks} from
-  {Microarray} {Data}: {Cold} {Shock} {Response} in {Saccharomyces} cerevisiae.
-\newblock Bulletin of Mathematical Biology. 2015;77(8):1457--1492.
-\newblock doi:{10.1007/s11538-015-0092-6}.
-
-\bibitem{skbio2020}
-scikit-bio~development team T. scikit-bio: A Bioinformatics Library for Data
-  Scientists, Students, and Developers; 2020.
-\newblock Available from: \url{http://scikit-bio.org}.
-
-\bibitem{D.2005}
-Bonchev DG, Rouvray DH.
-\newblock {Complexity in Chemistry, Biology, and Ecology, ser. Mathematical and
-  Computational Chemistry.}
-\newblock Springer; 2005.
-
-\bibitem{sklearn}
-Pedregosa F, Varoquaux G, Gramfort A, Michel V, Thirion B, Grisel O, et~al.
-\newblock Scikit-learn: Machine Learning in {P}ython.
-\newblock Journal of Machine Learning Research. 2011;12:2825--2830.
-
-\bibitem{Erdos1959}
-Erd{\"{o}}s P, R{\'{e}}nyi a.
-\newblock {On random graphs}.
-\newblock Publicationes Mathematicae. 1959;6:290--297.
-\newblock doi:{10.2307/1999405}.
-
-\bibitem{Watts1998}
-Watts DJ, Strogatz SH.
-\newblock {Collective dynamics of 'small-world' networks.}
-\newblock Nature. 1998;393(6684):440--2.
-\newblock doi:{10.1038/30918}.
-
-\bibitem{giererMeinh1972}
-Gierer A, Meinhardt H.
-\newblock A theory of biological pattern formation.
-\newblock Kybernetik. 1972;12(1):30--39.
-\newblock doi:{10.1007/BF00289234}.
-
-\bibitem{meinhardt2000}
-Meinhardt H, Gierer A.
-\newblock Pattern formation by local self-activation and lateral inhibition.
-\newblock BioEssays: News and Reviews in Molecular, Cellular and Developmental
-  Biology. 2000;22(8):753--760.
-\newblock doi:{10.1002/1521-1878(200008)22:8<753::AID-BIES9>3.0.CO;2-Z}.
-
-\bibitem{shoaf1984}
-Shoaf SA, Conway K, Hunt RK.
-\newblock Application of reaction-diffusion models to cell patterning in
-  {Xenopus} retina. {Initiation} of patterns and their biological stability.
-\newblock Journal of Theoretical Biology. 1984;109(3):299--329.
-\newblock doi:{10.1016/s0022-5193(84)80085-5}.
-
-\bibitem{ishihara_2006}
-Ishihara S, Kaneko K.
-\newblock Turing pattern with proportion preservation.
-\newblock Journal of Theoretical Biology. 2006;238(3):683--693.
-\newblock doi:{10.1016/j.jtbi.2005.06.016}.
-
-\bibitem{kitano2004}
-Kitano H.
-\newblock Biological robustness.
-\newblock Nature Reviews Genetics. 2004;5(11):826--837.
-\newblock doi:{10.1038/nrg1471}.
-
-\bibitem{lo_robust_2015}
-Lo WC, Zhou S, Wan FYM, Lander AD, Nie Q.
-\newblock Robust and precise morphogen-mediated patterning: trade-offs,
-  constraints and mechanisms.
-\newblock Journal of The Royal Society Interface. 2015;12(102):20141041.
-\newblock doi:{10.1098/rsif.2014.1041}.
-
-\bibitem{wagner_robustness_2008}
-Wagner A.
-\newblock Robustness and evolvability: a paradox resolved.
-\newblock Proceedings of the Royal Society B: Biological Sciences.
-  2008;275(1630):91--100.
-\newblock doi:{10.1098/rspb.2007.1137}.
-
-\bibitem{ishihara_cross_2005}
-Ishihara S, Fujimoto K, Shibata T.
-\newblock Cross talking of network motifs in gene regulation that generates
-  temporal pulses and spatial stripes.
-\newblock Genes to Cells. 2005;10(11):1025--1038.
-\newblock doi:{https://doi.org/10.1111/j.1365-2443.2005.00897.x}.
-
-\bibitem{Li2019}
-Li L, Rispoli R, Patient R, Ciau-Uitz A, Porcher C.
-\newblock {Etv6 activates vegfa expression through positive and negative
-  transcriptional regulatory networks in Xenopus embryos}.
-\newblock Nature Communications. 2019;10(1).
-\newblock doi:{10.1038/s41467-019-09050-y}.
-
-\end{thebibliography}
-
 \end{document}
